---
title: "2_CL_Data__Data_Cleaning__generates_AEI_No_NAs"
author: "Rachel Ledig"
date: "4/20/2021"
output: 
  html_document:
    toc: true
    toc_float: true
bibliography: /Volumes/RachelExternal/Thesis/Thesis/bib.bib
---
```{r include=FALSE}
knitr::opts_chunk$set(message = FALSE)
```


```{r Housekeeping, warning=FALSE, message=FALSE, include=FALSE}
library(brms)
library(rstan)
library(dplyr)
library(tidyr)
library(plotly)
library(ggplot2)
library(rethinking)
library(purrr)
library(gganimate)
library(wbstats)
library(viridis)
library(tm)
library(wbstats)
library(tibble)
library(stringr)
library(distill)
```


Here is the data necessary for this script:
```{r}
#original Siebert Data
aei <- read.csv2(
  "/Volumes/RachelExternal/Thesis/Data/SiebertData/HID_v10/Country_tenyear_ha.csv")
#import the data set with total country area, this comes from the Worldbank
countryarea <- read.csv2(
  "/Volumes/RachelExternal/Thesis/Data/LandArea/LandArea.csv")
#the ISO codes and regional data from Gapminder
iso <- read.csv2(
  "/Volumes/RachelExternal/Thesis/Thesis/Data/GAPminder_ISOcodes.csv")
#income here is 2011 usd, taken from Gapminder
income <- read.csv(
  "/Volumes/RachelExternal/Thesis/Thesis/Data/Gapminder_income_per_person_gdppercapita_ppp_inflation_adjusted.csv")
#population data also taken from Gapminder
population <- read.csv(
  "/Volumes/RachelExternal/Thesis/Thesis/Data/Gapminder_population_total.csv")
#cft taken from LPJmL output
cft <- read.csv(
  '/Volumes/RachelExternal/Thesis/Thesis/Data/cft_and_precipdata.csv')
#humidity, PET, and precipitation
hum_PET_prec <- read.csv(
  '/Volumes/RachelExternal/Thesis/Thesis/Data/hum_pet_prec.csv')
#Political system from our world in data
democracy <- read.csv("/Volumes/RachelExternal/Thesis/Data_upload_for_CL/OurWorldInData_political-regime-updated2016.csv")
#More political system data from center for systemic peace
syspeace <-
  read.csv2("/Volumes/RachelExternal/Thesis/Data_upload_for_CL/CenterForSystematicPeacePolity.csv")

```



Lets do a short introduction to what we're working with here. 

# Target Variable  
  
* $I$ - Irrigation Percentage - Downloaded the data from @siebertGlobalDataSet2015 data set in ha/country/timestep, then by dividing by the countries total area, the fraction of total land area that is irrigated in a given country is created. I will be working with data from 1910-2005. The first years of @siebertGlobalDataSet2015 are a bit off, and some of my data derived from LPJmL output only starts in 1901. To avoid further conflicts my study period will be 1910-2005 for the country level model.
```{r Setting up Country Level df, warning=FALSE}
aei <- as.data.frame(aei)

#rename the cols so that I can use them as number
colnames(aei) <- c("country", "ISO", "ID", "1900", "1910","1920","1930","1940","1950","1960","1970","1980","1985","1990","1995","2000","2005")

#tidying data to long format
aei <- pivot_longer(aei,cols = 4:17, names_to = "year", values_to = "aei_ha")
aei$aei_ha <- removePunctuation(aei$aei_ha)
aei$year <- as.numeric(aei$year)
aei$aei_ha <- as.numeric(aei$aei_ha)
aei$ISO <- as.factor(aei$ISO)
aei$country <- as.factor(aei$country)

#remove 1900
aei <- subset(aei, year != "1900")

#make a new col for year count
aei$yearcount <- (aei$year - 1910)

#remove data for the total global irrigated area
globeaei <- subset(aei, country == "WORLD")
aei= aei[!aei$country == "WORLD",]

str(aei)
```

      
Added to this to be able to calculate the % of land irrigated per country we need the country area. This data is taken from Worldbank.  

```{r message=FALSE, warning=FALSE}
#removing extra cols
countryarea <- countryarea[-c(265:270), ]
countryarea <- countryarea[, -c(1, 3:4)]

#fix the col names (in the two days since i found this little for loop I learned how to do this faster!)
for ( col in 2:ncol(countryarea)){
    colnames(countryarea)[col] <-  sub("X", "", colnames(countryarea)[col])
}
remove(col)

#making long data
countryarea <- 
  countryarea %>%  
  pivot_longer(!Country.Code, names_to = "year", values_to = "area_km") %>% 
  rename(ISO = Country.Code)

#fixing variable types
countryarea$ISO <- as.factor(countryarea$ISO)
countryarea$year <- as.numeric(countryarea$year)
str(countryarea)
```


For some reason, the data from Worldbank on each countries area only starts in 1961. For this analysis, data back to 1910 is needed. Country area from 1961 will be substituted for backwards, with the assumption that country area doesn't change drastically. Also here, merge the country area with the AEI df here.   

```{r Join CAr }
#combine these data sets
aei <- aei %>% merge(y = countryarea, by.x = c("ISO", "year"), all.x = TRUE)

#going to do the same thing as before so that I have area for all the other years
#assuming that political borders are the same as in the year 2005
aei <-
  aei %>% 
  group_by(ISO) %>% 
  fill(area_km, .direction = "up") %>% 
  ungroup()

rm(countryarea)
head(aei)
```

Some countries also have no area value. I'll go through and see if any of the countries with missing area values have non zero AEI_ha. If so, I'll replace them, as they didn't come through with the WorldBank data set but are online. Copying and pasting from various sources should be fine.


```{r Fixes for CAr}
#checking for countries with NA values in their area_km col
#looking primarily for countries that have non 0 AEI_ha
aei[is.na(aei$area_km),]
                
#fixing counties with no area
aei$area_km[which(aei$ISO == "GUF")] <- 83534 #French Guiana from Britannica
aei$area_km[which(aei$ISO == "GLP")] <- 1630 #Guadeloupe from Britannica 
aei$area_km[which(aei$ISO == "MTQ")] <- 1128 # Martinique from Britannica
aei$area_km[which(aei$ISO == "REU")] <- 2512 #Reunion from Britannica
aei$area_km[which(aei$ISO == "SMK")] <- 13812 + 10905 + 77589 #Montenegro + Kosovo + Serbia from Britannica
aei$area_km[which(aei$ISO == "TWN")] <- 35980 #Taiwan from CIA.gov
aei$area_km[which(aei$ISO == "SDN")] <- 1882000 #Sudan from UNDP
```

Now, transform the AEI_ha to $km^2$ and divide to get the % of total country area that is equipped for irrigation. There are still some countries that have 0 AEI_ha and no area_km value. I think I will assign the irrfrac/perc 0 to those countries, but lets wait a little while to do that, most are islands.  

```{r Creating Target Variable}
aei <- aei %>% mutate(irrperc = ((aei_ha/100)/area_km)*100, irrfrac = ((aei_ha/100)/area_km))
```

Here is a quick histogram of the target variable. A lot of 0s. 

```{r Hist of TVar}
hist(aei$irrperc, breaks = 200)
```

***

# Chosen Predictor Variables  

According to the FAO [@walkerFAOIrrigationDrainage1989], the choice of which irrigation, and by proxy whether to irrigate or not, comes from six factors. The predictor that will be used for this analysis are listed below:  

- Economic 

  - $M_c$ - GDP per Capita - Each country's GDP per Capita from Gapminder, which collects data from the UN and the Maddison project. This variable will be logged because as stated in *Statistical Rethinking* [@mcelreathStatisticalRethinkingBayesian2020] McElreath states (pg. 239), "We use the logarithm of it, because the logarithm of GDP is the *magnitude* of GDP. Since wealth generates wealth, it tends to be exponentially related to anything that increases it"
  
- Topographical   

  - $Ru_c$ - Represented here using the "rugged" data set from rethinking, in which a coeff of ruggedness for each country was assigned. This data was taken from @rileyTerrainRuggednessIndex1999
  
- Soil Type  

  - This will not be included in this analysis. Although, it could be useful there is no good way to quantify soil type on a country level basis. This will be included in the grid cell model, as regions can be easily integrated with the grid cell structure. 

- Water Supply    

  - $R_c$ - Total Precipitation -  Translated output from LPJmL. Summed across two dimensions: month and country. In $m^3$/country/year  
  - $H_c$ - Mean/Median Humidity - Translated output from LPJmL. Calculated from $Precip/PET$ for each grid cell then averaged (or median) for each country and year.  
  - $E_c$ - Mean/Median Potential Evapotranspiration - Translated output from LPJmL. Average and median were taken for each country and year.
- Crop Type  

  - Fraction of each crop grown per country -  Translated output from LPJmL. Individual crop fraction per grid cell multiplied by respective grid cell area, summed on a country level giving ha of each crop grown per country. This was then divided by total area of the country to achieve the % area occupied by a certain crop per country. Still working out how to work with this predictor. 

- Social Influences  

  - $P_c$ - Total Population - Each country's total population on a yearly basis. Taken from Gapminder. Again, this will be logged, due to the same argument for GDP, and centered.
  
All variables will be scaled or centered. 


First, the ISO codes and Regions to be able to hold some structure.  

```{r}

iso <- iso[, -c(2,7,10,12,13)] #removing extra info

#renaming some cols
iso <- 
  iso %>% 
  rename(ISO = GEO, country = name)

#fixing the variable types
iso$ISO <- as.factor(iso$ISO)
iso$country <- as.factor(iso$country)
iso$four_regions <- as.factor(iso$four_regions)
iso$eight_regions <- as.factor(iso$eight_regions)
iso$six_regions <- as.factor(iso$six_regions)
iso$six_regions <- as.factor(iso$six_regions)
iso$World.bank.region <- as.factor(iso$World.bank.region)
str(iso)

#past me helping future me
#fixing Macedonia, FYR to North Macedonia
iso$country <- as.character(iso$country)
iso[iso == "Macedonia, FYR"] <- "North Macedonia"
iso$country <- as.factor(iso$country)
```
Cool, we have multiple regions here that we could use in the future for some multilevel business. Also, latitude and longitude is here, which we can also try later. Obviously these lat and lon measurements will mean nothing for a country like China or the USA, but they are nice to have.    


## Income

Income data is coming from Gapminder, as they have put real effort in to filling all the gaps in income and population for the last 150 years. Income is in terms of the 2011 USD and is inflation adjusted.  

```{r}

income <- income[, c(1, 102:207)] #removing extra years

#making long data
income <- 
  income %>%  
  pivot_longer(!country, names_to = "year", values_to = "income") 

#still have the xs there
income$year <- str_remove_all(income$year, "[X]")
#transformation of the variable types
income$year <- as.numeric(income$year)
income$country <- as.factor(income$country)

#a summary
summary(income)

#past me helping future me again
#fixing Eswatini to Swaziland (even though Eswatini is the correct name)
income$country <- as.character(income$country)
income[income == "Eswatini"] <- "Swaziland"
income$country <- as.factor(income$country)

```


## Population

Population data is also taken from Gapminder. 
```{r}
#removing extra years
population <- population[, c(1, 102:207)] 
#making longer data
population <- 
  population %>%  
  pivot_longer(!country, names_to = "year", values_to = "population") 

#removing the Xs from the year strings
population$year <- str_remove_all(population$year, "[X]")

#fixing variable types
population$country <- as.factor(population$country)
population$year <- as.numeric(population$year)

summary(population)

#annnd again...
#fixing Eswatini to Swaziland (even though Eswatini is the correct name)
population$country <- as.character(population$country)
population[population == "Eswatini"] <- "Swaziland"
population$country <- as.factor(population$country)
```

So some issues arise here with the fact that we have different country data for each dataframe. Sometimes the names are capitalized, other times there are spaces instead of underscores. By using `anti_join()` I can see what won't be joined. Also this goes both ways, `anti_join()` returns rows of x that do not have a match in y.  

```{r}
isopop <- anti_join(iso, population, by = "country")
summary(isopop)

popiso <- anti_join(population, iso, by = "country")
summary(popiso)
```

Went back up and solved the issue of North Macedonia and Swaziland. There are still issues with Hong Kong and Taiwan as they are part of China. ISO has data for Hong Kong and Taiwan that population does not have.  I wont include them in the HTML but they are in the code.  

```{r include=FALSE}
isoinc <- anti_join(iso, income, by = "country")
summary(isoinc)

inciso <- anti_join(income, iso, by = "country")
summary(inciso)
```


```{r include=FALSE}
popinc <- anti_join(population, income, by = "country")
summary(popinc)

incpop <- anti_join(income, population, by = "country")
summary(incpop)
```


Alright, so I will left join ISO and population then again with income? Not sure that this will solve my problem, but lets try it.
```{r}
#remove all these rando tables
rm(popinc, inciso, incpop, popiso, isoinc, isopop)

#
iso <- left_join(iso, population, by = "country")
iso <- left_join(iso, income, by = c("country", "year"))

# alright 197 countries! whoop!
summary(iso)

rm(income, population)
```

Now, join ISO to AEI with a merge (cause in this case the dynamics of merge are clearer for me). Hoping to preserve all data! 

```{r}
aei <- merge(aei, iso, by = c("ISO","country", "year"), all.x = TRUE)
rm(iso)
summary(aei) 
str(aei)
```

Solved the issue, 232 factors for both countries and ISO, meaning that we have been using the original AEI file from Siebert as the main data frame to join to. Good. This is the data that needs to be maintained, NAs can pop up in other cols but not irrigated area.   

But I still would like total GDP for later so lets transform using population. 
```{r}
aei$income <- as.numeric(aei$income) 
aei$population <- as.numeric(aei$population)
#mutating for total GDP and log GDP
aei <-
  aei %>%
  mutate(GDPtot = income*population) 

str(aei)
summary(aei)
```



### Dealing with Some Issues

There are a lot of countries that have no data for the regions, but have non zero irrperc/frac which I will use in further analysis, so we need to replace the regions. 

```{r}
#view which countries don't have regions but have non zero irrprec
na_regions <- aei[is.na(aei$four_regions),]

#Aruba
aei <- aei %>%
  mutate(four_regions = replace(four_regions, ISO == 'ABW', 'americas'), 
         six_regions = replace(six_regions, ISO == 'ABW', 'america'), 
         eight_regions = replace(eight_regions, ISO == 'ABW', 'america_north')) %>%
#american samoa
  mutate(four_regions = replace(four_regions,ISO == 'ASM', 'asia'), 
         six_regions = replace(six_regions,ISO == 'ASM', 'east_asia_pacific'), 
         eight_regions = replace(eight_regions,ISO == 'ASM', 'east_asia_pacific')) %>%
#bermuda
  mutate(four_regions = replace(four_regions,ISO == 'BMU', 'americas'), 
         six_regions = replace(six_regions,ISO == 'BMU', 'america'), 
         eight_regions = replace(eight_regions,ISO == 'BMU', 'america_north')) %>% 
#cote d'ivoire
  mutate(four_regions = replace(four_regions,ISO == 'CIV', 'africa'), 
         six_regions = replace(six_regions,ISO == 'CIV', 'sub_saharan_africa'), 
         eight_regions = replace(eight_regions,ISO == 'CIV', 'africa_sub_saharan')) %>% 
#democratic republic of congo
  mutate(four_regions = replace(four_regions,ISO == 'COD', 'africa'), 
         six_regions = replace(six_regions,ISO == 'COD', 'sub_saharan_africa'), 
         eight_regions = replace(eight_regions,ISO == 'COD', 'africa_sub_saharan')) %>% 
#republic of congo
  mutate(four_regions = replace(four_regions,ISO == 'COG', 'africa'), 
         six_regions = replace(six_regions,ISO == 'COG', 'sub_saharan_africa'), 
         eight_regions = replace(eight_regions,ISO == 'COG', 'africa_sub_saharan')) %>% 
#cayman islands
  mutate(four_regions = replace(four_regions,ISO == 'CYM', 'americas'), 
         six_regions = replace(six_regions,ISO == 'CYM', 'america'), 
         eight_regions = replace(eight_regions,ISO == 'CYM', 'america_north')) %>%
#Faroe islands
  mutate(four_regions = replace(four_regions,ISO == 'FRO', 'europe'), 
         six_regions = replace(six_regions,ISO == 'FRO', 'europe_central_asia'), 
         eight_regions = replace(eight_regions,ISO == 'FRO', 'europe_west')) %>%
#micronesia
  mutate(four_regions = replace(four_regions,ISO == 'FSM', 'asia'), 
         six_regions = replace(six_regions,ISO == 'FSM', 'east_asia_pacific'), 
         eight_regions = replace(eight_regions,ISO == 'FSM', 'east_asia_pacific')) %>%
#gibraltar
  mutate(four_regions = replace(four_regions,ISO == 'GIB', 'europe'), 
         six_regions = replace(six_regions,ISO == 'GIB', 'europe_central_asia'), 
         eight_regions = replace(eight_regions,ISO == 'GIB', 'europe_west')) %>%
#Guadeloupe
  mutate(four_regions = replace(four_regions,ISO == 'GLP', 'americas'), 
         six_regions = replace(six_regions,ISO == 'GLP', 'america'), 
         eight_regions = replace(eight_regions,ISO == 'GLP', 'america_north')) %>%
#greenland
  mutate(four_regions = replace(four_regions,ISO == 'GRL', 'europe'), 
         six_regions = replace(six_regions,ISO == 'GRL', 'europe_central_asia'), 
         eight_regions = replace(eight_regions,ISO == 'GRL', 'europe_west')) %>%
#french Guiana 
  mutate(four_regions = replace(four_regions,ISO == 'GUF', 'americas'), 
         six_regions = replace(six_regions,ISO == 'GUF', 'america'), 
         eight_regions = replace(eight_regions,ISO == 'GUF', 'america_south')) %>%
#guam
  mutate(four_regions = replace(four_regions,ISO == 'GUM', 'asia'), 
         six_regions = replace(six_regions,ISO == 'GUM', 'east_asia_pacific'), 
         eight_regions = replace(eight_regions,ISO == 'GUM', 'east_asia_pacific')) %>%
#Kyrgyzstan
  mutate(four_regions = replace(four_regions,ISO == 'KGZ', 'asia'), 
         six_regions = replace(six_regions,ISO == 'KGZ', 'europe_central_asia'), 
         eight_regions = replace(eight_regions,ISO == 'KGZ', 'asia_west')) %>%
#laos
  mutate(four_regions = replace(four_regions,ISO == 'LAO', 'asia'), 
         six_regions = replace(six_regions,ISO == 'LAO', 'east_asia_pacific'), 
         eight_regions = replace(eight_regions,ISO == 'LAO', 'east_asia_pacific')) %>%
#Macedonia
  mutate(four_regions = replace(four_regions,ISO == 'MKD', 'europe'), 
         six_regions = replace(six_regions,ISO == 'MKD', 'europe_central_asia'), 
         eight_regions = replace(eight_regions,ISO == 'MKD', 'europe_east')) %>%
#Martinique 
  mutate(four_regions = replace(four_regions,ISO == 'MTQ', 'americas'), 
         six_regions = replace(six_regions,ISO == 'MTQ', 'america'), 
         eight_regions = replace(eight_regions,ISO == 'MTQ', 'america_north')) %>%
#new caledonia
  mutate(four_regions = replace(four_regions,ISO == 'NCL', 'asia'), 
         six_regions = replace(six_regions,ISO == 'NCL', 'east_asia_pacific'), 
         eight_regions = replace(eight_regions,ISO == 'NCL', 'east_asia_pacific')) %>%
#puerto rico
 mutate(four_regions = replace(four_regions,ISO == 'PRI', 'americas'), 
         six_regions = replace(six_regions,ISO == 'PRI', 'america'), 
         eight_regions = replace(eight_regions,ISO == 'PRI', 'america_north')) %>%
#Palestine (West Bank + Gaza strip)
 mutate(four_regions = replace(four_regions,ISO == 'PSE', 'asia'), 
         six_regions = replace(six_regions,ISO == 'PSE', 'middle_east_north_africa'), 
         eight_regions = replace(eight_regions,ISO == 'PSE', 'asia_west')) %>%
#french Polynesia
  mutate(four_regions = replace(four_regions,ISO == 'PYF', 'asia'), 
         six_regions = replace(six_regions,ISO == 'PYF', 'east_asia_pacific'), 
         eight_regions = replace(eight_regions,ISO == 'PYF', 'east_asia_pacific')) %>%
#Reunion
  mutate(four_regions = replace(four_regions,ISO == 'REU', 'africa'), 
         six_regions = replace(six_regions,ISO == 'REU', 'sub_saharan_africa'), 
         eight_regions = replace(eight_regions,ISO == 'REU', 'africa_sub_saharan')) %>%
#Serbia, Montenegro, Kosovo
  mutate(four_regions = replace(four_regions,ISO == 'SMK', 'europe'), 
         six_regions = replace(six_regions,ISO == 'SMK', 'europe_central_asia'), 
         eight_regions = replace(eight_regions,ISO == 'SMK', 'europe_east')) %>%
#Slovakia
  mutate(four_regions = replace(four_regions,ISO == 'SVK', 'europe'), 
         six_regions = replace(six_regions,ISO == 'SVK', 'europe_central_asia'), 
         eight_regions = replace(eight_regions,ISO == 'SVK', 'europe_east')) %>%
#Turks and Caicos Islands
  mutate(four_regions = replace(four_regions,ISO == 'TCA', 'americas'), 
         six_regions = replace(six_regions,ISO == 'TCA', 'america'), 
         eight_regions = replace(eight_regions,ISO == 'TCA', 'america_north')) %>%
#East Timor
  mutate(four_regions = replace(four_regions,ISO == 'TLS', 'asia'), 
         six_regions = replace(six_regions,ISO == 'TLS', 'east_asia_pacific'), 
         eight_regions = replace(eight_regions,ISO == 'TLS', 'east_asia_pacific')) %>%
#taiwan
  mutate(four_regions = replace(four_regions,ISO == 'TWN', 'asia'), 
         six_regions = replace(six_regions,ISO == 'TWN', 'east_asia_pacific'), 
         eight_regions = replace(eight_regions,ISO == 'TWN', 'east_asia_pacific')) %>%
#British virgin Islands
  mutate(four_regions = replace(four_regions,ISO == 'VGB', 'americas'), 
         six_regions = replace(six_regions,ISO == 'VGB', 'america'), 
         eight_regions = replace(eight_regions,ISO == 'VGB', 'america_north')) %>%
#US Virgin Islands
  mutate(four_regions = replace(four_regions,ISO == 'VIR', 'americas'), 
         six_regions = replace(six_regions,ISO == 'VIR', 'america'), 
         eight_regions = replace(eight_regions, ISO == 'VIR', 'america_north')) 

```
Next time I'll write a function.

There are also a lot of countries that have NA values for their income or population. Take a peek here.

```{r message=FALSE, warning=FALSE}
lotanas <-  aei[rowSums(is.na(aei)) > 0,]
lotanas %>% 
  group_by(country) %>% 
  select(income, population) %>%  
  summarise_all(funs(sum(is.na(.))))


```


The countries that don't have income also don't have population, so as least I am consistent in my missing data. I have to move on so I will just leave this here for a minute. Perhaps later we should fill these variables.  



## Water Supply

Again, precipitation is translated output from LPJmL. Precipitation was chosen because it could be summed across countries, as other data could not (like discharge or ET). I have summed the precipitation by country and year. Also included are the mean and median of Humidity and PET.
```{r Precip Manipulation, message=FALSE, warning=FALSE}
hum_PET_prec <- hum_PET_prec[,-c(1)]

#merge
aei <- merge(aei, hum_PET_prec, by = c("ISO", "year"), all.x = TRUE)
rm(hum_PET_prec)
```


## Crop Fraction

Uff. Then comes crop fraction. As mentioned above individual crop fraction per grid cell is multiplied by respective grid cell area, summed on a country level giving ha of each crop grown per country. This was then divided by total area of the country to achieve the % area occupied by a certain crop per country.
```{r CFT manipulation, message=FALSE, warning=FALSE}
#pull cft frac outta here
cft$ISO <- as.factor(cft$ISO)
cft <- cft[, -c(1)] #remove id col
cft <- subset(cft, year >= 1910)

cft <-
  cft %>%
  select(-c(4)) %>% #removing precip
  mutate_each(funs(.*cellaream2), c(4:35)) %>% #multiply all fractions with the cell area
  group_by(ISO, year) %>% #group in to country and year
  summarise_each(funs(sum)) %>% #take col sums for each crop
  mutate(across(c(3:34), .fns = ~./cellaream2)) #divide col sums by total country area

#removing categories with 0 (the biofuels and such)
cft <- cft[, -c(18, 19, 26, 34, 35)]

#and a merge
aei <- merge(aei, cft, by = c("ISO", "year"), all.x = TRUE)
rm(cft)
```

Idk how to show the data here. Any ideas on representation would be helpful here. 

## Topographical

Topographical data is represented here by the data set from @rileyTerrainRuggednessIndex1999 and is conviently available through the `library(rethinking)` from @mcelreathStatisticalRethinkingBayesian2020. It provides a ruggedness factor for each country. 

```{r}
#rugged
data("rugged")
rug <- rugged %>% select(c("isocode", "rugged"))
colnames(rug) <- c("ISO", "rugged")

#merge
aei <- merge(aei, rug, by = "ISO", all.x = TRUE)
rm(rug, rugged)


```


## Political System

This data is from [ourworldindata](https://ourworldindata.org/democracy#world-maps-of-political-regimes-over-200-years) and is the democracy index for the last 200 years. There are some gaps due to country codes and such. I will use the raw data used to create this data from the [Center for systematic peace](https://www.systemicpeace.org/inscrdata.html) to fill in the gaps where I can. 



```{r}
#merge democracy to aei

colnames(democracy) <- c("country", "ISO", "year", "polity")
democracy <- democracy[,-1]

polity <-
  aei %>% 
  select(ISO, year, aei_ha, country) %>% 
  left_join(., democracy, by = c("ISO", "year")) #merge based on aei ISO

missing <-
  polity %>% 
  subset(., is.na(.$polity)) %>% #select those with no polity
  mutate( month = "01", day = "01", 
          year = as.Date(with(., paste(year, month, day,sep="-")), 
                          "%Y-%m-%d")) %>% #fix the dates so we can merge 
  select(ISO, year, polity, country)

```

```{r}
#deal with center for systemic peace data
syspeace <- syspeace[, -c(8:20)] #take some rows
syspeace$date <- as.Date(with(syspeace, paste(byear, bmonth, bday,sep="-")), "%Y-%m-%d") #make some dates
syspeace <- syspeace[, -c(4:7)] 
colnames(syspeace) <- c("ISO", "country","polity", "year")

#quick flip for russia
syspeace$ISO <- replace(as.character(syspeace$ISO), syspeace$ISO == "USR", "RUS")
  
#merge syspeace with missing

aei<-
  merge(missing, syspeace, by = c("ISO", "year", "polity", "country"), all = TRUE) %>% 
  arrange(., ISO, year) %>% 
  group_by(ISO) %>% 
  fill(polity, .direction = "down") %>% 
  ungroup() %>% 
  subset(year == "1910-01-01" | 
           year == "1920-01-01" |
           year == "1930-01-01" |
           year == "1940-01-01" |
           year == "1950-01-01" |
           year == "1960-01-01" |
           year == "1970-01-01" |
           year == "1980-01-01" |
           year == "1985-01-01" |
           year == "1990-01-01" |
           year == "1995-01-01" |
           year == "2000-01-01" | 
           year == "2005-01-01" ) %>%  #ok filled a couple of gaps.
  select(ISO, year, polity) %>% 
  mutate(year = as.character(year), 
         year = str_replace(year, '-01-01', ""),
         year = as.numeric(year)) %>% 
  merge(., democracy, by = c("ISO", "year", "polity"), all = TRUE) %>% 
  merge(., aei, by = c("ISO", "year"), all.y = TRUE) %>% 
  subset(., select=c(1:2, 4:55, 3))
```

```{r}
aei <- aei %>%
  add_column(polity_case = NA)
aei[which(aei$polity>= -10 & aei$polity<= -6), 'polity_case'] <- "Autocracy"
aei[which(aei$polity>= -5 & aei$polity<= 0), 'polity_case'] <- "Closed_Anocracy"
aei[which(aei$polity>= 1 & aei$polity<= 5), 'polity_case'] <- "Open_Anocracy"
aei[which(aei$polity>= 6 & aei$polity<= 10), 'polity_case'] <- "Democracy"
aei[which(aei$polity== -20 ), 'polity_case'] <- "Colony/Non-State"
aei[which(aei$polity== -66 ), 'polity_case'] <- "Interruption"
aei[which(aei$polity== -77 ), 'polity_case'] <- "Interregnum"
aei[which(aei$polity== -88 ), 'polity_case'] <- "Transition"
aei$polity_case <- as.factor(aei$polity_case)
```



Save that!
```{r}
#write.csv(aei, "/Volumes/RachelExternal/Thesis/Data_upload_for_CL/AEI_NoStd.csv")
```



***




# References
