---
title: "R Notebook"
output: html_notebook
---
```{r}
library(tidyverse)
library(dplyr)
library(tm)
```



```{r}
aei2 <- read.csv2("/Volumes/RachelExternal/Thesis/Data/SiebertData/HID_v10/Country_tenyear_ha.csv")
aei2 <- as.data.frame(aei2)

#rename the cols so that I can use them as number
colnames(aei2) <- c("Country", "ISO", "ID", "1900", "1910","1920","1930","1940","1950","1960","1970","1980","1985","1990","1995","2000","2005")

#tidying data to long format
aei2 <- pivot_longer(aei2,cols = 4:17, names_to = "year", values_to = "aei_ha")
aei2$aei_ha <- removePunctuation(aei2$aei_ha)
aei2$year <- as.numeric(aei2$year)
aei2$aei_ha <- as.numeric(aei2$aei_ha)
aei2$ISO <- as.factor(aei2$ISO)
#subset the data so that its from the year 1960
aei2 <- aei2 %>% subset(year >= 1960)

#make a new col for year count
aei2$yearcount <- (aei2$year - 1960)

str(aei2)
```
```{r}
#income here is 2011 usd 
income <- read.csv("/Volumes/RachelExternal/Thesis/Thesis/DataWesternosprey/Gapminder_income_per_person_gdppercapita_ppp_inflation_adjusted.csv")
income <- income[, c(1, 162:207)]
#still messing with this and i need to shave off some cols still
income <- 
  income %>%  
  pivot_longer(!country, names_to = "year", values_to = "income") 

#still have the xs there
income$year <- str_remove_all(income$year, "[X]")
income$year <- as.numeric(income$year)
income$country <- as.factor(income$country)

iso <- read.csv2("/Volumes/RachelExternal/Thesis/Thesis/DataWesternosprey/GAPminder_ISOcodes.csv")
iso <- iso[, -c(2,7,10,12,13)]
iso <- 
  iso %>% 
  rename(ISO = GEO, country = name)

iso$ISO <- as.factor(iso$ISO)
iso$country <- as.factor(iso$country)
iso$four_regions <- as.factor(iso$four_regions)
iso$eight_regions <- as.factor(iso$eight_regions)
iso$six_regions <- as.factor(iso$six_regions)
iso$six_regions <- as.factor(iso$six_regions)
iso$World.bank.region <- as.factor(iso$World.bank.region)

income <- merge(income, iso, by = "country", all.x = TRUE)

population <- read.csv("/Volumes/RachelExternal/Thesis/Thesis/DataWesternosprey/Gapminder_population_total.csv")
population <- population[, c(1, 162:207)]
population <- 
  population %>%  
  pivot_longer(!country, names_to = "year", values_to = "population") 
population$year <- str_remove_all(population$year, "[X]")
population$country <- as.factor(population$country)
population$year <- as.numeric(population$year)

soceco <- merge(income, population, by = c("country", "year"))

```

```{r}

aei2 <- read.csv2("/Volumes/RachelExternal/Thesis/Data/SiebertData/HID_v10/Country_tenyear_ha.csv")
aei2 <- as.data.frame(aei2)

#rename the cols so that I can use them as number
colnames(aei2) <- c("Country", "ISO", "ID", "1900", "1910","1920","1930","1940","1950","1960","1970","1980","1985","1990","1995","2000","2005")

#tidying data to long format
aei2 <- pivot_longer(aei2,cols = 4:17, names_to = "year", values_to = "aei_ha")
aei2$aei_ha <- removePunctuation(aei2$aei_ha)
aei2$year <- as.numeric(aei2$year)
aei2$aei_ha <- as.numeric(aei2$aei_ha)
aei2$ISO <- as.factor(aei2$ISO)
#subset the data so that its from the year 1960
aei2 <- aei2 %>% subset(year >= 1960)

#make a new col for year count
aei2$yearcount <- (aei2$year - 1960)

str(aei2)

```
```{r}
df <- left_join(aei2, soceco, by = c("ISO", "year"))

```

```{r}
#import the data set with total country area
countryarea <- read.csv2("/Volumes/RachelExternal/Thesis/Data/LandArea/LandArea.csv")
#str(countryarea)

#first lets remove the little things at the end
countryarea <- countryarea[-c(265:270), ]
#lets remove extra things, as all we want is values and ISO codes
countryarea <- countryarea[, -c(1, 3:4)]
#fix the col names
for ( col in 2:ncol(countryarea)){
    colnames(countryarea)[col] <-  sub("X", "", colnames(countryarea)[col])
}
remove(col)
#oooh lets get pivot longer out
countryarea <- 
  countryarea %>%  
  pivot_longer(!Country.Code, names_to = "year", values_to = "area_km") %>% 
  rename(ISO = Country.Code)

countryarea$ISO <- as.factor(countryarea$ISO)
countryarea$year <- as.numeric(countryarea$year)


df <- left_join(df, countryarea, by = c("ISO", "year"))
```


```{r}
df <-
  df%>% 
  group_by(ISO) %>% 
  fill(area_km, .direction = "up") %>% 
  ungroup()

#checking for countries with NA values in their area_km col
#looking primarily for countres that have non 0 aei_ha
errors <- df[is.na(df$area_km),]
                
#fixing counrties with no area
df$area_km[which(df$ISO == "GUF")] <- 83534 #French Guiana from Britannica
df$area_km[which(df$ISO == "GLP")] <- 1630 #Guadeloupe from Britannica 
df$area_km[which(df$ISO == "MTQ")] <- 1128 # Martinique from Britannica
df$area_km[which(df$ISO == "REU")] <- 2512 #Reunion from Britannica
df$area_km[which(df$ISO == "SMK")] <- 13812 + 10905 + 77589 #Montenegro + Kosovo + Serbia from Britannica
df$area_km[which(df$ISO == "TWN")] <- 35980 #Tawian from CIA.gov
df$area_km[which(df$ISO == "SDN")] <- 1882000 #Sudan from UNDP

#check errors again then rm
rm(errors)
```


```{r}
df <- df %>% mutate(irrperc = ((aei_ha/100)/area_km)*100, irrfrac = ((aei_ha/100)/area_km))
```

```{r}
lotanas2 <-  df[rowSums(is.na(df)) > 0,]
lotanas2 %>% 
  group_by(ISO) %>% 
  select(income, population) %>%  
  summarise_all(funs(sum(is.na(.))))


```
OKKKK this has way way less countries with missing pop and income. ill use the gapminder dataset.




***
***
***
***
***
This is the data from world bank that is not as complete as the gapminder stuff
```{r Adding pop and GDP}
#get the data from the worldbank. 
popgdp <- wb(indicator = c("NY.GDP.PCAP.CD", "SP.POP.TOTL"), 
                       country = "countries_only", startdate = 1960, enddate = 2005)  %>% 
  # pull down mapping of countries to regions and join
  left_join(wbcountries() %>% 
              select(iso3c, region)) %>% 
  # spread the three indicators
  pivot_wider(id_cols = c("date", "iso3c"), names_from = indicator, values_from = value)

colnames(popgdp) <- c("year", "ISO", "gdppc", "poptot")

#fix the col types
popgdp$ISO <- as.factor(popgdp$ISO)
popgdp$year <- as.numeric(popgdp$year)
```
Join the data frames. 
```{r Join}
aei <- left_join(x = aei, y = popgdp, by =  c("ISO", "year"))
rm(popgdp)
```
Log and standardize GDP and Population. 
```{r}
#lets log and standardize gdp and poulation. 
#log (going to x+1 so that i dont have any pesky -infs) and standardize
aei$log_gdppc <- log(aei$gdppc)
logstdgdp <- aei %>% 
  filter(!is.na(log_gdppc)) %>% 
  select(c(ISO, year, log_gdppc)) %>% 
  mutate(log_gdp_std = log_gdppc / mean(log_gdppc))
aei <- logstdgdp %>% 
  select(ISO, year, log_gdp_std) %>% 
  merge(x = aei, y = ., by = c("ISO", "year"), all.x = TRUE)
rm(logstdgdp)
```
So, I've been sneaky with the NAs. I have removed them and standardized those that had values, disregarding the ones with NAs, and then remreging the dataframes, so that I have the option to get a sense of There are still a lot, but I'll leave them in for now. Perhaps I'll use categories of GDP instead of values, which would solve the problem of missing values for GDP. The NAs with population are tougher, but again I could transform things in to categories (Big, medium, small countries, etc). Let's see how this works with continuous variables though. I can identify where the NAs are though. 

```{r Checking for NAs, include=FALSE}
#have a lot of NAs still... grahhh
lotanas <-  aei[rowSums(is.na(aei)) > 0,]
lotanas %>% 
  group_by(Country) %>% 
  select(gdppc, poptot) %>%  
  summarise_all(funs(sum(is.na(.))))
```
Almost half the countries have an NA value for either population or GDP.. this might be tough to work with.. but I will leave it.  
However, the worldbank did categorize countries by incomes. There is no time series component to this though. For example, China is currently and upper middle income country but they did not start that way. Let me see if I can create summary statistics to replace the income groups with useful classifications.   

I am assuming that these classifications are based on the most recent data from world bank, so let me see if i can break it down and put some limits on each category. 

```{r}
countrymetadata <- 
  read.csv2("/Volumes/RachelExternal/Thesis/Data/LandArea/MetaData.csv") %>% 
  subset(select = c(1,3)) %>% 
  rename(ISO = Country.Code)
  

incats <- wb(indicator = c("NY.GDP.PCAP.CD", "SP.POP.TOTL"), 
                       country = "countries_only", startdate = 2005, enddate = 2021)  %>% 
  # pull down mapping of countries to regions and join
  left_join(wbcountries() %>% 
              select(iso3c, region)) %>% 
  # spread the three indicators
  pivot_wider(id_cols = c("date", "iso3c"), names_from = indicator, values_from = value)%>% 
  rename(ISO = iso3c, year = date) %>%
  filter(year == "2017") %>% 
  merge(countrymetadata, by = "ISO")

incats$IncomeGroup <- as.factor(incats$IncomeGroup)
  
incats %>% split(.$IncomeGroup) %>% map(summary)
```


Adding the larger regions. 
```{r Adding regions}
regional <- read.csv("/Volumes/RachelExternal/Thesis/Data/ISO-3166-Countries-with-Regional-Codes/all/all.csv")
regional <- 
  regional %>% 
  subset(select = c("alpha.3", "region")) %>% 
  rename(ISO = "alpha.3")

aei <- left_join(aei, regional, by = "ISO")
aei$region <- as.factor(aei$region)

rm(regional)
  
```


```{r Standardizing GDP and POP, include=FALSE}
#GDP and pop per capita
gdppop <- read.csv2("/Volumes/RachelExternal/Thesis/Thesis/DataWesternosprey/GDP_pop.csv", header = TRUE)
#limits
gdppop <- subset(gdppop, year >= 1950 & year <= 2005 & country != "Africa", select = c("countrycode", "year", "gdppc", "pop"))
colnames(gdppop) <- c("ISO", "year", "gdppc", "pop")
gdppop$gdppc <- as.numeric(gdppop$gdppc)
gdppop$pop <- as.numeric(gdppop$pop)
countrystats <- merge(countrystats, gdppop, on = c("ISO", year), all.x = TRUE)
rm(gdppop)
summary(countrystats)


```
