---
title: "Country Level Target Variable Investigaion"
author: "Rachel Ledig"
date: "4/20/2021"
output: 
  html_document:
    toc: true
    toc_float: true
bibliography: /Volumes/RachelExternal/Thesis/Thesis/bib.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
```


```{r include=FALSE}
library(ggplot2)
library(dplyr)
library(plotly)
```



```{r}
aei <- read.csv("/Volumes/RachelExternal/Thesis/Data_upload_for_CL/AEI_NoStd.csv")
```



# Time Series Evolution of Country Level Target Variable

Ok, so lets peek at how things change over time. Here is a graph of percentage of country area equipped for irrigation from 1960 - 2005.

```{r Irrigated Totals Graph, warning=FALSE}

pirrfrac <-
  ggplot(data = subset(aei, !is.na(irrperc)), 
         mapping = aes(x=year, y= irrperc, group = country, color = six_regions)) +
  geom_line() +
  scale_x_continuous() +
  theme(panel.grid = element_blank()) 

#run this separately, otherwise the plotly part doesn't work.
ggplotly(pirrfrac)


```
Although this graph has a lot of countries and can get quite messy, some general trends appear. South Asian countries seem to have high percentages of irrigated area per total land area (irrigation percentage). Then seemingly followed by East Asian and European countries. Africa and the Middle East do not seem to be irrigated as much. Be aware that you can these graphs are interactive, you may zoom, pan or click on the legend to isolate the regions. Although, a regional breakdown follows.



## Sub-Saharan Africa
  
```{r Africa Irrigated Totals Graph, echo=FALSE, warning=FALSE}
pirrssAF <-
  ggplot(data = subset(aei, six_regions == "sub_saharan_africa"), 
         mapping = aes(x=year, y= irrperc, color = country)) +
  geom_line() +
  scale_x_continuous() +
  theme(panel.grid = element_blank()) 

#run this separately, otherwise the plotly part doesn't work.
ggplotly(pirrssAF)
```
  
In general, Africa has low levels of irrigated area. Swaziland, Sao Tome and Principe, Reunion and Mauritius lead here with majority of others falling below the levels of these countries.

## Europe and Centrtal Asia
  
```{r Europe Irrigated Totals Graph, echo=FALSE, warning=FALSE}
pirrEU <-
  ggplot(data = subset(aei, six_regions == "europe_central_asia"), 
         mapping = aes(x=year, y= irrperc, color = country)) +
  geom_line() +
  scale_x_continuous() +
  theme(panel.grid = element_blank()) 

#run this separately, otherwise the plotly part doesn't work.
ggplotly(pirrEU)
```
  
Europe contains some countries that reach higher percentages of irrigated area. Countries such as the Netherlands, Romania, Albania, and Azerbaijan have reached peaks in the past (roughly 10-15% irrigated area) but seem to be in a decline towards the end our study period. Other countries such as Italy and Denmark have had more stable rises with less down turn toward. Unexpectedly, Russia has a very low percentage of irrigated area, however at this point, the assumption is that either Russia's enormous size cancels out its irrigated area, or that Russia receives enough precipitation for the crops it grows that irrigation is not necessary. Further analysis should reveal this.  

## Americas

```{r Americas Irrigated Totals Graph, echo=FALSE, warning=FALSE}
pirrAM <-
  ggplot(data = subset(aei, six_regions == "america"), 
         mapping = aes(x=year, y= irrperc, color = country)) +
  geom_line() +
  scale_x_continuous() +
  theme(panel.grid = element_blank()) 

#run this separately, otherwise the plotly part doesn't work.
ggplotly(pirrAM)
```
  
In the Americas, the % of area irrigated by countries is lower than Asia or Europe. Peaks here are Cuba and the Dominican Republic at around roughly 6-8% of their total countries being irrigated. There are some islands (Barbados and St. Lucia) that have sharp increases in irrigation fraction, but this could be because they are islands and any irrigation at all would induce the ration of irrigated land to total land spike. Counties that have experienced a steady increase, but are not the leaders in overall irrigation fraction are countries that are big and produce a lot of produce (USA, Mexico, Chile) The general trends of irrigation increase seem to be upward with perhaps a slight leveling off toward the end of the study period (for some countries) but lacking the distinct downturn that is seen in Europe.   
  
## East Asia and Oceania

```{r East Asia and Oceania Irrigated Totals Graph, echo=FALSE, warning=FALSE}
pirrEAsia <- 
  ggplot(data = subset(aei, six_regions == "east_asia_pacific"), 
         mapping = aes(x=year, y= irrperc, color = country)) +
  geom_line() +
  scale_x_continuous() +
  theme(panel.grid = element_blank()) 

#run this separately, otherwise the plotly part doesn't work.
ggplotly(pirrEAsia)
```
  
The East Asian countries are quite high in % of irrigated area, when compared to the Americas or Africa. The four countries that reach upwards of (roughly) 10% or more of their total area irrigated toward the end of the study period are Vietnam, North Korea, South Korea, and Thailand. Other countries that have high irrfracs are China, Japan and the Philippines. It is expected that for this region the irrigation fraction would be high due to the amount of rice produced. 

Oceania, due to its large amount of islands, does not have a lot of irrigation, majority of countries have 0% irrigation. New Zealand (around 2%) and Australia (less than 1%) have some irrigation. Although this could have a multitude of reasons why irrigation is so low here, regardless of the fact that Australia and New Zealand are very populous countries with mouths to feed. Australia is very large and perhaps its enormity cancels out the irrigated land it does have. 
  
## South Asia
```{r South Asia Irrigated Totals Graph, echo=FALSE, warning=FALSE}
pirrSAsia <-
  ggplot(data = subset(aei, six_regions == "south_asia"), 
         mapping = aes(x=year, y= irrperc, color = country)) +
  geom_line() +
  scale_x_continuous() +
  theme(panel.grid = element_blank()) 

#run this separately, otherwise the plotly part doesn't work.
ggplotly(pirrSAsia)
```
Very few countries here, countries in this area are large. Very high levels of irrigation here coming from most countries. Bangladesh, Pakistan, and India are heavily irrigated compared to their land area. 


## Middle East and North Africa
```{r Middle East and North Africa Irrigated Totals Graph, echo=FALSE, warning=FALSE}
pirrME <-
  ggplot(data = subset(aei, six_regions == "middle_east_north_africa"), 
         mapping = aes(x=year, y= irrperc, color = country)) +
  geom_line() +
  scale_x_continuous() +
  theme(panel.grid = element_blank()) 

#run this separately, otherwise the plotly part doesn't work.
ggplotly(pirrME)
```

Irrigation is also prevalent here (but not as much as in south Asia). These countries are dry and therefore would need irrigation to have domestic food supplies. Lebanon, Iraq and Israel lead here. Oil producing countries have little irrigation, they are dry and their GDP is high, perhaps importation is the easiest option for them?

***

Overall, this might be a good overview of what irrperc looks like for each region. 
```{r message=FALSE, warning=TRUE}

theme_set(theme_minimal())

ggplot(subset(aei, !is.na(irrperc)),
       aes(x=six_regions, y=irrperc, fill=six_regions)) + 
  geom_boxplot(show.legend = FALSE)  + 
  coord_flip() +
  theme(legend.title = element_blank())
```
