---
title: "R Notebook"
output: html_notebook
---
Necessary: 
d <- data

### Housekeeping    
```{r Housekeeping, warning=FALSE, message=FALSE}
library(brms)
library(rstan)
library(dplyr)
library(tidyr)
library(plotly)
library(ggplot2)
library(rethinking)
library(purrr)
library(gganimate)
library(wbstats)
library(viridis)
library(tm)
library(wbstats)
library(tibble)
```

## Data  
  

So data has been collected, cleaned and organized. Lets do a short introduction to what we're working with here. 

#### Target Variable  
  
* $Y$ - Irrigation Percentage - the percentage of the country that is irrigated. This is calculated by grouping cells by country name and year to calculate the sum of the irrigated area, then divide by the total calculated area of the country. I wont scale the outcome variable as its already scaled from 0 to 1. 

```{r include=FALSE}
#total irrigated fraction per country per year
#no need to divide for km2 conversion, as final is unitless
#remove data before 1950
# countrystats <- d %>%  group_by(countryname, ISO, year) %>% summarise(IrrFrac = ((sum(Frac*cellaream2))/sum(cellaream2)))
# countrystats$IrrFrac <- round(countrystats$IrrFrac, digits = 5)
# countrystats <- as.data.frame(countrystats)
# countrystats$year <- as.character(countrystats$year)
# countrystats$year <- as.numeric(countrystats$year)
# countrystats <- subset(countrystats, year >= 1960, select = c("ISO", "year", "countryname", "IrrFrac"))
# 
# countrystats <- 
#   countrystats %>% 
#   mutate(yearcount = year - 1960)


```
This data is taken directly from the Siebert data set.  
```{r Setting up CL df}
aei <- read.csv2("/Volumes/RachelExternal/Thesis/Data/SiebertData/HID_v10/Country_tenyear_ha.csv")
aei <- as.data.frame(aei)

#rename the cols so that I can use them as number
colnames(aei) <- c("Country", "ISO", "ID", "1900", "1910","1920","1930","1940","1950","1960","1970","1980","1985","1990","1995","2000","2005")

#tidying data to long format
aei <- pivot_longer(aei,cols = 4:17, names_to = "year", values_to = "aei_ha")
aei$aei_ha <- removePunctuation(aei$aei_ha)
aei$year <- as.numeric(aei$year)
aei$aei_ha <- as.numeric(aei$aei_ha)
aei$ISO <- as.factor(aei$ISO)
#subset the data so that its from the year 1960
aei <- aei %>% subset(year >= 1960)

#make a new col for year count
aei$yearcount <- (aei$year - 1960)

str(aei)
```
Data taken from Worldbank.
```{r adding C Area, warning=FALSE}
#import the data set with total country area
countryarea <- read.csv2("/Volumes/RachelExternal/Thesis/Data/LandArea/LandArea.csv")
#str(countryarea)

#first lets remove the little things at the end
countryarea <- countryarea[-c(265:270), ]
#lets remove extra things, as all we want is values and ISO codes
countryarea <- countryarea[, -c(1, 3:4)]
#fix the col names
for ( col in 2:ncol(countryarea)){
    colnames(countryarea)[col] <-  sub("X", "", colnames(countryarea)[col])
}
remove(col)
#oooh lets get pivot longer out
countryarea <- 
  countryarea %>%  
  pivot_longer(!Country.Code, names_to = "year", values_to = "area_km")

countrymetadata <- read.csv2("/Volumes/RachelExternal/Thesis/Data/LandArea/MetaData.csv")
#removing cols 4 and 5
countrymetadata <- countrymetadata[, -c(4:5)]
#join both country info dataframes
countryarea <- 
  countryarea %>% 
  left_join(y = countrymetadata, by = "Country.Code" )

#change the col names and fix the variable types
colnames(countryarea) <- c("ISO", "year", "area_km", "subregion", "incomegroup")
countryarea$year <- as.numeric(countryarea$year)
countryarea$ISO <- as.factor(countryarea$ISO)

#subset data
countryarea <- countryarea %>% subset(year <= 2005 ) 

head(countryarea)
head(countrymetadata)

```

```{r Join CAr }
#combine these data sets and calculate the irrfrac per country
aei <- aei %>% left_join(y = countryarea, by = c("ISO", "year") )
rm(countryarea, countrymetadata)
str(aei)
head(aei)
```

For some reason 1960 (the data starts in 1961) there is an issue with the registration of the total country area. Fixing it by using the value from 1970. Some countries also have no area value. I'll go through and see if any of the countries with missing area values have non zero aei_ha. If so, I'll replace them, as they didn't come through with the WorldBank data set but are online. Copying and pasting from various sources should be fine.
```{r Fixes for CAr}
#fixing the issue of 1960
aei <-
  aei %>% 
  group_by(ISO) %>% 
  fill(area_km, .direction = "up") %>% 
  ungroup()

#checking for countries with NA values in their area_km col
#looking primarily for countres that have non 0 aei_ha
errors <- aei[is.na(aei$area_km),]
                
#fixing counrties with no area
aei$area_km[which(aei$ISO == "GUF")] <- 83534 #French Guiana from Britannica
aei$area_km[which(aei$ISO == "GLP")] <- 1630 #Guadeloupe from Britannica 
aei$area_km[which(aei$ISO == "MTQ")] <- 1128 # Martinique from Britannica
aei$area_km[which(aei$ISO == "REU")] <- 2512 #Reunion from Britannica
aei$area_km[which(aei$ISO == "SMK")] <- 13812 + 10905 + 77589 #Montenegro + Kosovo + Serbia from Britannica
aei$area_km[which(aei$ISO == "TWN")] <- 35980 #Tawian from CIA.gov
aei$area_km[which(aei$ISO == "SDN")] <- 1882000 #Sudan from UNDP

#check errors again then rm
rm(errors)
```


Now, transform the AEI_ha to km^2 and divide to get the % of total country area that is equipped for irrigation. There are still some countries that have 0 aei_ha and no area_km value. I think I will assign the irrfrac/perc 0 to those countries, but lets wait a little while to do that, most are islands.  
```{r Creating Target Variable}
aei <- aei %>% mutate(irrperc = ((aei_ha/100)/area_km)*100, irrfrac = ((aei_ha/100)/area_km))

```

Here is a quick histogram of the target variable. Lotta 0s. 
```{r Hist of TVar}
hist(aei$irrperc, breaks = 20)
```

##### Percentage of Country Area Equipped for Irrigation (1960 - 2005)

```{r Irrigated Totals Graph, warning=FALSE}
pirrfrac <- plot_ly(aei, x = ~year, y = ~irrperc, color = ~region, 
                    type = 'scatter', mode = 'lines', 
                    text = ~paste("Country :", Country,"<br> Fraction :", irrfrac,  "<br> Year :", year),
                    textposition = "auto", hoverinfo = "text")
pirrfrac <- pirrfrac %>% layout(autosize = F, height = 500, width = 1000) 
pirrfrac <- pirrfrac %>% layout(title = 'Total Irrigation Percentage per Country',
         xaxis = list(title = 'Year'),
         yaxis = list(title = '% Irrigated Land of Total Area'))
pirrfrac
```
Although this graph has a lot of countries and can get quite messy, some general trends appear. Asian countries seem to have high percentages of irrigated area per total land area (irrigation percentage). Then seemingly followed by European countries. Africa and Oceania do not seem to be irrigated as much. Be aware that you can these graphs are interactive, you may zoom, pan or click on the legend to isolate the regions. Although, a regional breakdown follows, just need to add some predictor variables first. Hang tight.

***

### Chosen Predictor Variables  

According to the FAO, the choice of which irrigation, and by proxy whether to irrigate or not, comes from six factors. The predictor that will be used for this analysis are listed below:  

- Economic 

  - $M_c$ - GDP per Capita - Each country's GDP per Capita from Worldbank. This variable will be logged because as stated in *Statistical Rethinking* McElreath states (pg. 239), "We use the logarithm of it, because the logarithm of GDP is the *magnitude* of GDP. Since wealth generates wealth, it tends to be exponentially related to anything that increases it"
  
- Topographical   

  - $Ru_c$ - Represented here using the "rugged" data set from rethinking. This data was taken from Riley et al. 1999. 
  
- Soil Type  

  - This will not be included in this analysis. Although, it could be useful there is no good way to quantify soil type on a country level basis. This will be included in the grid cell model, as regions can be easily integrated with the grid cell structure. 

- Water Supply    

  - $R_c$ - Total Precipitation -  Translated output from LPJmL. Summed across two dimensions: month and country. In mm/country/year
- Crop Type   

- Social Influences  

  - $P_c$ - Total Population - Each country's total population on a yearly basis. Taken from https://ourworldindata.org/grapher/population. Again, this will be logged, due to the same argument for GDP, and centered.
  
All variables will be scaled or centered. 

```{r Adding pop and GDP}
#get the data from the worldbank. 
popgdp <- wb(indicator = c("NY.GDP.PCAP.CD", "SP.POP.TOTL"), 
                       country = "countries_only", startdate = 1960, enddate = 2005)  %>% 
  # pull down mapping of countries to regions and join
  left_join(wbcountries() %>% 
              select(iso3c, region)) %>% 
  # spread the three indicators
  pivot_wider(id_cols = c("date", "iso3c"), names_from = indicator, values_from = value)

colnames(popgdp) <- c("year", "ISO", "gdppc", "poptot")

#fix the col types
popgdp$ISO <- as.factor(popgdp$ISO)
popgdp$year <- as.numeric(popgdp$year)
```
Join the data frames. 
```{r Join}
aei <- left_join(x = aei, y = popgdp, by =  c("ISO", "year"))
rm(popgdp)
```
Adding the larger regions. 
```{r Adding regions}
regional <- read.csv("/Volumes/RachelExternal/Thesis/Data/ISO-3166-Countries-with-Regional-Codes/all/all.csv")
regional <- 
  regional %>% 
  subset(select = c("alpha.3", "region")) %>% 
  rename(ISO = "alpha.3")

aei <- left_join(aei, regional, by = "ISO")
aei$region <- as.factor(aei$region)

rm(regional)
  
```


```{r Standardizing GDP and POP, include=FALSE}
#GDP and pop per capita
gdppop <- read.csv2("/Volumes/RachelExternal/Thesis/Thesis/DataWesternosprey/GDP_pop.csv", header = TRUE)
#limits
gdppop <- subset(gdppop, year >= 1950 & year <= 2005 & country != "Africa", select = c("countrycode", "year", "gdppc", "pop"))
colnames(gdppop) <- c("ISO", "year", "gdppc", "pop")
gdppop$gdppc <- as.numeric(gdppop$gdppc)
gdppop$pop <- as.numeric(gdppop$pop)
countrystats <- merge(countrystats, gdppop, on = c("ISO", year), all.x = TRUE)
rm(gdppop)
summary(countrystats)

#gotta transform POP in to total
countrystats <- countrystats %>%  mutate(poptot = pop *1000)
countrystats$log_pop <- log(countrystats$poptot)
logstdpop <- countrystats %>% 
  filter(!is.na(log_pop)) %>% 
  select(c(ISO, year, log_pop)) %>% 
  mutate(log_pop_std = log_pop / mean(log_pop))
logstdpop <- logstdpop %>% 
  select(ISO, year, log_pop_std) 
countrystats <- merge(countrystats, logstdpop, by = c("ISO", "year"), all.x = TRUE)
rm(logstdpop)

#lets log and standardize gdp and poulation. 
#log (going to x+1 so that i dont have any pesky -infs) and standardize
countrystats$log_gdppc <- log(countrystats$gdppc +1)
logstdgdp <- countrystats %>% 
  filter(!is.na(log_gdppc)) %>% 
  select(c(ISO, year, log_gdppc)) %>% 
  mutate(log_gdp_std = log_gdppc / mean(log_gdppc))
logstdgdp <- logstdgdp %>% 
  select(ISO, year, log_gdp_std) 
countrystats <- merge(countrystats, logstdgdp, by = c("ISO", "year"), all.x = TRUE)
rm(logstdgdp)
```
So, I've been sneaky with the NAs. I have removed them and standardized those that had values, disregarding the ones with NAs, and then remreging the dataframes, so that I have the option to get a sense of There are still a lot, but I'll leave them in for now. Perhaps I'll use categories of GDP instead of values, which would solve the problem of missing values for GDP. The NAs with population are tougher, but again I could transform things in to categories (Big, medium, small countries, etc). Let's see how this works with continuous variables though. I can identify where the NAs are though. 

```{r Checking for NAs, include=FALSE}
#have a lot of NAs still... grahhh
lotanas <-  countrystats[rowSums(is.na(countrystats)) > 0,]
lotanas %>% 
  group_by(countryname) %>% 
  select(gdppc, pop) %>%  
  summarise_all(funs(sum(is.na(.))))
```
Majority of the NAs are on islands or eastern European countries after the fall of the USSR.  
Pressing on!  

```{r include=FALSE}

#precipitation
precip <- d %>% group_by(countryname, year) %>% summarise(precip = (sum(precipmm)))
countrystats <- merge(countrystats, precip, on = c("countryname", "year"))
rm(precip)
#scale
countrystats$precip_std <- countrystats$precip / max(countrystats$precip)


#rugged
data("rugged")
rug <- rugged %>% select(c("isocode", "rugged")) 
colnames(rug) <- c("ISO", "rugged")
countrystats <- merge(countrystats, rug, on = "ISO")
rm(rug, rugged)
#scale
countrystats$rugged_std <- countrystats$rugged / max(countrystats$rugged)
```


***

### Regional Breakdown of % of Area Irrigated  
  
```{r Africa Irrigated Totals Graph, echo=FALSE, warning=FALSE}
pirafrica <- plot_ly(subset(aei, region == "Africa" ), x = ~year, y = ~irrfrac, color = ~Country, 
                    type = 'scatter', mode = 'lines', 
                    text = ~paste("Country :", Country,"<br> Fraction :", irrfrac,  "<br> Year :", year),
                    textposition = "auto", hoverinfo = "text")
pirafrica <- pirafrica %>% layout(autosize = F, height = 500, width = 1000) 
pirafrica <- pirafrica %>% layout(title = 'Total Irrigation Fraction per African Countries',
         xaxis = list(title = 'Year'),
         yaxis = list(title = '% Irrigated Land of Total Area'))
pirafrica
```
  
In general, Africa has low levels of irrigated area. Swaziland, Morocco and Egypt lead with around 2.5-3% of total area being irrigated. However, as with all of the data for irrigated area, many countries have volatile curves that peak and fall repeatedly over the course of the last century. Here Morocco and Tunisia stand out.  
  
```{r Europe Irrigated Totals Graph, echo=FALSE, warning=FALSE}
pireuro <- plot_ly(subset(aei, region == "Europe" ), x = ~year, y = ~irrfrac, color = ~Country, 
                    type = 'scatter', mode = 'lines', 
                    text = ~paste("Country :", Country,"<br> Fraction :", irrfrac,  "<br> Year :", year),
                    textposition = "auto", hoverinfo = "text")
pireuro <- pireuro %>% layout(autosize = F, height = 500, width = 1000) 
pireuro <- pireuro %>% layout(title = 'Total Irrigation Fraction per European Countries',
         xaxis = list(title = 'Year'),
         yaxis = list(title = '% Irrigated Land of Total Area'))
pireuro
```
  
Europe contains some countries that reach higher percentages of irrigated area. Countries such as the Netherlands, Romania and Bulgaria have reached peaks in the past (roughly 10-15% irrigated area) but were in a decline towards the end our my study period. Other countries such as Italy and Portugal have had more stable rises with less down turn toward. Unexpectedly, Russia has a very low percentage of irrigated area, however at this point, the assumption is that either Russia's enormous size cancels out its irrigated area, or that Russia receives enough precipitation for the crops it grows that irrigation is not necessary. Further analysis should reveal this.  

```{r Americas Irrigated Totals Graph, echo=FALSE, warning=FALSE}
piramerica <- plot_ly(subset(aei, region == "Americas" ), x = ~year, y = ~irrfrac, color = ~Country, 
                    type = 'scatter', mode = 'lines', 
                    text = ~paste("Country :", Country,"<br> Fraction :", irrfrac,  "<br> Year :", year),
                    textposition = "auto", hoverinfo = "text")
piramerica <- piramerica %>% layout(autosize = F, height = 500, width = 1000) 
piramerica <- piramerica %>% layout(title = 'Total Irrigation Fraction per American Countries',
         xaxis = list(title = 'Year'),
         yaxis = list(title = '% Irrigated Land of Total Area'))
piramerica
```
  
In the Americas, the % of area irrigated by countries is lower than Asia or Europe. Peaks here are Cuba and the Dominican Republic at around roughly 3-4% of their total countries being irrigated. These peak countries are cluttered around the equator, and although this is interesting it is not indicative of anything at the moment. The general trends of irrigation increase seem to be upward with perhaps a slight leveling off toward the end of the study period (for some countries) but lacking the distinct downturn that is seen in Europe.   
  
```{r Asia Irrigated Totals Graph, echo=FALSE, warning=FALSE}
pirasia <- plot_ly(subset(aei, region == "Asia" ), x = ~year, y = ~irrfrac, color = ~Country, 
                    type = 'scatter', mode = 'lines', 
                    text = ~paste("Country :", Country,"<br> Fraction :", irrfrac,  "<br> Year :", year),
                    textposition = "auto", hoverinfo = "text")
pirasia <- pirasia %>% layout(autosize = F, height = 500, width = 1000) 
pirasia <- pirasia %>% layout(title = 'Total Irrigation Fraction per Asian Countries',
         xaxis = list(title = 'Year'),
         yaxis = list(title = '% Irrigated Land of Total Area'))
pirasia
```
  
The Asian countries are really high in % of irrigated area. The four countries that reach upwards of (roughly) 15% or more of their total area irrigated toward the end of the study period are Bangladesh, India, Pakistan, and Azerbaijan, with Bangladesh reaching almost 30%. Some countries have interesting and volatile trajectories however. For example, seems to have 10 year cycles that peak and fall repeatdly toward the end of the study period.  
  
```{r Oceania Irrigated Totals Graph, echo=FALSE, warning=FALSE}
pirocean <- plot_ly(subset(aei, region == "Oceania" ), x = ~year, y = ~irrfrac, color = ~Country, 
                    type = 'scatter', mode = 'lines', 
                    text = ~paste("Country :", Country,"<br> Fraction :", irrfrac,  "<br> Year :", year),
                    textposition = "auto", hoverinfo = "text")
pirocean <- pirocean %>% layout(autosize = F, height = 500, width = 1000) 
pirocean <- pirocean %>% layout(title = 'Total Irrigation Fraction per Oceanian Countries',
         xaxis = list(title = 'Year'),
         yaxis = list(title = '% Irrigated Land of Total Area'))
pirocean
```
  
Oceania, due to its large amount of islands, does not have a lot of irrigation, majority of countries have 0% irrigation. New Zealand and Australia have some irrigation, with percentage of irrigated area at their peaks to be somewhere around 0.4% and 1.6%, respectively. Although this could have a multitude of reasons why irrigation is so low here, regardless of the fact that Australia and New Zealand are very populous countries with mouths to feed. Australia is very large and perhaps its enormity cancels out the irrigated land it does have.   

*** 
Lets do some stuff from Solomon Kurz's other bookdown, [_Applied Longitudinal Data Analysis in brms and the tidyverse_](https://bookdown.org/content/4253/exploring-longitudinal-data-on-change.html), to see if we can visualize our target variable a bit better. I'm also using some techniques from Hadley Wickham's nice youtube video, [_Managing many models with R_](https://www.youtube.com/watch?v=rz3_FDVt9eg&t=802s). We can 
```{r Nesting for First Mods}
by_ISO <-
  aei %>%
  filter(!is.na(irrfrac)) %>% 
  group_by(ISO) %>%
  nest()
```
Doing a little prior plotting, I've messed around a bit and settled on some semi-sensible priors assuming a gaussian distribution for both the parameter and slope. I am assuming that our intercept is normally distributed with around a mean of 2 and a standard variation of 2, our beta coef is centered around 0.01 with a sd of 0.1. This produces irrperc values within an acceptable range. There are some negative values here. Perhaps a prior that is bounded by 0 would be a better fit for this, but experiments with log normal distributions have proved difficult. Also there are few countries that have negative trajectoriies of irrigaton expansion, some do have a decrease towards the end of the study period.
```{r Prior Predict}
set.seed(17)
N <- 50
a <- rnorm(N , 2, 2)
b <- rnorm( N , 0., 0.1 )

plot( NULL , xlim=range(aei$yearcount) , ylim=c(-50,50) , xlab="year" , ylab="Irrigation Percentage" )
abline( h=0 , lty=2 )
abline( h=272 , lty=1 , lwd=0.5 )
for ( i in 1:N ) curve( a[i] + b[i]*x ,
from=min(aei$yearcount) , to=max(aei$yearcount) , add=TRUE , col=col.alpha("black",0.2) )
```
These don't look too bad. There are some lines that predict negative values but in general they seem to 

```{r First Fits}
fit1.1 <-
  brm(data = by_ISO$data[[1]],
      formula = irrperc ~ yearcount,
      control = list(adapt_delta = 0.99),
      prior = c(prior(normal(2,2), class = Intercept),
                prior(normal(0.01, 0.1), class = b)),
      iter = 4000, chains = 4, cores = 4,
      seed = 2)
      # file = "/Volumes/RachelExternal/Thesis/Thesis/fits/fit1.1")

print(fit1.1, digits = 4)

```
Alright, so the summary here looks fine. Everything converged and we have responses from year. However, the slope is just not much above zero. Perhaps a check of the priors and pp_check will give us more insight. 
```{r Prior Sum}
prior_summary(fit1.1)
```
brms uses flat priors for the slope (class = b) and a Student-T distobution for the intercept and sigma. These are weak and shouldnt have had much effect on the results. 
```{r Update Fit}
fit1.2 <-
  update(fit1.1, 
         newdata = by_ISO$data[[2]],
         file = "/Volumes/RachelExternal/Thesis/Thesis/fits/fit1.2")
print(fit1.2, digits = 4)
```
Alright. Plan is to apply this to all countries. Let's give it a shot.
```{r Map fit for all countries}
models <- 
  by_ISO %>%
  mutate(model = map(data, ~update(fit1.1, newdata = ., seed = 2)))
```


```{r Mean Structure}

mean_structure <-
  models %>% 
  mutate(coefs = map(model, ~ posterior_summary(.)[1:2, 1:2] %>% 
                       data.frame() %>% 
                       rownames_to_column("coefficients"))) %>% 
  unnest(coefs) %>% 
  select(-data, -model) %>% 
  unite(temp, Estimate, Est.Error) %>% 
  pivot_wider(names_from = coefficients,
              values_from = temp) %>% 
  separate(b_Intercept, into = c("init_stat_est", "init_stat_sd"), sep = "_") %>% 
  separate(b_yearcount, into = c("rate_change_est", "rate_change_sd"), sep = "_") %>% 
  mutate_if(is.character, ~ as.double(.) %>% round(digits = 2)) %>% 
  ungroup()

summary(mean_structure)
```

```{r Residual Variance}
residual_variance <-
  models %>% 
  mutate(residual_variance = map_dbl(model, ~ posterior_summary(.)[3, 1])^2) %>% 
  mutate_if(is.double, round, digits = 2) %>% 
  select(ISO, residual_variance)

summary(residual_variance)
```

```{r Bayesian Rsq}


r2 <-
  models %>% 
  mutate(r2 = map_dbl(model, ~ bayes_R2(., robust = T)[1])) %>% 
  mutate_if(is.double, round, digits = 2) %>% 
  select(ISO, r2)

summary(r2)
```

```{r Plotting CL irrfrac vs time}
abline <- 
  by_ISO %>% 
  unnest(data) %>% 
  
  ggplot(aes(x = yearcount, y = irrperc, group = ISO)) +
  geom_point() +
  geom_abline(data = mean_structure,
              aes(intercept = init_stat_est,
                  slope = rate_change_est, group = ISO),
              color = "blue") +
  scale_x_continuous() +
  coord_cartesian(ylim = c(0, 35)) +
  theme(panel.grid = element_blank()) +
  facet_wrap(~ISO, ncol = 2)

ggsave("/Volumes/RachelExternal/Thesis/Thesis/abline.png", abline, height = 200,limitsize = FALSE, dpi = 300 )
```
Alright. I think that this did an ok job of plotting regression for each countries irrperc trajectory over time. For some of the countries (ex. Bangladesh) the prior was way to restrictive to capture the high slope and intercept values. 
```{r}
lmlines<-
  by_ISO %>%
  unnest(data) %>% 
  ggplot(aes(x = yearcount, y = irrperc, group = ISO)) +
  stat_smooth(method = "lm", se = T) +
  geom_point() +
  scale_x_continuous() +
  theme(panel.grid = element_blank()) + 
  facet_wrap(~ISO, ncol = 2, scales = "free_y") 

ggsave("/Volumes/RachelExternal/Thesis/Thesis/lmlines.png", lmlines, height = 200,limitsize = FALSE, dpi = 300 )
```





  
### Mean and Median % of Irrigated Area by Region  
  
It is clear from the graphs below that Asian countries have the highest rates of irrigation per country, followed by Europe, the Americas, Africa and finally Oceania. The same trends are present when means and medians are taken. Interestingly, the downturn in Europe is very apparent in these graphs, showing that the downturn of irrigated area in Europe began around the early 1990s.  
```{r Mean and median, message=FALSE, warning=FALSE}
##lets take some summary statistics
#removing No_Land
summary(subset(countrystats, countryname== "No_Land") )
countrystats <-countrystats[!(countrystats$countryname=="No_Land"),]

meansmeds <- countrystats %>% group_by(Regions, year) %>% summarize(median = median(IrrFrac), mean = mean(IrrFrac)) 

ggplot(meansmeds) + geom_line(aes(y = median, x = year, group = Regions, color = Regions)) + theme(legend.title = element_blank()) +
  scale_x_discrete(breaks = seq(1901, 2005, 10)) +
  xlab("Years") +
  ylab("% of Area Irrigated") +  
  ggtitle(" Median % of Area Irrigated by Region ") 

ggplot(meansmeds) + geom_line(aes(y = mean, x = year, group = Regions, color = Regions)) + theme(legend.title = element_blank()) +
  scale_x_discrete(breaks = seq(1901, 2005, 10)) +
  xlab("Years") +
  ylab("% of Area Irrigated") +  
  ggtitle("Mean % of Area Irrigated by Region ") 
```
***

We can check out the data like this.
```{r}
countrystats %>% 
  select(log_gdp_std, IrrFrac, log_pop_std, precip_std, rugged_std) %>% 
  pairs()
```
Uff this is a mess. but somethings have patterns... but they are tough to see.     
Looking at the irrigation fraction column we can see a couple of things.    
The highest levels of irrigation are in countries with lowest ruggedness. Also unsurprisingly, countries with lowest precipitation have high irrigation fractions. Log_std_gdp doesn't seem to have much of an effect, everything looks a little blobby over there, or at least I cannot seem to sort much out. The same thing happens for GDP, but maybe countries with a slightly above average population have high levels of irrigation.   

But, now model building!

#### Beginning to Model

Ok lets specify something and see if we can get it to run. 
```{r}

cs <- countrystats[ complete.cases(c(countrystats$log_gdp_std, countrystats$log_pop_std)) , ]

m1 <- brm(formula = bf(IrrFrac ~ year,
                         zi ~ 1),
            data = cs,
            family = zero_inflated_beta)

```

```{r}
save(m1, file = "m1.rda")
summary(m1)
```
What does this all mean.. it takes a little while to run the model here too. Not crazy long tho. 



