---
title: "Country Level Zero Inflation Fits"
output: 
  html_document:
    toc: true
    toc_float: true
bibliography: /Volumes/RachelExternal/Thesis/Thesis/bib.bib
    
---

```{r Loaded Packages, include=FALSE}
library(knitr)
library(brms)
library(tidyverse)
library(dplyr)
library(tidyr)
library(ggplot2)
```


Necessary files:
```{r}
#read the file, if you dont have it from the previous markdown. 
aei <- read.csv("/Volumes/RachelExternal/Thesis/Data_upload_for_CL/AEI.csv")
```

# Eight Region Model

Jumping in to the zero inflated models. Some helpful stuff here has been [_How to analyze visual analog (slider) scale data?_](https://mvuorre.github.io/posts/2019-02-18-analyze-analog-scale-ratings-with-zero-one-inflated-beta-models/) by Matti @vuorreSometimesHowAnalyze2019. 

## The Model

Following what he has done, specify a `bf()`. Here we are using regions as the grouping factor, just to get things going. I've specified no priors here. Getting things running is the goal here. There is no time series component here, ignoring this for now.  
```{r 8 region ZIB 1}
#specifying the brms formula
zi_eight_regions_bf <- 
  bf(
  irrfrac ~ eight_regions,
  phi ~ eight_regions,
  zi ~ eight_regions,
  family = zero_inflated_beta()
)
```


```{r 8 region ZIB 2, echo=TRUE, results='hide'}
#and the model
zi_eight_regions <- 
  brm(
  formula = zi_eight_regions_bf,
  data = aei,
  cores = 4,
  seed = 17,
  file = "/Volumes/RachelExternal/Thesis/Thesis/fits/CL_Zero_Inflated_Fits/zero_inflated_eight_regions"
)
```


```{r 8 region ZIB 3}
print(zi_eight_regions)
```
This took about an hour to run. But rhats converged. 

## Priors

```{r}
prior_summary(zi_eight_regions)
```


Does `get_prior()` give me something different here?
```{r}
get_prior(zi_eight_regions_bf, aei, family = zero_inflated_beta())
```
Nah, exactly the same thing.  

### Visualization of b coeff Prior - Uniform

Alright, weve got flat priors for all of the b coeffs, a student t for the intercept (and phi's intercept), and a logistic prior for zi. We can do better than flat priors, the student t probably doesn't do too poor of a job. lets see if we can visualize some of these. 

```{r}
x <- seq(0,1,0.01)
curve(dunif(x, min = 0, max = 1))
```






Hahaha, Art. Yeah, so this is what is influencing the b_ coeffs, which is pretty much everything in the fit. We've got a decent amount of data with this model formula so this shouldnt have any influence on the estimates. 

### Visualization of Overall Intercept and Phi Prior - Student T 

For the overall intercept and the phi? The students t using the specified degrees of freedom in the prior. 
```{r}
curve(dt(x, 3)*2.5) # a scaled student t 
```

Alright this one doesn't look too too bad, considering we are very zero inflated, it does capture the general structure of our data which would be fine since this is the default prior for the intercept. But as the probability increases, this prior is probably a bit too generous (weak). I am still unsure how this prior would behave for the phi paramater. 

### Visualization of Overall Zi Prior - Logarithmic

Finally the logarithmic prior for the zi paramater.
```{r}
curve(dlogis(x, location = 0, scale = 1))
```

Ok, a similar thing is captured here as for the student t. This prior is weakly informing the model that the probability of 0s in our data set is closer to 0 than it is to 1, which seems a fair assumtion here.  

We can probably do better than this, both interms of the priors and the brms formula specified. But lets check out the posterior, that has been transfromed.


## Posterior

### Untransformed Densities for Post Estimates

Ok, so here are some of the posterior estimates for the predictors, they are still in the logit and log scale so the distrobutions are not representative of the propbability scale. 
```{r}
mcmc_plot(zi_eight_regions, pars = c("b_phi_Intercept", 
                        "b_phi_eight_regionsafrica_sub_saharan",
                        "b_phi_eight_regionsamerica_north",
                        "b_phi_eight_regionsasia_west",
                        "b_phi_eight_regionseast_asia_pacific",
                        "b_phi_eight_regionseurope_east", 
                        "b_phi_eight_regionseurope_west"), 
          type = "dens")
```

```{r}
mcmc_plot(zi_eight_regions, pars = "b_zi_", type = "dens")
```
```{r}
mcmc_plot(zi_eight_regions, pars = "b_phi_", type = "dens")
```
Alright. Need to work out a bit of what is being communicated here. Majority of these plots seem similarly wide. Some of the Zi paramaters seem bounded by zero on either the top or the bottom end. Perhaps I am having a hard time with these because I have not transformed them in to the response scale. Let's do that now. 

### Transformed Posterior

All of the summary output is transformed via the logit ($\mu, zi$) or log ($\phi$) link. We need to transform them in to the response scale, use `exp` to transform $\phi$ and `plogis` for $\mu$ and $zi$ to be able to interpret them better. I am assuming that all parameters that are $\phi$ related have the same  log link and need to be transformed via `exp` and others need`plogis` as they use the logit link. Is this a fair assumption?

```{r}
posterior_samples(zi_eight_regions, pars = "b_")[,1:24] %>% 
  #the old fashioned way, brute force.
  mutate_at(c("b_phi_Intercept", "b_phi_eight_regionsafrica_sub_saharan", 
              "b_phi_eight_regionsamerica_north", 
              "b_phi_eight_regionsamerica_north",
              "b_phi_eight_regionsasia_west", 
              "b_phi_eight_regionseast_asia_pacific",
              "b_phi_eight_regionseurope_east", 
              "b_phi_eight_regionseurope_west"), exp) %>% 
  mutate_at(vars(-c("b_phi_Intercept", "b_phi_eight_regionsafrica_sub_saharan", 
              "b_phi_eight_regionsamerica_north", 
              "b_phi_eight_regionsamerica_north",
              "b_phi_eight_regionsasia_west", 
              "b_phi_eight_regionseast_asia_pacific",
              "b_phi_eight_regionseurope_east", 
              "b_phi_eight_regionseurope_west")), plogis) %>% 
  posterior_summary() %>% 
  as.data.frame() %>% 
  rownames_to_column("Parameter") %>% 
  kable(digits = 2) 
```
Alright, so this has been transformed back in to the response scale. Some things are still a bit confusing to me here. One region (north africa) is encapsulated by the population level Intercept, zi_intercept, and phi_intercept. The phi_intercept is huge here, 44.95, but is similar to others as the difference between the intercept and the groups is captured by the specific "b_phi" value. 

Here "b_zi" is a measure of the proportions of zeros present in the data set, as it is just a bernoulli distrobution? Is this correct? and How could it be that for europe west and east asia pacific that the probability of a country having an irrigation value of 0 is 1? Is this a correct interpretation?

### Conditional Effects

We can also check out the conditional effects for this, they are also automatically transformed to the response scale, yay for brms!

```{r}
conditional_effects(zi_eight_regions, dpar = "mu")
```
Ok so this is the spread of mu, the average for the response, of irrigated land per country. Remember this is represented from 0 to 1. These don't seem to be too off in a conceptual sense. Sub-saharan Africa and south America have lower rates of irrigation than northern africa (its dry?) and north america (its big). Highest rates of irrigation are found in west asia, which also makes a sense, as these areas are usually fertile and dry, hence the need for irrigation. Then europe, east and west, along with east asia have expectedly high rates of irrigation. 

Some things to remember here though: each of these groups have a different number of countries that conclusions are being based off. (Ignore the NAs, they have zero irrigation fraction as per the Siebert Data).
```{r}
aei %>% 
  group_by(eight_regions) %>% 
  count()
```
So yeah, here Sub Saharan Africa has 352 samples as where South America has 96. Countries' sizes vary and regions are not equally divided by a certain number of countries. 

### Posterior Predictive Check

quick posterior check!
```{r}
pp_check(zi_eight_regions)
```
ahh this makes me so happy, as it looks so much better than the gaussian models I had been messing around with earlier.

***




# Simple Model with Sensible Predictors

Ok, let's try something here that is a bit more sensible in terms of predictors. Lets keep things linear with no hierarchial structure and try to set up some priors if we can. 

## The Model

```{r Formula Specs}
zi_simple_preds_1_bf <- 
  bf(
  irrfrac ~ yearcount,
  phi ~ yearcount,
  zi ~ yearcount + area_km,
  family = zero_inflated_beta()
)
```
Ok so here we assume that the irrigation fraction depends on the year, the dispersion coeff ($\phi$) is dependent on the year as well, and that the probability of a 0 is a linear function of the year and the size of the country. All are continuous.  

```{r Model 1, echo=TRUE, results='hide'}
#and the model
zi_simple_preds_1 <- 
  brm(
  formula = zi_simple_preds_1_bf,
  data = aei,
  cores = 4,
  seed = 17,
  file = "/Volumes/RachelExternal/Thesis/Thesis/fits/CL_Zero_Inflated_Fits/zi_simple_preds_1"
)
```

```{r}
print(zi_simple_preds_1)
```
Hmm.. This did not go the way that I expected at all, and I cant really use the output here as the Rhats have not converged to 1 and my effective sample sizes are laughable. 

## The Even Simpler Simple Model

Ok, so divergent transitions are through the roof. Lets knock out the $\phi$ parameter, cause I wanna model something I understand at this point. haha. 

```{r}
zi_simple_preds_2_bf <- 
  bf(
  irrfrac ~ yearcount,
  zi ~ yearcount,
  family = zero_inflated_beta() 
)
```

```{r Model 2, echo=TRUE, results='hide'}
#and the model
zi_simple_preds_2 <- 
  brm(
  formula = zi_simple_preds_2_bf,
  data = aei,
  cores = 4,
  seed = 17,
  file = "/Volumes/RachelExternal/Thesis/Thesis/fits/CL_Zero_Inflated_Fits/zi_simple_preds_2"
)
```
No yelling about divergent transitions here, I would be mad if it had.

```{r}
print(zi_simple_preds_2)
```

```{r}
zi_simple_preds_3_bf <- 
  bf(
  irrfrac ~ yearcount,
  zi ~ yearcount,
  family = zero_inflated_beta() 
)
```

```{r Model 3, echo=TRUE, results='hide'}
#and the model
zi_simple_preds_3 <- 
  brm(
  formula = zi_simple_preds_3_bf,
  data = aei,
  cores = 4,
  seed = 17,
  file = "/Volumes/RachelExternal/Thesis/Thesis/fits/CL_Zero_Inflated_Fits/zi_simple_preds_3"
)
```

```{r}
print(zi_simple_preds_3)
```
```{r}
zi_simple_preds_4_bf <- 
  bf(
  irrfrac ~ yearcount,
  zi ~ area_km_std,
  family = zero_inflated_beta() 
)
```

```{r Model 4, echo=TRUE, results='hide'}
#and the model
zi_simple_preds_4 <- 
  brm(
  formula = zi_simple_preds_4_bf,
  data = aei,
  cores = 4,
  seed = 17,
  file = "/Volumes/RachelExternal/Thesis/Thesis/fits/CL_Zero_Inflated_Fits/zi_simple_preds_4"
)
```

```{r}
print(zi_simple_preds_4)
```


## Alright, Lets Get Down to Business

Ok, lets think about this seriously and see if I can take something from this. 

```{r}
zi_mod_1_bf <-
  bf(
    irrfrac ~ yearcount + log_pop_std + log_income_std + rugged_sc,
    phi ~ 1, 
    zi ~ area_km_std, 
    family = zero_inflated_beta()
  )

get_prior(zi_mod_1_bf, data = aei)
```
```{r More models 1, echo=TRUE, results='hide'}
zi_mod_1 <-
  brm(
  formula = zi_mod_1_bf,
  data = aei,
  cores = 4,
  seed = 17,
  file = "/Volumes/RachelExternal/Thesis/Thesis/fits/CL_Zero_Inflated_Fits/zi_mods_1"
)
```
```{r}
print(zi_mod_1)
```



```{r}
zi_mod_2_bf <-
  bf(
    irrfrac ~ yearcount + log_pop_std + log_income_std + rugged_sc,
    zi ~ area_km_std, 
    family = zero_inflated_beta()
  )
```

```{r More models 2, echo=TRUE, results='hide'}
zi_mod_2 <-
  brm(
  formula = zi_mod_2_bf,
  data = aei,
  cores = 4,
  seed = 17,
  file = "/Volumes/RachelExternal/Thesis/Thesis/fits/CL_Zero_Inflated_Fits/zi_mods_2"
)
```
```{r}
print(zi_mod_2)
```
## The Posterior

```{r}
posterior_samples(zi_mod_1, pars = "b_")[,1:8] %>% 
  #the old fashioned way, brute force.
  mutate_at(c("b_phi_Intercept"), exp) %>% 
  mutate_at(vars(-c("b_phi_Intercept")), plogis) %>% 
  posterior_summary() %>% 
  as.data.frame() %>% 
  rownames_to_column("Parameter") %>% 
  kable(digits = 2)
```


```{r}
conditional_effects(zi_mod_1)
```

```{r}
posterior_samples(zi_mod_2)[,1:8] %>% 
  #the old fashioned way, brute force.
  mutate_at(c("phi"), exp) %>% 
  mutate_at(vars(-c("phi")), plogis) %>% 
  posterior_summary() %>% 
  as.data.frame() %>% 
  rownames_to_column("Parameter") %>% 
  kable(digits = 5) 

```
```{r}
conditional_effects(zi_mod_2)
```
# Its not time yet but, multilevel?
```{r}
zi_multi_1_bf <-
  bf(
    irrfrac ~ yearcount + log_pop_std + log_income_std + rugged_sc + (1 | four_regions ),
    zi ~ area_km_std + (1 | four_regions), 
    family = zero_inflated_beta()
  )
```

```{r Multi Mods , echo=TRUE, results='hide'}
zi_multi_1 <-
  brm(
  formula = zi_multi_1_bf,
  data = aei,
  cores = 4,
  seed = 17,
  file = "/Volumes/RachelExternal/Thesis/Thesis/fits/CL_Zero_Inflated_Fits/zi_multi_1"
)
```
Some Divergent Transitions here
```{r}
print(zi_multi_1)
```

### Diagnostics

### Transformed Summary Statistics

### Posterior Predictive Check

### Conditional Effects