---
title: "Data Exploration"
author: "Rachel Ledig"
date: "11/9/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Housekeeping
```{r message=FALSE, warning=FALSE}
setwd("/Volumes/RachelExternal/Thesis")

library(ggplot2)
library(brms)
library(rstan)
library(tidyverse)
library(tidybayes)
library(rgdal)
library(raster)
library(dplyr)
library(purrr)
library(sp)
library(sf)
library(rworldmap)
library(tidyr)
library(rnaturalearth)
library(rnaturalearthdata)
library(plotly)
library(magclass)
library(viridis)
library(hrbrthemes)

iFol="/Volumes/RachelExternal/Data_upload_/DataFabian/Data/"
oFol="/Volumes/RachelExternal/Thesis/DataFabian/plots/"
load(file=paste0(iFol,"data.RData"))
load(file=paste0(iFol,"datacftfrac.RData"))
load(file=paste0(iFol,"countryData_67420.RData"))
```

Data Exploration
```{r}
globalIrrigArea <- as.data.frame(globalIrrigArea)
globalIrrigArea$year <- c(1901:2005)
ggplot(globalIrrigArea, aes(year, globalIrrigArea)) + geom_point() + xlab("Year") + ylab("Irrigated Area (Mha)") 
```
This graph freaks me out.. things look too linear. Is this reminant of the fact
that all of this is coming from LPJmL? 

Let me create one large data set with all variables for each cell.
```{r}


irrFracdf <- as.data.frame(irrFrac)
irrFracdf$lat <- lat #adding spacial coords
irrFracdf$lon <- lon
irrFracdf <- irrFracdf[, c(106, 107, 1:105)]#flipping the order
names(irrFracdf)[3:107] <- c("1901":"2005")
names(irrFracdf)[1] <- "lat"
names(irrFracdf)[2] <- "lon"
irrfrac_long <- gather(irrFracdf, year, measurement, "1901":"2005", factor_key=TRUE)
colnames(irrfrac_long)[4] <- "Frac"
head(irrfrac_long, 20)


```
Lets put all the other data in long format as well. 
```{r}

d <- irrfrac_long
rm(irrfrac_long, irrFrac, irrFracdf)

#correct calculation of cell area
cellarea <- (111194.9/2)*(111194.9/2)*cos(lat/180*pi) # cellarea in m2

#Adding cell area, still m2
cellareadf <- as.data.frame(cellarea)
#repeating the cell area for all 105 years
cellarealong <- cellareadf[rep(seq_len(nrow(cellareadf)), 105), ] 
d$cellaream2 <- cellarealong
rm(cellarea, cellareadf, cellarealong)

#adding full country data
#repeating the full country data for all 105 years
fullCountryDatalong <- fullCountryData[rep(seq_len(nrow(fullCountryData)), 105), ]
d$COW <- fullCountryDatalong$COW
d$REG <- fullCountryDatalong$REG
d$countryname <- fullCountryDatalong$Countryname
d$ISO <- fullCountryDatalong$ISO
rm(fullCountryData, fullCountryDatalong)

#adding GDP
lpjGDPpcdf <- as.data.frame(lpjGDPpc)
#transforming from wide to long
lpjGDPpc_long <- gather(lpjGDPpcdf, year, GDP, V1:V105)
d$GDP <- lpjGDPpc_long$GDP
rm(lpjGDPpc, lpjGDPpc_long, lpjGDPpcdf)

#adding popdens
popdensdf <- as.data.frame(popdens)
#selecting just 1901-2005. This variable has data until 2011.
popdensdf <- subset(popdensdf, select = c(1:105))
#transforming from wide to long
popdens_long <- gather(popdensdf, year, popdens, V1:V105)
d$popdens <- popdens_long$popdens
rm(popdens, popdens_long, popdensdf)

#adding distance to next irrigated cell
distNextIrrdf <- as.data.frame(distNextIrr)
#transforming from wide to long
distNextIrr_long <- gather(distNextIrrdf, year, distnextirr, V1:V105)
d$distnextirr <- distNextIrr_long$distnextirr
rm(distNextIrr, distNextIrr_long, distNextIrrdf)

#adding mean and median increase in yield
meanincreasedf <- as.data.frame(mean_increase)
#transforming from wide to long
meanincrease_long <- gather(meanincreasedf, year, meanincrease, V1:V105)
d$meanincrease <- meanincrease_long$meanincrease
rm(mean_increase, meanincrease_long, meanincreasedf)


medianincreasedf <- as.data.frame(median_increase)
medianincrease_long <- gather(medianincreasedf, year, medianincrease, V1:V105)
d$medianincrease <- medianincrease_long$medianincrease
rm(medianincrease_long, medianincreasedf, median_increase)

#Add precipitation, Needs to be summed yearly. we don't have monthly data for anything else..
#I wont delete the data set cause for the hierarchical model perhaps the monthly description will be helpful

precipdf <- dimSums(prec_landuse, dim = 2) #we wanna sum over the months so that precip is in (mm/year)
precipdf <- as.data.frame(precipdf)
prec_long <- gather(precipdf, year, precipmm, V1:V105)
d$precipmm <- prec_long$precipmm
rm(prec_long, precipdf)

rm(lat, lon, evap_landuse, prec_landuse, runoff_landuse, interc_landuse)


#how to deal with cft frac... Cant really put it in long format... so ill leave it weird

# 1 - Temperate cereals (wheat, rye, barley; wheat)
# 2 - Rice (paddy rice; rice)
# 3 - Maize (maize for food; maize)
# 4 - Tropical cereals (millet, sorghum; millet)
# 5 - Pulses (pulses; field peas)
# 6 - Temperate roots (sugar beet; sugar beet)
# 7 - Tropical roots (cassava; cassava)
# 8 - Sunflower (sunflower; sunflower)
# 9 - Soybean (soybean; soybean)
# 10 - Groundnuts (groundnuts; groundnuts)
# 11 - Rapeseed (rapeseed; rapeseed)
# 12 - Sugarcane (sugarcane: sugarcane)
# 13 - others (potatoes, oil palm, citrus, date palm, grapes/vine, cotton, cocoa, coffee, other perennial crops, other annual crops; managed grassland)
# 14 - managed grasslands (pastures; managed grasslands)
# 15 - bio-energy grass
# 16 - bio-energy tree



cftfracdf <- as.data.frame(cftfrac)
#name the cols for the different crop codes.
crops <- c("Temperate_cereals", "Rice" ,"Maize" ,"Tropical_cereals" ,"Pulses" ,"Temperate_roots" ,"Tropical_roots", "Sunflower" ,"Soybean" ,"Groundnuts" ,"Rapeseed" ,"Sugarcane" ,"Others" ,"Managed_Grasslands" ,"Bio-energy_Grass" ,"Bio-energy_Tree")
#repeating the col names for all 105 years
colnames(cftfracdf) <- rep(crops, times = 210)
#separating the dataframe every 32nd col, to create yearly data
cftla <- lapply(seq(1, ncol(cftfracdf), by=32), function(i) 
     cftfracdf[i: pmin((i+31), ncol(cftfracdf))])
#couldnt figure this out any other way, so naming all of the separated dfs and 
#saving them as objects
nam <- "cftyear_"
val <- c(1:length(cftla))
for(i in 1:length(val)){
    assign(
        paste(nam, val, sep = "")[i], cftla[[i]]
    ) } 
#finally combining all year's dfs vertically
cftla_long <- do.call(bind_rows, mget(paste0("cftyear_", c(1:105))))
#and binding them to the big df, d
d <- bind_cols(d, cftla_long)
rm(list = paste0("cftyear_", seq(1,105,1)))
rm(cftfrac, cftfracdf, cftla, cftla_long)
rm(crops, val)


head(d, 20)
#write_csv(d, "/Volumes/RachelExternal/Thesis/Thesis/Data_longish.csv")
```


Let me do some summary statistics. Aha! This will come back in handy!
```{r}
#setting the quantiles
# p <- c(0.025, 0.5, 0.975)
# p_names <- map_chr(p, ~paste0(.x*100, "%"))
# p_funs <- map(p, ~partial(quantile, probs = .x, na.rm = TRUE)) %>% set_names(nm = p_names) 
# p_funs
# 
# 
# quants <-
#   d %>% 
#   group_by(countryname, year) %>% 
#   summarize_at(vars(Frac, GDP, distnextirr, cellaream2, popdens, medianincrease, precipmm), 
#                funs(!!!p_funs))

#reorder so this is more readable
#this is annoying, how to fix this...
#quants <- quants[, c(1,2,3,9,15,4,10,16,5,11,17,6,12,18,7,13,19,8,14,20)]

```

Ok this is annoying... can I plot it instead?
```{r}
setwd("/Volumes/RachelExternal/Thesis/Thesis/Plots")

countries = unique(d$countryname)
country_plots = list()

#ok for precipitation. Lets check out the quantiles and ranges
for(country_ in countries) {
  country_plots[[country_]] = ggplot(d %>% filter(countryname == country_), aes(x=year, y=`precipmm`)) +
    geom_boxplot() + 
    theme(axis.text.x = element_text(angle = 90)) + 
    ggtitle(country_) + 
    ylab("Grid Cell Percip (mm/year)") + 
    xlab("Year") + 
    theme(plot.title = element_text(size = 32, face = "bold", hjust=0.5), axis.title.x = element_text(size = 18, face = "bold"), axis.title.y = element_text(size = 18, face = "bold"))
  ggsave(country_plots[[country_]], file=paste0("PrecipVar_", country_,".png"), width = 44.45, height = 27.78, units = "cm", dpi=300)
}
```

```{r}
#mean yield increase
for(country_ in countries) {
  country_plots[[country_]] = ggplot(d %>% filter(countryname == country_), aes(x=year, y=`meanincrease`)) +
    geom_boxplot() + 
    theme(axis.text.x = element_text(angle = 90)) + 
    ggtitle(country_) + 
    ylab("Mean Yield Increase") + 
    xlab("Year") + 
    theme(plot.title = element_text(size = 32, face = "bold", hjust=0.5), axis.title.x = element_text(size = 18, face = "bold"), axis.title.y = element_text(size = 18, face = "bold"))
  ggsave(country_plots[[country_]], file=paste0("Meanincrease", country_,".png"), width = 44.45, height = 27.78, units = "cm", dpi=300)
}
```


```{r}
#this is for population density
for(country_ in countries) {
  country_plots[[country_]] = ggplot(d %>% filter(countryname == country_), aes(x=year, y=`popdens`)) +
    geom_boxplot() + 
    theme(axis.text.x = element_text(angle = 90)) + 
    ggtitle(country_) + 
    ylab("Grid Cell Population Density (people/km2)") + 
    xlab("Year") + 
    theme(plot.title = element_text(size = 32, face = "bold", hjust=0.5), axis.title.x = element_text(size = 18, face = "bold"), axis.title.y = element_text(size = 18, face = "bold"))
  ggsave(country_plots[[country_]], file=paste0("CellPopdens", country_,".png"), width = 44.45, height = 27.78, units = "cm", dpi=300)
}
```

```{r}
#dealing with cftfrac totals

#here I've summed all crop fractions per cell, both rainfed and irrigated
d$cftfracsum <- rowSums(d[ , c(16:47)], na.rm=TRUE)
max(d$cftfracsum) #theoretically this should be 1...
quantile(d$cftfracsum, probs = c(0, 0.025, 0.5, 0.975, 1 ))

for(country_ in countries) {
  country_plots[[country_]] = ggplot(d %>% filter(countryname == country_), aes(x=year, y=`cftfracsum`)) +
    geom_boxplot() + 
    theme(axis.text.x = element_text(angle = 90)) + 
    ggtitle(country_) + 
    ylab("Grid Cell Total Crop Fraction") + 
    xlab("Year") + 
    theme(plot.title = element_text(size = 32, face = "bold", hjust=0.5), axis.title.x = element_text(size = 18, face = "bold"), axis.title.y = element_text(size = 18, face = "bold"))
  ggsave(country_plots[[country_]], file=paste0("Total Crop Fraction", country_,".png"), width = 44.45, height = 27.78, units = "cm", dpi=300)
}

```

Lets break down cft frac a bit more and check it numbers match. 
```{r}
#lets check out just RAINFED totals
d$cftrf <- rowSums(d[ , c(16:31)], na.rm=TRUE)
max(d$cftrf) #Again slightly high, but makes sense that the value in some cells would be near 1.
mean(d$cftrf)
quantile(d$cftrf, probs = c(0, 0.025, 0.5, 0.975, 1 ))



#lets check out just IRRIGATED totals and see if they match "Frac"
d$cftirr <- rowSums(d[ , c(32:47)], na.rm=TRUE)
max(d$cftirr) #Again slightly high
mean(d$cftirr)
quantile(d$cftirr, probs = c(0, 0.025, 0.5, 0.975, 1 ))

all(d$Frac == d$cftirr) #sick, they are the same

```
Can i change these plots so that i have a stacked bar chart for each country for each year?
Gonna need to do a bit more data wrangling here to make this make sense.
```{r}

for(country_ in countries) {
  country_plots[[country_]] = d %>%
  subset(countryname == country_, select = c(3, 8, 16:47)) %>% 
  gather("croptype", "fraction", 3:34) %>% 
  group_by(year, croptype) %>%
  summarize_at(vars(fraction),
               funs(!!!p_funs, mean)) %>%  
  ggplot(aes(x= year, y= `50%`, group = croptype)) +
  geom_line(aes(y = `5%`), colour = "grey50", linetype = "dotted") +
  geom_line(aes(y = `95%`), colour = "grey50", linetype = "dotted") +
  geom_line(aes(y = mean), colour = "red") +
  geom_line() + 
  ggtitle(country_) +
  scale_x_discrete(breaks=seq(1901,2005,10)) +
  facet_wrap(~croptype, scale = "free_y", ncol = 4)
 ggsave(country_plots[[country_]], file=paste0("CftFrac", country_,".png"), width = 44.45, height = 70, units = "cm", dpi=300)
}




  
```

Ok, let's leave this as it is for now. I will add more data as I go along.  What are some ideas to explore this data.
Maybe?
- Irrigated Area vs. Country
– Irrigated area vs Crop Type
– Irrigated Area vs. Temp or humidity 
– Irrigated area vs. GDP

Lets do irrigated area per country.
steps
1. group by countries
2. calculate irrgated area $(sum(irrfrac(i) * cell area(i))/sum(cell area(country))$
3. do this for all years
4. plot in line graph for major countries
```{r}

#total irrigated fraction per country per year
#no need to divide for km2 conversion, as final is unitless
irrfractots <- d %>% group_by(countryname, year) %>% summarise(fractot = ((sum(Frac*cellaream2))/sum(cellaream2))*100)

#just checking that I did the above operation correct
d %>% group_by(ISO, year) %>% mutate(c = sum(cellaream2)) %>% summarise(c)
d %>% group_by(ISO, year) %>% mutate(f = sum(Frac*cellaream2)) %>% summarise(f)


#irrfracggp <- ggplot(irrfractots, aes(x=year,y=fractot, group = countryname, color = countryname)) + geom_line() + geom_point()
#irrfracggp <- irrfracggp + theme(legend.title = element_blank()) +
  #xlab("Year") +
  #ylab("% of Irrigated Area") +  
  #ggtitle("Fraction of area irrigated by country per year") 


```

```{r}
pirrfrac <- plot_ly(irrfractots, x = ~year, y = ~fractot, color = ~countryname, type = 'scatter', mode = 'lines')
pirrfrac <- pirrfrac %>% layout(autosize = F, height = 500) 
pirrfrac <- pirrfrac %>% layout(title = 'Total Irrigation Fraction per Country',
         xaxis = list(title = 'Year'),
         yaxis = list(title = '% Irrigated Land of Total Area'))
hide_legend(pirrfrac)
```


Ok so this one works.. I just need to figure out the colors and shit
lets do some more, GDP perhaps?

So total GDP per country is calculated by:
$sum(lpjGDP(GDP/capita) * population density(capita/km^2) * Cell Area(m^2)*0.000001) = GDP$

So total GDP per capita is calculated by:
$sum(lpjGDP(GDP/capita) * population density(capita/km^2) * Cell Area(m^2)*0.000001)/sum(sum(Population Density(capita/m^2)*cell area(m^2)*0.000001)) = GDP$

GDP has a particular feature. It is the same for all cells of a particular country. So it cannot be summed in the way that population density can be.

```{r}

#converted to km2
GDPtot <- d %>% drop_na() %>% group_by(countryname, year) %>% summarise(gdptotal = GDP*sum(popdens*cellaream2*0.000001)) %>% unique()

GDPcap <- d %>% drop_na() %>% group_by(countryname, year) %>% summarise(gdpcap = (sum(GDP*popdens*cellaream2*0.000001))/sum(popdens*cellaream2*0.000001))

#My attempt at a simple ggplot. Plotly worked better here.
#GDPggp <- ggplot(GDPtot, aes(x=year,y=gdptotal, group = countryname)) + geom_line() + geom_point()
#GDPggp <- GDPggp + theme(legend.title = element_blank()) +
  #xlab("Years") +
  #ylab("GDP (2011 USD)") +  
  #ggtitle("GDP per Country Per year") 

```

```{r}
pGDPtot <- plot_ly(GDPtot, x = ~year, y = ~gdptotal, color = ~countryname, type = 'scatter', mode = 'lines')
pGDPtot <- pGDPtot %>% layout(autosize = F, height = 500) %>% layout(title = 'GDP per Country',
         xaxis = list(title = 'Year'),
         yaxis = list(title = 'GDP (2011 USD)'))
hide_legend(pGDPtot)

pGDPcap <- plot_ly(GDPcap, x = ~year, y = ~gdpcap, color = ~countryname, type = 'scatter', mode = 'lines')
pGDPcap <- pGDPcap %>% layout(autosize = F, height = 500) %>% layout(title = 'Total GDP per Capita per Country',
         xaxis = list(title = 'Year'),
         yaxis = list(title = 'GDP/cap (2011 USD)'))
hide_legend(pGDPcap)

```


It would be cool to compare this to other GDP data. Ok it looks as if the US is three trillion off, but this could be due to the conversion of the 2011 dollar? 

And another, population by country?
$sum(Population Density(capita/km^2)*cell area(m^2)*0.000001)$
I wanna check out overall population density? So, I'll divide it by total area?
$sum(Population Density(capita/km^2)*cell area(m^2)*0.000001)/sum(cell area (m^2)*0.000001)$

```{r}
POPtot <- d %>% group_by(countryname, year) %>% summarise(poptotal = (sum(popdens*(cellaream2*0.000001))))

POPdenstot <- d %>% group_by(countryname, year) %>% summarise(popdenstotal = (sum(popdens*cellaream2*0.000001))/sum(cellaream2*0.000001))

head(POPtot)
head(POPdenstot)

#POPggp <- ggplot(POPtot, aes(x=year,y=poptotal, group = ISO)) + geom_line() + geom_point()
#POPggp <- POPggp + theme(legend.title = element_blank()) +
  #xlab("Year") +
  #ylab("Population") +  
  #ggtitle("Population per Country per Year") 


#POPDENSggp <- ggplot(POPdenstot, aes(x=year,y=popdenstotal, group = ISO)) + geom_line() + geom_point()
#POPDENSggp <- POPDENSggp + theme(legend.title = element_blank()) +
  #xlab("Year") +
  #ylab("Overall Population Density (capita/m^2") +  
  #ggtitle("Overall Population Density per Year per Country") 

```

```{r}
pPOP <- plot_ly(POPtot, x = ~year, y = ~poptotal, color = ~countryname, type = 'scatter', mode = 'lines')
pPOP <- pPOP %>% layout(autosize = F, height = 500) %>% layout(title = 'Population per Country',
         xaxis = list(title = 'Year'),
         yaxis = list(title = 'Population'))
hide_legend(pPOP)
```

```{r}
#idk if this is helpful
pPOPdens <- plot_ly(POPdenstot, x = ~year, y = ~popdenstotal, color = ~countryname, type = 'scatter', mode = 'lines')
pPOPdens <- pPOPdens %>% layout(autosize = F, height = 500) %>% layout(title = 'Overall Population Density per Country',
         xaxis = list(title = 'Year'),
         yaxis = list(title = 'Population Density (capita/m^2)'))
hide_legend(pPOPdens)
```
WTF is going on with antigua and barbuda?

Lets sum precipitation so that we can construct a little model.
```{r}
PrecipTot <- d %>% group_by(countryname, year) %>% summarise(PrecipTot = (sum(popdens*(cellaream2*0.000001))))
```




