---
title: "Fixing PET and Calculating Humidity"
date: "April 15, 2021"
output: html_document
bibliography: /Volumes/RachelExternal/Thesis/Thesis/bib.bib
---

```{r include=FALSE}
library(dplyr)
library(tidyr)
```
Lets begin with loading things.. Files you wont have. Just ask me? Annnndd... remove all the stuff you don't need so things don't get crazy on you (hopefully?).

```{r}
load("/Volumes/RachelExternal/Thesis/Thesis/Original_Grid_Data_Scripts/data/data.RData")
load("/Volumes/RachelExternal/Thesis/Thesis/Original_Grid_Data_Scripts/data/pet_data.RData")
load("/Volumes/RachelExternal/Thesis/Thesis/Original_Grid_Data_Scripts/data/countryData_67420.RData")

rm(distNextIrr, irrFrac, lpjGDPpc, mean_increase, median_increase, popdens, discharge_landuse, globalIrrigArea, lat, lon, runoff_landuse, transp_blue_landuse, transp_green_landuse)
```

So `fullCountryData` has all of the info we need to create the structure of this data frame, let's start working things in to it. 
```{r}
structure <- fullCountryData
rm(fullCountryData)

#flip structure around so its long
structure <- do.call("rbind", replicate(105, structure, simplify = FALSE))

#add a col for year
structure$year <- rep(1901:2005, each=67420)
```

Ok so the game plan here: PET data has the structure [1:67420, 1:12, 1:105], meaning that it has a layer for every month of every year for all 67420 grid cells. Precipitation has the same structure. We want average humidity and average PET for a country for a year. 

In @neumannExploringGlobalIrrigation2011, humidity is calculated as: $Precip/PET$. We shall follow the same logic here. First lets make a data frame that maintains all of the resolution contained in the mm/mo/year dataset of total_PET and prec_landuse

```{r}
PET <- as.data.frame(total_PET)
rm(total_PET)

Precip <- as.data.frame(prec_landuse)
rm(prec_landuse)
```

Nothing has crashed yet, on thin ice though.  

So PET and precip are in the same format. Can I just divide them element wise to calculate humidity? (I put the 1 in there so we don't get inf values :) )

```{r}
humidity <- Precip/(PET+1)
```

We need months.
```{r}
mo <- seq(1,12,1)
```

And lets separate this the clunkiest way possible. I am proud I found such a horrible solution. 
```{r}
#repeating the col names for all 105 years
colnames(humidity) <- rep(mo, 105)
#separating the dataframe every 32nd col, to create yearly data
humla <- lapply(seq(1, ncol(humidity), by=12), function(i) 
     humidity[i: pmin((i+11), ncol(humidity))])
#couldnt figure this out any other way, so naming all of the separated dfs and 
#saving them as objects
nam <- "humyear_"
val <- c(1:length(humla))
for(i in 1:length(val)){
    assign(
        paste(nam, val, sep = "")[i], humla[[i]]
    ) } 
#finally combining all year's dfs vertically
humidity_long <- do.call(bind_rows, mget(paste0("humyear_", c(1:105))))
#and binding them to the big df
humidity_full <- bind_cols(structure, humidity_long)
rm(list = paste0("humyear_", seq(1,105,1)))
rm(humidity, humla, humidity_long)


names(humidity_full)[6:17] <- paste("Humidity", 1:ncol(humidity_full[,6:17]), sep = "")

#long data
humidity_full <- 
    humidity_full %>% 
    pivot_longer(cols = Humidity1:Humidity12,
                 names_to = "Months", 
                 values_to = "Humidity")

humidity_full$year <- as.factor(humidity_full$year)

#i hate this solution. 
humidity_full <- subset(humidity_full, year == "1901"| year == "1910"| year == "1920"| year == "1930"| year == "1940"| year == "1950"| year == "1960"| year == "1970"| year == "1980"| year == "1985"| year == "1990"| year == "1995"| year == "2000"| year == "2005")


```

lets do it again for PET only

```{r}

#repeating the col names for all 105 years
colnames(PET) <- rep(mo, 105)
#separating the dataframe every 32nd col, to create yearly data
PETla <- lapply(seq(1, ncol(PET), by=12), function(i) 
     PET[i: pmin((i+11), ncol(PET))])
#couldnt figure this out any other way, so naming all of the separated dfs and 
#saving them as objects
nam <- "PETyear_"
val <- c(1:length(PETla))
for(i in 1:length(val)){
    assign(
        paste(nam, val, sep = "")[i], PETla[[i]]
    ) } 
#finally combining all year's dfs vertically
PET_long <- do.call(bind_rows, mget(paste0("PETyear_", c(1:105))))
#and binding them to the big df
PET_full <- bind_cols(structure, PET_long)
rm(list = paste0("PETyear_", seq(1,105,1)))
rm(PET, PETla, PET_long)

names(PET_full)[6:17] <- paste("PET", 1:ncol(PET_full[,6:17]), sep = "")

#long data
PET_full <- 
    PET_full %>% 
    pivot_longer(cols = PET1:PET12,
                 names_to = "Months", 
                 values_to = "PET")

PET_full$year <- as.factor(PET_full$year)

#i hate this solution. 
PET_full <- subset(PET_full, year == "1901"| year == "1910"| year == "1920"| year == "1930"| year == "1940"| year == "1950"| year == "1960"| year == "1970"| year == "1980"| year == "1985"| year == "1990"| year == "1995"| year == "2000"| year == "2005")

```

And finally for precip. But as Fabian mentioned before, I need to sum to $m^3/year$. Knowing that $1 mm = 1 \frac{L}{m^2}$ we can get somewhere. Ill throw cell area in to the dataframe too.

```{r}
#repeating the col names for all 105 years
colnames(Precip) <- rep(mo, 105)
#separating the dataframe every 32nd col, to create yearly data
Precipla <- lapply(seq(1, ncol(Precip), by=12), function(i) 
     Precip[i: pmin((i+11), ncol(Precip))])
#couldnt figure this out any other way, so naming all of the separated dfs and 
#saving them as objects
nam <- "Precipyear_"
val <- c(1:length(Precipla))
for(i in 1:length(val)){
    assign(
        paste(nam, val, sep = "")[i], Precipla[[i]]
    ) } 
#finally combining all year's dfs vertically
Precip_long <- do.call(bind_rows, mget(paste0("Precipyear_", c(1:105))))
#and binding them to the big df
Precip_full <- bind_cols(structure, Precip_long)
rm(list = paste0("Precipyear_", seq(1,105,1)))
rm(Precip, Precipla, Precip_long)

names(Precip_full)[6:17] <- paste("Precipmm", 1:ncol(Precip_full[,6:17]), sep = "")

#and cell area
cellarea <- as.data.frame(cellarea)
cellarea <- do.call("rbind", replicate(105, cellarea, simplify = FALSE))
Precip_full <- cbind(cellarea, Precip_full)
rm(cellarea)

#long data
Precip_full <- 
    Precip_full %>% 
    pivot_longer(cols = Precipmm1:Precipmm12,
                 names_to = "Months", 
                 values_to = "Precip")

Precip_full$year <- as.factor(Precip_full$year)

#i hate this solution. 
Precip_full <- subset(Precip_full, year == "1901"| year == "1910"| year == "1920"| year == "1930"| year == "1940"| year == "1950"| year == "1960"| year == "1970"| year == "1980"| year == "1985"| year == "1990"| year == "1995"| year == "2000"| year == "2005")
```

Stragglers.
```{r}
rm(total_PET_allcrops, interc_landuse, evap_landuse)
```


For some reason it lists all of the years as the factor levels but it has done the subsetting. Idk why, but its getting late so I just accept it at this point.  


***

Alright, game time.  

So humidity (remember $Precip/PET$ which is in the units of mm and mm/mo respectively) cannot be summed, we need to take the average of all grid cells during all months for each country and year.The same goes for PET, we will take the average of all data cells per country during one year. 


First, Humidity we just need the mean and the median of all the values for a year and for a country. This should do it.
```{r}
humidity_full <-
    humidity_full %>% 
    group_by(ISO, year) %>% 
    summarise(meanHumid = mean(Humidity), 
              medHumid = median(Humidity))

head(humidity_full, 20)
    
```

Now, PET. Same procedure as humidity.
```{r}
PET_full <-
    PET_full %>% 
    group_by(ISO, year) %>% 
    summarise(meanPET = mean(PET), 
              medPET = median(PET))

head(PET_full, 20)
    
```

Precippp.... I work out the units here just for me to keep things clear.
$$
\begin{aligned}
1 mm  &= \frac{1 L}{m^2} \\
&= \frac{1 L}{m^2} * cellarea(m^2) \\
&= 1L \\
&= 0.001 m^3
\end{aligned}
$$
So I'm gonna multiply things by their cell area then average/take the median. 

```{r}

Precip_full <- 
    Precip_full %>% 
    mutate(precipm3 = Precip*cellarea*0.001) %>% 
    group_by(ISO,year) %>% 
    summarise(cubM_precip = sum(precipm3))
    

head(Precip_full, 20)
```
All looks good. Lets merge these and save them as a csv file. 
```{r}
hum_pet_prec <-
    merge(humidity_full, PET_full) %>% 
    merge(Precip_full)

head(hum_pet_prec, 20)

#write.csv(hum_pet_prec, "/Volumes/RachelExternal/Thesis/Thesis/Data/hum_pet_prec.csv")
```
Wahoo!!
