---
title: "Country Level Predictor Standardization"
author: "Rachel Ledig"
date: "4/20/2021"
output: 
  html_document:
    toc: true
    toc_float: true
bibliography: /Volumes/RachelExternal/Thesis/Thesis/bib.bib
---
```{r include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)

library(ggplot2)
library(plyr)
library(dplyr)
library(tidyverse)
library(gridExtra)
library(plotly)
```


```{r data and functions}
aei_no_std <- read.csv("/Volumes/RachelExternal/Thesis/Data_upload_for_CL/AEI_NoStd.csv") #some data
source("/Volumes/RachelExternal/Thesis/Thesis/Thesis_Functions.R") #some functions
```

# Standardization


We've got a lot of things here to log, scale or center. 

What do I mean by scaling or centering? 

**Normalizing**: Shifting the range of the predictor between [0,1] using the formula: $\frac{y_i-min(y)}{max(y)-min(y)}$. This is useful because for some predictors it is helpul to have a notion of how big or small they are with the constraints of 0 and 1.    

**Centering**: Centering the mean of the predictor on 0 or 1 using the formula: $\frac{y_i-mean(y)}{2sd(y)}$. The classic with a twist. By using 2 standard deviations, 95% of the data is contained within 2 sd, convient. 

Why must I do this? 

Well, a couple of reasons. First and foremost it makes the specification of priors easier, as the distribution of the parameter already centered around the mean and most of the data points are contained within one standard deviation to each side. Another reason is that it makes the interpretation of the coeficients a bit easier, as you can clearly tell which have positive or negative effects.

A lot of these predictors I assume accumulate exponentially. Population, Income and total GDP are obvious ones. Some others seem to be exponential as well, given the distribution of their histograms, such as majority of the crop fractions. However, variables like irrperc/irrfrac and the crop fractions have a hard time being transformed using a log transformation because they are so zero inflated. As Log(0) = -Inf, this causes a lot of problems. I will keep them in their original scale. 


- _population_ - Log and Center
- _income_ - Log and Center
- _GDP_ - Log and Center
- _Median Humidity_ - Center
- _Median PET_ - Center
- _Precip_ - Center
- _Ruggedness_ - Scale
- _Crop Fractoins_ - Center

```{r}
summary(aei_no_std)
```


```{r}
#these functions are coming from Thesis_Functions.R, loaded in the beginning

aei_std <-
  aei_no_std %>% 
  select(-c(20, 21,24,28)) %>% #removing mean Humid and PET
  mutate(across(c(17:19, 23), log)) %>% 
  #centering by dividing by 2sd, so that 95% of the distribution is from [-1,1]
  mutate(across(c(8, 17:19, 20:23, 25:50), centered)) %>% 
  mutate(across(c(24, 51), normalized)) %>% 
  mutate(polity_case = as.factor(polity_case))

summary(aei_std)

cols <- c("ISO","country",  "four_regions", "six_regions", "eight_regions", "World.bank.region", "polity_case")

aei_std <- aei_std  %>% 
  mutate_each_(funs(factor(.)),cols)

str(aei_std)
  

#write.csv(aei_std, "/Volumes/RachelExternal/Thesis/Data_upload_for_CL/AEI_Std_polity.csv")
```


# Histograms



Lets plot these before and after being transformed.

## Main Predictors
```{r predictor graphs, echo=FALSE}
pop_raw <-
  aei_no_std %>% 
  drop_na(population) %>% 
  ggplot(aes(x=population, color=six_regions)) +
  geom_histogram(fill="white", position="dodge", bins = 100) + 
  scale_color_brewer(palette="Dark2") + ggtitle("Total Population")
pop_std <-
  aei_std %>% 
  drop_na(population) %>% 
  ggplot(aes(x=population, color=six_regions)) +
  geom_histogram(fill="white", position="dodge", bins = 100) + 
  scale_color_brewer(palette="Dark2") + ggtitle("Total Population - Log and Center")
grid.arrange(pop_raw, pop_std, nrow = 2)

income_raw <-
  aei_no_std %>% 
  drop_na(income) %>% 
  ggplot(aes(x=income, color=six_regions)) +
  geom_histogram(fill="white", position="dodge", bins = 100) + 
  scale_color_brewer(palette="Dark2") + ggtitle("Income per Capita")
income_std <-
  aei_std %>% 
  drop_na(income) %>% 
  ggplot(aes(x=income, color=six_regions)) +
  geom_histogram(fill="white", position="dodge", bins = 100) + 
  scale_color_brewer(palette="Dark2") + ggtitle("Income per Capita - Log and Center")
grid.arrange(income_raw, income_std, nrow = 2)


GDP_raw <-
  aei_no_std %>% 
  drop_na(GDPtot) %>% 
  ggplot(aes(x=GDPtot, color=six_regions)) +
  geom_histogram(fill="white", position="dodge", bins = 100) + 
  scale_color_brewer(palette="Dark2") + ggtitle("Total GDP")
GDP_std <-
  aei_std %>% 
  drop_na(GDPtot) %>% 
  ggplot(aes(x=GDPtot, color=six_regions)) +
  geom_histogram(fill="white", position="dodge", bins = 100) + 
  scale_color_brewer(palette="Dark2") + ggtitle("Total GDP - Log and Center")
grid.arrange(GDP_raw, GDP_std, nrow = 2)


medhum_raw <-
  aei_no_std %>% 
  drop_na(medHumid) %>% 
  ggplot(aes(x=medHumid, color=six_regions)) +
  geom_histogram(fill="white", position="dodge", bins = 100) + 
  scale_color_brewer(palette="Dark2") + ggtitle("Median Humidity")
medhum_std <-
  aei_std %>% 
  drop_na(medHumid) %>% 
  ggplot(aes(x=medHumid, color=six_regions)) +
  geom_histogram(fill="white", position="dodge", bins = 100) + 
  scale_color_brewer(palette="Dark2") + ggtitle("Median Humidity - Center")
grid.arrange(medhum_raw, medhum_std, nrow = 2)


medpet_raw <-
  aei_no_std %>% 
  drop_na(medPET) %>% 
  ggplot(aes(x=medPET, color=six_regions)) +
  geom_histogram(fill="white", position="dodge", bins = 100) + 
  scale_color_brewer(palette="Dark2") + ggtitle("Median PET")
medpet_std <-
  aei_std %>% 
  drop_na(medPET) %>% 
  ggplot(aes(x=medPET, color=six_regions)) +
  geom_histogram(fill="white", position="dodge", bins = 100) + 
  scale_color_brewer(palette="Dark2") + ggtitle("Median PET - Center")
grid.arrange(medpet_raw, medpet_std, nrow = 2)

precip_raw <-
  aei_no_std %>% 
  drop_na(cubM_precip) %>% 
  ggplot(aes(x=cubM_precip, color=six_regions)) +
  geom_histogram(fill="white", position="dodge", bins = 100) + 
  scale_color_brewer(palette="Dark2") + ggtitle("Precipitation m^3")
precip_std <-
  aei_std %>% 
  drop_na(cubM_precip) %>% 
  ggplot(aes(x=cubM_precip, color=six_regions)) +
  geom_histogram(fill="white", position="dodge", bins = 100) + 
  scale_color_brewer(palette="Dark2") + ggtitle("Precipitation - Log and Center")
grid.arrange(precip_raw, precip_std, nrow = 2)

rugged_raw <-
  aei_no_std %>% 
  drop_na(rugged) %>% 
  ggplot(aes(x=rugged, color=six_regions)) +
  geom_histogram(fill="white", position="dodge", bins = 100) + 
  scale_color_brewer(palette="Dark2") + ggtitle("Ruggedness") 
rugged_std <-
  aei_std %>% 
  drop_na(rugged) %>% 
  ggplot(aes(x=rugged, color=six_regions)) +
  geom_histogram(fill="white", position="dodge", bins = 100) + 
  scale_color_brewer(palette="Dark2") + ggtitle("Ruggedness - Scaled")
grid.arrange(rugged_raw, rugged_std, nrow = 2)
  
```

The transformations of the predictors look good, and those variables that were logged and then centered look great, much more normally distributed 

## Crop Fractions

What about crop fractions?

```{r temperate cereals, echo=FALSE}
# temp_cereals_rain_raw <-
#   aei_no_std %>% 
#   drop_na(Temperate_cereals) %>% 
#   ggplot(aes(x=Temperate_cereals, color=six_regions)) +
#   geom_histogram(fill="white", position="dodge", bins = 100) + 
#   scale_color_brewer(palette="Dark2") + ggtitle("Temprate Cereals - Rainfed")
# temp_cereals_irr_raw <-
#   aei_no_std %>% 
#   drop_na(Temperate_cereals.1) %>% 
#   ggplot(aes(x=Temperate_cereals.1, color=six_regions)) +
#   geom_histogram(fill="white", position="dodge", bins = 100) + 
#   scale_color_brewer(palette="Dark2") + ggtitle("Temprate Cereals - Irrigated")
# 
# 
# temp_cereals_rain_std <-
#   aei_std %>% 
#   drop_na(Temperate_cereals) %>% 
#   ggplot(aes(x=Temperate_cereals, color=six_regions)) +
#   geom_histogram(fill="white", position="dodge", bins = 100) + 
#   scale_color_brewer(palette="Dark2") + ggtitle("Temprate Cereals - Rainfed - Centered")
# temp_cereals_irr_std <-
#   aei_std %>% 
#   drop_na(Temperate_cereals.1) %>% 
#   ggplot(aes(x=Temperate_cereals.1, color=six_regions)) +
#   geom_histogram(fill="white", position="dodge", bins = 100) + 
#   scale_color_brewer(palette="Dark2") + ggtitle("Temprate Cereals - Irrigated - Centered")
# 
# 
# grid.arrange(temp_cereals_rain_raw,temp_cereals_irr_raw, temp_cereals_rain_std, temp_cereals_irr_std, nrow = 4)
```

```{r rice, echo=FALSE}
# rice_rain_raw <-
#   aei_no_std %>% 
#   drop_na(Rice) %>% 
#   ggplot(aes(x=Rice, color=six_regions)) +
#   geom_histogram(fill="white", position="dodge", bins = 100) + 
#   scale_color_brewer(palette="Dark2") + ggtitle("Rice - Rainfed")
# rice_irr_raw <-
#   aei_no_std %>% 
#   drop_na(Rice.1) %>% 
#   ggplot(aes(x=Rice.1, color=six_regions)) +
#   geom_histogram(fill="white", position="dodge", bins = 100) + 
#   scale_color_brewer(palette="Dark2") + ggtitle("Rice - Irrigated")
# 
# 
# rice_rain_std <-
#   aei_std %>% 
#   drop_na(Rice) %>% 
#   ggplot(aes(x=Rice, color=six_regions)) +
#   geom_histogram(fill="white", position="dodge", bins = 100) + 
#   scale_color_brewer(palette="Dark2") + ggtitle("Rice - Rainfed - Centered")
# rice_irr_std <-
#   aei_std %>% 
#   drop_na(Rice.1) %>% 
#   ggplot(aes(x=Rice.1, color=six_regions)) +
#   geom_histogram(fill="white", position="dodge", bins = 100) + 
#   scale_color_brewer(palette="Dark2") + ggtitle("Rice - Irrigated - Centered")
# 
# 
# grid.arrange(rice_rain_raw, rice_irr_raw, rice_rain_std, rice_irr_std, nrow = 4)
```

```{r maize, echo=FALSE}
# maize_rain_raw <-
#   aei_no_std %>% 
#   drop_na(Maize) %>% 
#   ggplot(aes(x=Maize, color=six_regions)) +
#   geom_histogram(fill="white", position="dodge", bins = 100) + 
#   scale_color_brewer(palette="Dark2") + ggtitle("Maize - Rainfed")
# maize_irr_raw <-
#   aei_no_std %>% 
#   drop_na(Maize.1) %>% 
#   ggplot(aes(x=Maize.1, color=six_regions)) +
#   geom_histogram(fill="white", position="dodge", bins = 100) + 
#   scale_color_brewer(palette="Dark2") + ggtitle("Maize - Irrigated")
# 
# maize_rain_std <-
#   aei_std %>% 
#   drop_na(Maize) %>% 
#   ggplot(aes(x=Maize, color=six_regions)) +
#   geom_histogram(fill="white", position="dodge", bins = 100) + 
#   scale_color_brewer(palette="Dark2") + ggtitle("Maize - Rainfed - Centered")
# maize_irr_std <-
#   aei_std %>% 
#   drop_na(Maize.1) %>% 
#   ggplot(aes(x=Maize.1, color=six_regions)) +
#   geom_histogram(fill="white", position="dodge", bins = 100) + 
#   scale_color_brewer(palette="Dark2") + ggtitle("Maize - Irrigated - Centered")
# 
# grid.arrange(maize_rain_raw,maize_irr_raw, maize_rain_std, maize_irr_std, nrow = 4)
```

```{r tropical cereals, echo=FALSE}
# tc_rain_raw <-
#   aei_no_std %>% 
#   drop_na(Tropical_cereals) %>% 
#   ggplot(aes(x=Tropical_cereals, color=six_regions)) +
#   geom_histogram(fill="white", position="dodge", bins = 100) + 
#   scale_color_brewer(palette="Dark2") + ggtitle("Tropical Cereals - Rainfed")
# tc_irr_raw <-
#   aei_no_std %>% 
#   drop_na(Tropical_cereals.1) %>% 
#   ggplot(aes(x=Tropical_cereals.1, color=six_regions)) +
#   geom_histogram(fill="white", position="dodge", bins = 100) + 
#   scale_color_brewer(palette="Dark2") + ggtitle("Tropical Cereals - Irrigated")
# 
# tc_rain_std <-
#   aei_std %>% 
#   drop_na(Tropical_cereals) %>% 
#   ggplot(aes(x=Tropical_cereals, color=six_regions)) +
#   geom_histogram(fill="white", position="dodge", bins = 100) + 
#   scale_color_brewer(palette="Dark2") + ggtitle("Tropical Cereals - Rainfed - Centered")
# tc_irr_std <-
#   aei_std %>% 
#   drop_na(Tropical_cereals.1) %>% 
#   ggplot(aes(x=Tropical_cereals.1, color=six_regions)) +
#   geom_histogram(fill="white", position="dodge", bins = 100) + 
#   scale_color_brewer(palette="Dark2") + ggtitle("Tropical Cereals - Irrigated - Centered")
# 
# grid.arrange(tc_rain_raw,tc_irr_raw, tc_rain_std, tc_irr_std, nrow = 4)
```

```{r pulses, echo=FALSE}
# pulses_rain_raw <-
#   aei_no_std %>% 
#   drop_na(Pulses) %>% 
#   ggplot(aes(x=Pulses, color=six_regions)) +
#   geom_histogram(fill="white", position="dodge", bins = 100) + 
#   scale_color_brewer(palette="Dark2") + ggtitle("Pulses - Rainfed")
# pulses_irr_raw <-
#   aei_no_std %>% 
#   drop_na(Pulses.1) %>% 
#   ggplot(aes(x=Pulses.1, color=six_regions)) +
#   geom_histogram(fill="white", position="dodge", bins = 100) + 
#   scale_color_brewer(palette="Dark2") + ggtitle("Pulses - Irrigated")
# 
# pulses_rain_std <-
#   aei_std %>% 
#   drop_na(Pulses) %>% 
#   ggplot(aes(x=Pulses, color=six_regions)) +
#   geom_histogram(fill="white", position="dodge", bins = 100) + 
#   scale_color_brewer(palette="Dark2") + ggtitle("Pulses - Rainfed - Centered")
# pulses_irr_std <-
#   aei_std %>% 
#   drop_na(Pulses.1) %>% 
#   ggplot(aes(x=Pulses.1, color=six_regions)) +
#   geom_histogram(fill="white", position="dodge", bins = 100) + 
#   scale_color_brewer(palette="Dark2") + ggtitle("Pulses - Irrigated - Centered")
# 
# grid.arrange(pulses_rain_raw,pulses_irr_raw, pulses_rain_std, pulses_irr_std, nrow = 4)
```

```{r temperate roots, echo=FALSE}
# tr_rain_raw <-
#   aei_no_std %>% 
#   drop_na(Temperate_roots) %>% 
#   ggplot(aes(x=Temperate_roots, color=six_regions)) +
#   geom_histogram(fill="white", position="dodge", bins = 100) + 
#   scale_color_brewer(palette="Dark2") + ggtitle("Temprate Roots - Rainfed")
# tr_irr_raw <-
#   aei_no_std %>% 
#   drop_na(Temperate_roots.1) %>% 
#   ggplot(aes(x=Temperate_roots.1, color=six_regions)) +
#   geom_histogram(fill="white", position="dodge", bins = 100) + 
#   scale_color_brewer(palette="Dark2") + ggtitle("Temprate Roots - Irrigated")
# 
# tr_rain_std <-
#   aei_std %>% 
#   drop_na(Temperate_roots) %>% 
#   ggplot(aes(x=Temperate_roots, color=six_regions)) +
#   geom_histogram(fill="white", position="dodge", bins = 100) + 
#   scale_color_brewer(palette="Dark2") + ggtitle("Temprate Roots - Rainfed - Centered")
# tr_irr_std <-
#   aei_std %>% 
#   drop_na(Temperate_roots.1) %>% 
#   ggplot(aes(x=Temperate_roots.1, color=six_regions)) +
#   geom_histogram(fill="white", position="dodge", bins = 100) + 
#   scale_color_brewer(palette="Dark2") + ggtitle("Temprate Roots - Irrigated - Centered")
# 
# grid.arrange(tr_rain_raw,tr_irr_raw, tr_rain_std, tr_irr_std, nrow = 4)
```

```{r tropical roots, echo=FALSE}
# trr_rain_raw <-
#   aei_no_std %>% 
#   drop_na(Tropical_roots) %>% 
#   ggplot(aes(x=Tropical_roots, color=six_regions)) +
#   geom_histogram(fill="white", position="dodge", bins = 100) + 
#   scale_color_brewer(palette="Dark2") + ggtitle("Tropical Roots - Rainfed")
# 
# trr_rain_std <-
#   aei_std %>% 
#   drop_na(Tropical_roots) %>% 
#   ggplot(aes(x=Tropical_roots, color=six_regions)) +
#   geom_histogram(fill="white", position="dodge", bins = 100) + 
#   scale_color_brewer(palette="Dark2") + ggtitle("Tropical Roots - Rainfed - Centered")
# 
# grid.arrange(trr_rain_raw, trr_rain_std, nrow = 2)
```

```{r sunflower, echo=FALSE}
# sf_rain_raw <-
#   aei_no_std %>% 
#   drop_na(Sunflower) %>% 
#   ggplot(aes(x=Sunflower, color=six_regions)) +
#   geom_histogram(fill="white", position="dodge", bins = 100) + 
#   scale_color_brewer(palette="Dark2") + ggtitle("Sunflower - Rainfed")
# sf_irr_raw <-
#   aei_no_std %>% 
#   drop_na(Sunflower.1) %>% 
#   ggplot(aes(x=Sunflower.1, color=six_regions)) +
#   geom_histogram(fill="white", position="dodge", bins = 100) + 
#   scale_color_brewer(palette="Dark2") + ggtitle("Sunflower - Irrigated")
# 
# sf_rain_std <-
#   aei_std %>% 
#   drop_na(Sunflower) %>% 
#   ggplot(aes(x=Sunflower, color=six_regions)) +
#   geom_histogram(fill="white", position="dodge", bins = 100) + 
#   scale_color_brewer(palette="Dark2") + ggtitle("Sunflower - Rainfed - Centered")
# sf_irr_std <-
#   aei_std %>% 
#   drop_na(Sunflower.1) %>% 
#   ggplot(aes(x=Sunflower.1, color=six_regions)) +
#   geom_histogram(fill="white", position="dodge", bins = 100) + 
#   scale_color_brewer(palette="Dark2") + ggtitle("Sunflower - Irrigated - Centered")
# 
# grid.arrange(sf_rain_raw,sf_irr_raw, sf_rain_std, sf_irr_std, nrow = 4)
```

```{r soy, echo=FALSE}
# soy_rain_raw <-
#   aei_no_std %>% 
#   drop_na(Soybean) %>% 
#   ggplot(aes(x=Soybean, color=six_regions)) +
#   geom_histogram(fill="white", position="dodge", bins = 100) + 
#   scale_color_brewer(palette="Dark2") + ggtitle("Soy - Rainfed")
# soy_irr_raw <-
#   aei_no_std %>% 
#   drop_na(Soybean.1) %>% 
#   ggplot(aes(x=Soybean.1, color=six_regions)) +
#   geom_histogram(fill="white", position="dodge", bins = 100) + 
#   scale_color_brewer(palette="Dark2") + ggtitle("Soy - Irrigated")
# 
# soy_rain_std <-
#   aei_std %>% 
#   drop_na(Soybean) %>% 
#   ggplot(aes(x=Soybean, color=six_regions)) +
#   geom_histogram(fill="white", position="dodge", bins = 100) + 
#   scale_color_brewer(palette="Dark2") + ggtitle("Soy - Rainfed - Centered")
# soy_irr_std <-
#   aei_std %>% 
#   drop_na(Soybean.1) %>% 
#   ggplot(aes(x=Soybean.1, color=six_regions)) +
#   geom_histogram(fill="white", position="dodge", bins = 100) + 
#   scale_color_brewer(palette="Dark2") + ggtitle("Soy - Irrigated - Centered")
# 
# grid.arrange(soy_rain_raw,soy_irr_raw, soy_rain_std, soy_irr_std, nrow = 4)
```

```{r groundnuts, echo=FALSE}
# gns_rain_raw <-
#   aei_no_std %>% 
#   drop_na(Groundnuts) %>% 
#   ggplot(aes(x=Groundnuts, color=six_regions)) +
#   geom_histogram(fill="white", position="dodge", bins = 100) + 
#   scale_color_brewer(palette="Dark2") + ggtitle("Groundnuts - Rainfed")
# gns_irr_raw <-
#   aei_no_std %>% 
#   drop_na(Groundnuts.1) %>% 
#   ggplot(aes(x=Groundnuts.1, color=six_regions)) +
#   geom_histogram(fill="white", position="dodge", bins = 100) + 
#   scale_color_brewer(palette="Dark2") + ggtitle("Groundnuts - Irrigated")
# 
# gns_rain_std <-
#   aei_std %>% 
#   drop_na(Groundnuts) %>% 
#   ggplot(aes(x=Groundnuts, color=six_regions)) +
#   geom_histogram(fill="white", position="dodge", bins = 100) + 
#   scale_color_brewer(palette="Dark2") + ggtitle("Groundnuts - Rainfed - Centered")
# gns_irr_std <-
#   aei_std %>% 
#   drop_na(Groundnuts.1) %>% 
#   ggplot(aes(x=Groundnuts.1, color=six_regions)) +
#   geom_histogram(fill="white", position="dodge", bins = 100) + 
#   scale_color_brewer(palette="Dark2") + ggtitle("Groundnuts - Irrigated - Centered")
# 
# grid.arrange(gns_rain_raw,gns_irr_raw, gns_rain_std, gns_irr_std, nrow = 4)
```

```{r rapeseed, echo=FALSE}
# rps_rain_raw <-
#   aei_no_std %>% 
#   drop_na(Rapeseed) %>% 
#   ggplot(aes(x=Rapeseed, color=six_regions)) +
#   geom_histogram(fill="white", position="dodge", bins = 100) + 
#   scale_color_brewer(palette="Dark2") + ggtitle("Rapeseed - Rainfed")
# rps_irr_raw <-
#   aei_no_std %>% 
#   drop_na(Rapeseed.1) %>% 
#   ggplot(aes(x=Rapeseed.1, color=six_regions)) +
#   geom_histogram(fill="white", position="dodge", bins = 100) + 
#   scale_color_brewer(palette="Dark2") + ggtitle("Rapeseed - Irrigated")
# 
# rps_rain_std <-
#   aei_std %>% 
#   drop_na(Rapeseed) %>% 
#   ggplot(aes(x=Rapeseed, color=six_regions)) +
#   geom_histogram(fill="white", position="dodge", bins = 100) + 
#   scale_color_brewer(palette="Dark2") + ggtitle("Rapeseed - Rainfed - Centered")
# rps_irr_std <-
#   aei_std %>% 
#   drop_na(Rapeseed.1) %>% 
#   ggplot(aes(x=Rapeseed.1, color=six_regions)) +
#   geom_histogram(fill="white", position="dodge", bins = 100) + 
#   scale_color_brewer(palette="Dark2") + ggtitle("Rapeseed - Irrigated - Centered")
# 
# grid.arrange(rps_rain_raw,rps_irr_raw, rps_rain_std, rps_irr_std, nrow = 4)
```

```{r sugar, echo=FALSE}
# sugar_rain_raw <-
#   aei_no_std %>% 
#   drop_na(Sugarcane) %>% 
#   ggplot(aes(x=Sugarcane, color=six_regions)) +
#   geom_histogram(fill="white", position="dodge", bins = 100) + 
#   scale_color_brewer(palette="Dark2") + ggtitle("Sugarcane - Rainfed")
# sugar_irr_raw <-
#   aei_no_std %>% 
#   drop_na(Sugarcane.1) %>% 
#   ggplot(aes(x=Sugarcane.1, color=six_regions)) +
#   geom_histogram(fill="white", position="dodge", bins = 100) + 
#   scale_color_brewer(palette="Dark2") + ggtitle("Sugarcane - Irrigated")
# 
# sugar_rain_std <-
#   aei_std %>% 
#   drop_na(Sugarcane) %>% 
#   ggplot(aes(x=Sugarcane, color=six_regions)) +
#   geom_histogram(fill="white", position="dodge", bins = 100) + 
#   scale_color_brewer(palette="Dark2") + ggtitle("Sugarcane - Rainfed - Centered")
# sugar_irr_std <-
#   aei_std %>% 
#   drop_na(Sugarcane.1) %>% 
#   ggplot(aes(x=Sugarcane.1, color=six_regions)) +
#   geom_histogram(fill="white", position="dodge", bins = 100) + 
#   scale_color_brewer(palette="Dark2") + ggtitle("Sugarcane - Irrigated - Centered")
# 
# grid.arrange(sugar_rain_raw,sugar_irr_raw, sugar_rain_std, sugar_irr_std, nrow = 4)
```



```{r others, echo=FALSE}
# others_rain_raw <-
#   aei_no_std %>% 
#   drop_na(Others) %>% 
#   ggplot(aes(x=Others, color=six_regions)) +
#   geom_histogram(fill="white", position="dodge", bins = 100) + 
#   scale_color_brewer(palette="Dark2") + ggtitle("Others - Rainfed")
# others_irr_raw <-
#   aei_no_std %>% 
#   drop_na(Others.1) %>% 
#   ggplot(aes(x=Others.1, color=six_regions)) +
#   geom_histogram(fill="white", position="dodge", bins = 100) + 
#   scale_color_brewer(palette="Dark2") + ggtitle("Others - Irrigated")
# 
# others_rain_std <-
#   aei_std %>% 
#   drop_na(Others) %>% 
#   ggplot(aes(x=Others, color=six_regions)) +
#   geom_histogram(fill="white", position="dodge", bins = 100) + 
#   scale_color_brewer(palette="Dark2") + ggtitle("Others - Rainfed - Centered")
# others_irr_std <-
#   aei_std %>% 
#   drop_na(Others.1) %>% 
#   ggplot(aes(x=Others.1, color=six_regions)) +
#   geom_histogram(fill="white", position="dodge", bins = 100) + 
#   scale_color_brewer(palette="Dark2") + ggtitle("Others - Irrigated - Centered")
# 
# grid.arrange(others_rain_raw,others_irr_raw, others_rain_std, others_irr_std, nrow = 4)
```

```{r managed grass, echo=FALSE}
# mg_rain_raw <-
#   aei_no_std %>% 
#   drop_na(Managed_Grasslands) %>% 
#   ggplot(aes(x=Managed_Grasslands, color=six_regions)) +
#   geom_histogram(fill="white", position="dodge", bins = 100) + 
#   xlim(-1,1) +
#   scale_color_brewer(palette="Dark2") + ggtitle("Managed Grassland - Rainfed")
# mg_irr_raw <-
#   aei_no_std %>% 
#   drop_na(Managed_Grasslands.1) %>% 
#   ggplot(aes(x=Managed_Grasslands.1, color=six_regions)) +
#   geom_histogram(fill="white", position="dodge", bins = 100) + 
#   theme(legend.position = "none") + xlim(-1,1) +
#   scale_color_brewer(palette="Dark2") + ggtitle("Managed Grassland - Irrigated")
# 
# mg_rain_std <-
#   aei_std %>% 
#   drop_na(Managed_Grasslands) %>% 
#   ggplot(aes(x=Managed_Grasslands, color=six_regions)) +
#   geom_histogram(fill="white", position="dodge", bins = 100) + 
#   theme(legend.position = "none") + xlim(-1,1) +
#   scale_color_brewer(palette="Dark2") + ggtitle("Managed Grassland - Rainfed - Centered")
# mg_irr_std <-
#   aei_std %>% 
#   drop_na(Managed_Grasslands.1) %>% 
#   ggplot(aes(x=Managed_Grasslands.1, color=six_regions)) +
#   geom_histogram(fill="white", position="dodge", bins = 100) + 
#   theme(legend.position = "none") + xlim(-1,1) +
#   scale_color_brewer(palette="Dark2") + ggtitle("Managed Grassland - Irrigated - Centered")
# 
# grid.arrange(mg_rain_raw,mg_irr_raw, mg_rain_std, mg_irr_std, nrow = 4)
```
