---
bibliography: "/Volumes/RachelExternal/Thesis/Thesis/references.bib"
---
# Data

Input data from Fabian and LPJML

```{r}

usethis::use_blank_slate()

iFol="/Volumes/RachelExternal/Thesis/Data_and_Plots/Grid_Data_Scripts/"
oFol="/Volumes/RachelExternal/Thesis/Data_and_Plots/Plots/"
load(file=paste0(iFol,"data.RData"))
load(file=paste0(iFol,"cftfrac.RData"))
load(file=paste0(iFol,"countryData_67420.RData"))
load(file=paste0(iFol,"pet_data.RData"))
load(file=paste0(iFol,"landfrac_67420.RData"))
source(file=paste0(iFol,"lpjmliotools_20201104.R"))

rm(transp_blue_landuse, transp_green_landuse, runoff_landuse, discharge_landuse, evap_landuse, interc_landuse, mean_increase)

```

These datasets contain the following data. Not all predictors will be used.

data.Rdata contains:

-   distNextIrr [1:67420, 1:105]: distance (in km) for each gridcell to next irrigated cell at that time

-   irrFrac: fraction (0..1) of cell that is irrigated lat/lon: latitude (-90..90)/longitude (-180..180) of each lpjml gridcell

-   lpjGDPpc: GDP per capita (2011 US dollars/cap) for each lpj-grid

-   mean_increase [1:67420, 1:105]: mean increase in productivity (delta gC/m2) in each gridcell across all crops

-   median_increase [1:67420, 1:105]: median increase in productivity (delta gC/m2) in each gridcell across all crops

-   popdens [1:67420, 1:111]: population density (cap/km2) starting 1901 to 2011

-   globalIrrigArea: Mha discharge_landuse [1:67420, 1:12, 1:105]: (hm)3/d == (1.000.000 m3)/d

-   prec_landuse [1:67420, 1:12, 1:105]: monthly precipitation (mm/month)

-   runoff_landuse [1:67420, 1:12, 1:105]: amount of water (fed by precipitation) that is not consumed and is handed over to the next gridcell in the riverrouting (mm/month)

-   transp_green_landuse [1:67420, 1:12, 1:105]: amount of green water (fed by precipitation) that is transpired by plants (mm/month)

-   transp_blue_landuse [1:67420, 1:12, 1:105]: amount of blue water (fed by irrigation) that is transpired by plants (mm/month)

-   evap_landuse [1:67420, 1:12, 1:105]: amount of water (fed by precipitation) that evaporates on bare soil (mm/month)

-   interc_landuse [1:67420, 1:12, 1:105]: amount of water (fed by precipitation) that evaporates from plants leaves (is intercepted -- never reaches the ground) (mm/month)

cftfrac.RData contains:

-   cftfrac [1:67420, 1:32, 1:105]: fraction of gridcell cultivated with the given cft (crop functional type); cft 1-16 are the rainfed fractions, 17-32 the irrigated ones - the sum of which you have already in irrFrac

1.  Temperate cereals (wheat, rye, barley; wheat)

2.  Rice (paddy rice; rice)

3.  Maize (maize for food; maize)

4.  Tropical cereals (millet, sorghum; millet)

5.  Pulses (pulses; field peas)

6.  Temperate roots (sugar beet; sugar beet)

7.  Tropical roots (cassava; cassava)

8.  Sunflower (sunflower; sunflower)

9.  Soybean (soybean; soybean)

10. Groundnuts (groundnuts; groundnuts)

11. Rapeseed (rapeseed; rapeseed)

12. Sugarcane (sugarcane: sugarcane)

13. others (potatoes, oil palm, citrus, date palm, grapes/vine, cotton, cocoa, coffee, other perennial crops, other annual crops; managed grassland)

14. managed grasslands (pastures; managed grasslands)

15. bio-energy grass

16. bio-energy tree

countryData_67420.RData contains:

-   fullCountryData - data.frame, which contains for every grid cell the LPJmL country id (COW), the LPJmL regional code (REG) - both defined in "include/managepar.h", the more of less official english Countryname, and the international 3 character ISO code

    example usage: `countryIrrfrac2005=aggregateLPJmLdata2Country(input=irrFrac[,105]*cellarea,cowList = fullCountryData ISO,aggMethod = "sum")`

    `plotCountryData(data = countryIrrfrac2005/10^6,sty="log",cowList = fullCountryData$ISO,file = paste0(oFol,"countryIrrFrac2005.png"),title = "Country specific total irrigated area",legendtitle = "km2")`

pet_data.RData contains

-   total_PET [1:67420, 1:12, 1:105]: Monthly Potential Evapotranspiration (mm/month) for each gridcell and year under historic landuse

-   total_PET_allcrops [1:67420, 1:12, 1:105]: Monthly Potential Evapotranspiration (mm/month) for each gridcell and year under synthetic allcrops conditions (not sure if this is at all helpful)

```{r packages}
library(tidyverse)
library(dplyr)
library(agrmt)
```


# Dataframe Assembly
```{r irrfrac}
aei <- as.data.frame(irrFrac)
rm(irrFrac)
colnames(aei) <- seq(1901, 2005, by = 1)
rownames(aei) <- seq(1, 67420, by =1)
aei$id <- seq(1, 67420, by =1)

meta <- as.data.frame(fullCountryData)
rownames(meta) <- seq(1, 67420, by =1)
meta$id <- seq(1, 67420, by =1)


aei <- meta %>% 
  select(Countryname, ISO, id) %>% 
  right_join(., aei, by = "id")
rm(meta)

```

Adding some extra meta data here for further analysis down the road.

```{r extrameta}
#some metadata for later analysis
extrameta <- read.csv2("/Volumes/RachelExternal/Thesis/Data_and_Plots/Country_Level_Data/GAPminder_ISOcodes.csv")

extrameta %>% 
  select(ISO, name) %>% 
  anti_join(aei, ., by = "ISO") %>% 
  select(ISO, Countryname) %>% 
  unique()
#ok 44 countries wont have this extra metadata. majority are island nations, thats ok. I also will take them out later when I do a cut based on number of cells per country.


aei <- 
  extrameta %>%
  select(-name, -Longitude, -Latitude) %>% 
    mutate_if(is.character,as.factor) %>% 
  right_join(., aei, by = "ISO")

rm(extrameta)
```

Add cellarea, landfrac, cropland, latitude and longitude to this dataframe

```{r landfrac}

landfr <- as.data.frame(landfrac)
#this is horozontal. rotate it. 
landfr <- pivot_longer(landfr, everything())
landfr$id <- seq(1, 67420, by =1)

aei <- 
  landfr %>%
  select(-name) %>% 
  rename(landfrac = value) %>% 
  left_join(., aei, by = "id")

```

```{r cellarea}

#fix cellarea before we start
cellarea <- (111194.9/2)*(111194.9/2)*cos(lat/180*pi) 
cellar <- as.data.frame(cellarea)
cellar$id <- seq(1, 67420, by =1)

aei <- 
  cellar %>%
  left_join(., aei, by = "id")

aei$landarea <- aei$landfrac*aei$cellarea #still in m2
```

```{r latlon}
#lat
la <- as.data.frame(lat)
la$id <- seq(1, 67420, by =1)

aei <- 
  la %>%
  left_join(., aei, by = "id")

#lon
lo <- as.data.frame(lon)
lo$id <- seq(1, 67420, by =1)

aei <- 
  lo %>%
  left_join(., aei, by = "id")
```

```{r cleanup 1}
#remove a bunch of junk
rm(lat, lon, la, lo, cellar, cellarea, landfr, landfrac)
```

Now pivot longer to add additional time series data

```{r dfpivot}

aei <- aei %>% 
  pivot_longer(`1901`:`2005`, names_to = "years", values_to = "irrfrac")
```

Add cropland, distance to next irrigated cell, population density, precip and PET. all need to be manipulated first and precip and PET need to be summed to yearly values.

```{r cropland}
cropland = apply(cftfrac[,1:32,],c(1,3),sum)
cropland <- as.data.frame(cropland)
colnames(cropland) <- seq(1901, 2005, by = 1)
cropland$id <- seq(1, 67420, by=1)

cropland <- cropland %>% 
  pivot_longer(-id, names_to = "years", values_to = "cropland")

aei <- aei %>% 
  right_join(., cropland, by = c("years", "id"))

rm(cropland)
```

```{r distnextirr}
dist <- as.data.frame(distNextIrr)
colnames(dist) <- seq(1901, 2005, by = 1)
dist$id <- seq(1, 67420, by =1)  

dist <- dist %>% 
  pivot_longer(-id, names_to = "years", values_to = "dist")

aei <- aei %>% 
  right_join(., dist, by = c("years", "id"))

rm(dist, distNextIrr)
```

```{r median_increase}
medinc <- as.data.frame(median_increase)
colnames(medinc) <- seq(1901, 2005, by = 1)
medinc$id <- seq(1, 67420, by =1)  

medinc <- medinc %>% 
  pivot_longer(-id, names_to = "years", values_to = "medinc")

aei <- aei %>% 
  right_join(., medinc, by = c("years", "id"))

rm(medinc, median_increase)
```


```{r popdens}
pop <- as.data.frame(popdens)
colnames(pop) <- seq(1901, 2011, by = 1)
pop$id <- seq(1, 67420, by =1)  

pop <- pop %>% 
  select(-c(`2006`:`2011`)) %>% 
  pivot_longer(-id, names_to = "years", values_to = "popdens")

aei <- aei %>% 
  right_join(., pop, by = c("years", "id"))

rm(pop, popdens)
```

```{r gdppc}
gdppc <- as.data.frame(lpjGDPpc)
colnames(gdppc) <- seq(1901, 2005, by = 1)
gdppc$id <- seq(1, 67420, by =1)  

gdppc <- gdppc %>% 
  pivot_longer(-id, names_to = "years", values_to = "gdppc")

aei <- aei %>% 
  right_join(., gdppc, by = c("years", "id"))

rm(gdppc, lpjGDPpc)
```

```{r precip}
prec = apply(prec_landuse[,1:12,],c(1,3),sum) #sums values accross the monthly dimension
prec <- as.data.frame(prec)
colnames(prec) <- seq(1901, 2005, by = 1)
prec$id <- seq(1, 67420, by =1)  

prec <- prec %>% 
  pivot_longer(-id, names_to = "years", values_to = "precip")

aei <- aei %>% 
  right_join(., prec, by = c("years", "id"))

rm(prec, prec_landuse)

```

```{r pet}
pet = apply(total_PET[,1:12,],c(1,3),sum) #sums values accross the monthly dimension
pet <- as.data.frame(pet)
colnames(pet) <- seq(1901, 2005, by = 1)
pet$id <- seq(1, 67420, by =1)  

pet <- pet %>% 
  pivot_longer(-id, names_to = "years", values_to = "pet")

aei <- aei %>% 
  right_join(., pet, by = c("years", "id"))

rm(pet, total_PET)
```

Turning year back numeric and adding a year count col. Removing data before 1960 as well. Limited data for several predictors exists prior to this.

```{r}
aei$years <- as.numeric(aei$years)
aei$yearcount <- aei$years - 1960
aei <- aei %>% 
  subset(years >= 1960)
```

Now add additional predictors such as

-   income,

-   percentage of gdp that comes from agriculture,

-   ruggedness

-   others

```{r ruggedness}
library(rethinking)
data("rugged")
aei <- rugged %>% select(c("isocode", "rugged")) %>% 
  rename(ISO = isocode) %>% 
  left_join(aei, ., by = c("ISO"))

rm(rugged)
```

```{r cerealyields}
library(wbstats)
aei <- wb_data("AG.YLD.CREL.KG", start_date = 1960, end_date = 2005) %>% 
  select(iso3c, date, AG.YLD.CREL.KG) %>% 
  rename(ISO = iso3c, years = date, crl_yld = AG.YLD.CREL.KG) %>% 
  select(ISO, years, crl_yld) %>% 
  left_join(aei, ., by = c("ISO", "years")) 
```

```{r perc_ag_gdp}
aei <- wb_data("NV.AGR.TOTL.ZS", start_date = 1960, end_date = 2005) %>% 
  select(iso3c, date, NV.AGR.TOTL.ZS) %>% 
  rename(ISO = iso3c, years = date, ag_gdp = NV.AGR.TOTL.ZS) %>% 
  select(ISO, years, ag_gdp) %>% 
  left_join(aei, ., by = c("ISO", "years")) 
```

Democracy

```{r democracy}
aei <- read.csv2("/Volumes/RachelExternal/Thesis/Data_and_Plots/Country_Level_Data/Bjørnskov-Rode-Democracy-data.csv") %>% 
  select(-c(6:12,14:19,21:22,24:51)) %>% 
  rename(ISO = country.isocode, years = year) %>% 
  select(ISO, years, DD.regime, DD.category, Democracy) %>% 
  mutate(DD.regime = replace_na(DD.regime, 6)) %>% 
  left_join(aei, ., by = c("ISO", "years")) 
  
 
```


#Variable transformation

- Irrigation fraction as a proportion of cropland, not cell area
- Humidity, precipitation divided by PET. Both are in the same units of mm/yr

As many cell value for cropland are 0, and dividing by zero produced NaNs, all NaN values in irrcrop will be replaced with 0s. 

```{r variable transformations}

aei$irrcrop <- aei$irrfrac/aei$cropland
aei$humid <- aei$precip/aei$pet

aei$irrcrop[is.nan(aei$irrcrop)]<-0
```



```{r}
summary(aei)

write_csv(aei, "/Volumes/RachelExternal/Thesis/Data_and_Plots/aei_full.csv")
```


# Removal of countries with < 5 cells

Finally, remove countries that have 5 or less cells.

```{r islandtchau!}
counts <- 
  fullCountryData %>% 
  group_by(ISO) %>% 
  count() %>% 
  subset(n <= 5)  %>% 
  select(ISO) 

counts <- as.character(counts$ISO)

#for fun
`%notin%` <- Negate(`%in%`)

aei_clean <- subset(aei, ISO %notin% counts)

summary(aei_clean)

write_csv(aei_clean, "/Volumes/RachelExternal/Thesis/Data_and_Plots/aei_clean.csv")
```

# Removing cells with 0 cropland

```{r}

nocrop <- aei_clean %>% dplyr::group_by(id) %>% filter(cropland > 0) #check for no cropland
zerocrop <- aei_clean %>% dplyr::group_by(id) %>% filter(cropland == 0) #looks great


#extract cell ids
ids <- unique(nocrop$id)


aei_nocrop <- subset(aei_clean, subset = id %in% ids)




p1 <- ggplot(aei_nocrop, aes(x=irrfrac)) + 
  geom_histogram(bins = 100) + ggtitle("irrfrac w/ cells removed")

p2 <- ggplot(aei_clean, aes(x=irrfrac)) + 
  geom_histogram(bins = 100) + ggtitle("irrfrac w/ cells included")

gridExtra::grid.arrange(p1, p2, ncol = 2)



write_csv(aei_nocrop, "/Volumes/RachelExternal/Thesis/Data_and_Plots/aei_nocrop.csv")
```

```{r}
fullids <- aei_clean %>% dplyr::select(id, lat, lon) %>%  unique()
nocropids <- aei_nocrop %>% dplyr::select(id, lat, lon) %>%  unique()

library(ggmap)
library(raster)
library(ggplot2)

world <- map_data("world") # we already did this, but we can do it again

ggplot() + geom_polygon(data = world, aes(x=long, y = lat, group = group), fill = NA, color = "black") + 
  geom_point(data = fullids, aes(x=lon, y=lat), size=1,shape=".", color = "red") + 
  geom_point(data = nocropids, aes(x=lon, y=lat), size=1,shape=".", color = "blue")  

ggsave(file = "/Volumes/RachelExternal/Thesis/Data_and_Plots/nocropland.png", dpi = 500, width = 12, height = 8)
```



#Sampling grid with min distance

```{r}
library(sp)
library(spatialEco)

coordinates(nocropids) <- ~ lon+lat

# Subsample with a 2 degree (or 4 cell) minimum sample spread 
# length(nocropids) = 43696 * 0.05 = 2185 sample size
sub.nocropids <- subsample.distance(nocropids, size = 2185, d = 2, echo = F)  

library(ggmap)
library(raster)
library(ggplot2)

world <- map_data("world") # we already did this, but we can do it again

ggplot() + geom_polygon(data = world, aes(x=long, y = lat, group = group), fill = NA, color = "black") + 
  geom_point(data = as.data.frame(nocropids), aes(x=lon, y=lat), size=3,shape=".", color = "red") + 
  geom_point(data = as.data.frame(sub.nocropids), aes(x=lon, y=lat), size=3,shape=".", color = "blue") 


# Check distances	
dm <- spDists(sub.nocropids)
diag(dm) <- NA
cat("\n", "Min distance for subsample", min(dm, na.rm=TRUE), "\n")  
cat("Max distance for subsample", max(dm, na.rm=TRUE), "\n") 
```
Extract cell ids
```{r}
finalids <- sub.nocropids$id

aei_nocrop_samp <- subset(aei_nocrop, subset = id %in% finalids)

write.csv(aei_nocrop_samp, "/Volumes/RachelExternal/Thesis/Data_and_Plots/aei_nocropland_sample.csv")

summary(aei_nocrop_samp)
```



# Slimming to 5 year time steps
No need for this if we have decreased the 

```{r}
#aei_small <- aei_nocrop %>% subset(., subset = yearcount %in% c("0",  "5", "10", "15", "20", "25", "30", "35", "40", "45"))

#write.csv(aei_small, "/Volumes/RachelExternal/Thesis/Data_and_Plots/aei_small.csv")

#summary(aei_small)
```
  
# Histograms


```{r histdist}
hist(aei_nocrop_samp$irrfrac, breaks = 50)
```


```{r histdist}
hist(aei_nocrop_samp$dist, breaks = 1000)
plot(aei_nocrop_samp$dist, aei_nocrop_samp$irrfrac)
```

```{r histmedinc}
hist(aei_nocrop_samp$medinc, breaks = 100000, xlim = c(-1, 1))
plot(aei_nocrop_samp$medinc, aei_nocrop_samp$irrfrac)
```

```{r histpopdens}
hist(aei_nocrop_samp$popdens, breaks = 100)
plot(aei_nocrop_samp$popdens, aei_nocrop_samp$irrfrac)
```

```{r histgdp}
hist(aei_nocrop_samp$gdppc, breaks = 100)
plot(aei_nocrop_samp$gdppc, aei_nocrop_samp$irrfrac)
```

```{r histprecip}
hist(aei_nocrop_samp$precip, breaks = 100)
plot(aei_nocrop_samp$precip, aei_nocrop_samp$irrfrac)
```

```{r histpet}
hist(aei_nocrop_samp$pet, breaks = 100)
plot(aei_nocrop_samp$pet, aei_nocrop_samp$irrfrac)
```

```{r}
hist(aei_nocrop_samp$humid, breaks = 100)
plot(aei_nocrop_samp$humid, aei_nocrop_samp$irrfrac)
```


```{r histrugg}
hist(aei_nocrop_samp$rugged, breaks = 100)
plot(aei_nocrop_samp$rugged, aei_nocrop_samp$irrfrac)
```



# `pairs()`



```{r}
pairs(~irrfrac+log(gdppc)+log(popdens)+rugged,data=aei_nocrop_samp) 
```

```{r}
pairs(~irrfrac + precip + humid + pet,data=aei_nocrop_samp) 
```

I will be using precip. It fits better conceptually and the relationship here looks stronger than pet. 

```{r}
pairs(~irrfrac + dist + medinc + Democracy + DD.regime,data=aei_nocrop_samp) 
```


