---
bibliography: "/Volumes/RachelExternal/Thesis/Thesis/references.bib"
---
```{r packages}
library(tidyverse)
library(dplyr)
library(agrmt)
library(devtools)
#install_github('mcooper/moranfast')
library(moranfast)
library(Rcpp)
library(sp)
library(geoR)
library(gstat)
library(agrmt)
```

# Data

Input data from Fabian and LPJML

```{r}

usethis::use_blank_slate()

iFol="/Volumes/RachelExternal2/Thesis/Data_and_Plots/Grid_Data_Scripts/"
oFol="/Volumes/RachelExternal2/Thesis/Data_and_Plots/Plots/"
load(file=paste0(iFol,"data.RData"))
load(file=paste0(iFol,"cftfrac.RData"))
load(file=paste0(iFol,"countryData_67420.RData"))
load(file=paste0(iFol,"pet_data.RData"))
load(file=paste0(iFol,"landfrac_67420.RData"))
source(file=paste0(iFol,"lpjmliotools_20201104.R"))

rm(transp_blue_landuse, transp_green_landuse, runoff_landuse, discharge_landuse, evap_landuse, interc_landuse, mean_increase)

```

These datasets contain the following data. Not all predictors will be used.

data.Rdata contains:

-   distNextIrr [1:67420, 1:105]: distance (in km) for each gridcell to next irrigated cell at that time

-   irrFrac: fraction (0..1) of cell that is irrigated lat/lon: latitude (-90..90)/longitude (-180..180) of each lpjml gridcell

-   lpjGDPpc: GDP per capita (2011 US dollars/cap) for each lpj-grid

-   mean_increase [1:67420, 1:105]: mean increase in productivity (delta gC/m2) in each gridcell across all crops

-   median_increase [1:67420, 1:105]: median increase in productivity (delta gC/m2) in each gridcell across all crops

-   popdens [1:67420, 1:111]: population density (cap/km2) starting 1901 to 2011

-   globalIrrigArea: Mha discharge_landuse [1:67420, 1:12, 1:105]: (hm)3/d == (1.000.000 m3)/d

-   prec_landuse [1:67420, 1:12, 1:105]: monthly precipitation (mm/month)

-   runoff_landuse [1:67420, 1:12, 1:105]: amount of water (fed by precipitation) that is not consumed and is handed over to the next gridcell in the riverrouting (mm/month)

-   transp_green_landuse [1:67420, 1:12, 1:105]: amount of green water (fed by precipitation) that is transpired by plants (mm/month)

-   transp_blue_landuse [1:67420, 1:12, 1:105]: amount of blue water (fed by irrigation) that is transpired by plants (mm/month)

-   evap_landuse [1:67420, 1:12, 1:105]: amount of water (fed by precipitation) that evaporates on bare soil (mm/month)

-   interc_landuse [1:67420, 1:12, 1:105]: amount of water (fed by precipitation) that evaporates from plants leaves (is intercepted -- never reaches the ground) (mm/month)

cftfrac.RData contains:

-   cftfrac [1:67420, 1:32, 1:105]: fraction of gridcell cultivated with the given cft (crop functional type); cft 1-16 are the rainfed fractions, 17-32 the irrigated ones - the sum of which you have already in irrFrac

1.  Temperate cereals (wheat, rye, barley; wheat)

2.  Rice (paddy rice; rice)

3.  Maize (maize for food; maize)

4.  Tropical cereals (millet, sorghum; millet)

5.  Pulses (pulses; field peas)

6.  Temperate roots (sugar beet; sugar beet)

7.  Tropical roots (cassava; cassava)

8.  Sunflower (sunflower; sunflower)

9.  Soybean (soybean; soybean)

10. Groundnuts (groundnuts; groundnuts)

11. Rapeseed (rapeseed; rapeseed)

12. Sugarcane (sugarcane: sugarcane)

13. others (potatoes, oil palm, citrus, date palm, grapes/vine, cotton, cocoa, coffee, other perennial crops, other annual crops; managed grassland)

14. managed grasslands (pastures; managed grasslands)

15. bio-energy grass

16. bio-energy tree

countryData_67420.RData contains:

-   fullCountryData - data.frame, which contains for every grid cell the LPJmL country id (COW), the LPJmL regional code (REG) - both defined in "include/managepar.h", the more of less official english Countryname, and the international 3 character ISO code

    example usage: `countryIrrfrac2005=aggregateLPJmLdata2Country(input=irrFrac[,105]*cellarea,cowList = fullCountryData ISO,aggMethod = "sum")`

    `plotCountryData(data = countryIrrfrac2005/10^6,sty="log",cowList = fullCountryData$ISO,file = paste0(oFol,"countryIrrFrac2005.png"),title = "Country specific total irrigated area",legendtitle = "km2")`

pet_data.RData contains

-   total_PET [1:67420, 1:12, 1:105]: Monthly Potential Evapotranspiration (mm/month) for each gridcell and year under historic landuse

-   total_PET_allcrops [1:67420, 1:12, 1:105]: Monthly Potential Evapotranspiration (mm/month) for each gridcell and year under synthetic allcrops conditions (not sure if this is at all helpful)




# Dataframe Assembly
```{r irrfrac}
aei <- as.data.frame(irrFrac)
rm(irrFrac)
colnames(aei) <- seq(1901, 2005, by = 1)
rownames(aei) <- seq(1, 67420, by =1)
aei$id <- seq(1, 67420, by =1)

meta <- as.data.frame(fullCountryData)
rownames(meta) <- seq(1, 67420, by =1)
meta$id <- seq(1, 67420, by =1)


aei <- meta %>% 
  dplyr::select(Countryname, ISO, id) %>% 
  right_join(., aei, by = "id")
rm(meta)

```

Adding some extra meta data here for further analysis down the road.

```{r extrameta}
#some metadata for later analysis
extrameta <- read.csv2("/Volumes/RachelExternal2/Thesis/Data_and_Plots/Country_Level_Data/GAPminder_ISOcodes.csv")

extrameta %>% 
  dplyr::select(ISO, name) %>% 
  anti_join(aei, ., by = "ISO") %>% 
  dplyr::select(ISO, Countryname) %>% 
  unique()
#ok 44 countries wont have this extra metadata. majority are island nations, thats ok. I also will take them out later when I do a cut based on number of cells per country.


aei <- 
  extrameta %>%
  dplyr::select(-name, -Longitude, -Latitude) %>% 
    mutate_if(is.character,as.factor) %>% 
  right_join(., aei, by = "ISO")

rm(extrameta)
```

Add cellarea, landfrac, cropland, latitude and longitude to this dataframe

```{r landfrac}

landfr <- as.data.frame(landfrac)
#this is horozontal. rotate it. 
landfr <- pivot_longer(landfr, everything())
landfr$id <- seq(1, 67420, by =1)

aei <- 
  landfr %>%
  dplyr::select(-name) %>% 
  rename(landfrac = value) %>% 
  left_join(., aei, by = "id")

```

```{r cellarea}

#fix cellarea before we start
cellarea <- (111194.9/2)*(111194.9/2)*cos(lat/180*pi) 
cellar <- as.data.frame(cellarea)
cellar$id <- seq(1, 67420, by =1)

aei <- 
  cellar %>%
  left_join(., aei, by = "id")

aei$landarea <- aei$landfrac*aei$cellarea #still in m2
```

```{r latlon}
#lat
la <- as.data.frame(lat)
la$id <- seq(1, 67420, by =1)

aei <- 
  la %>%
  left_join(., aei, by = "id")

#lon
lo <- as.data.frame(lon)
lo$id <- seq(1, 67420, by =1)

aei <- 
  lo %>%
  left_join(., aei, by = "id")
```

```{r cleanup 1}
#remove a bunch of junk
rm(lat, lon, la, lo, cellar, cellarea, landfr, landfrac)
```

Now pivot longer to add additional time series data

```{r dfpivot}

aei <- aei %>% 
  pivot_longer(`1901`:`2005`, names_to = "years", values_to = "irrfrac")
```

Add cropland, distance to next irrigated cell, population density, precip and PET. all need to be manipulated first and precip and PET need to be summed to yearly values.

```{r cropland}
cropland = apply(cftfrac[,1:32,],c(1,3),sum)
cropland <- as.data.frame(cropland)
colnames(cropland) <- seq(1901, 2005, by = 1)
cropland$id <- seq(1, 67420, by=1)

cropland <- cropland %>% 
  pivot_longer(-id, names_to = "years", values_to = "cropland")

aei <- aei %>% 
  right_join(., cropland, by = c("years", "id"))

rm(cropland)
```


```{r distnextirr}
dist <- as.data.frame(distNextIrr)
colnames(dist) <- seq(1901, 2005, by = 1)
dist$id <- seq(1, 67420, by =1)  

dist <- dist %>% 
  pivot_longer(-id, names_to = "years", values_to = "dist")

aei <- aei %>% 
  right_join(., dist, by = c("years", "id"))

rm(dist, distNextIrr)
```

```{r median_increase}
medinc <- as.data.frame(median_increase)
colnames(medinc) <- seq(1901, 2005, by = 1)
medinc$id <- seq(1, 67420, by =1)  

medinc <- medinc %>% 
  pivot_longer(-id, names_to = "years", values_to = "medinc")

aei <- aei %>% 
  right_join(., medinc, by = c("years", "id"))

rm(medinc, median_increase)
```


```{r popdens}
pop <- as.data.frame(popdens)
colnames(pop) <- seq(1901, 2011, by = 1)
pop$id <- seq(1, 67420, by =1)  

pop <- pop %>% 
  dplyr::select(-c(`2006`:`2011`)) %>% 
  pivot_longer(-id, names_to = "years", values_to = "popdens")

aei <- aei %>% 
  right_join(., pop, by = c("years", "id"))

rm(pop, popdens)
```

```{r gdppc}
gdppc <- as.data.frame(lpjGDPpc)
colnames(gdppc) <- seq(1901, 2005, by = 1)
gdppc$id <- seq(1, 67420, by =1)  

gdppc <- gdppc %>% 
  pivot_longer(-id, names_to = "years", values_to = "gdppc")

aei <- aei %>% 
  right_join(., gdppc, by = c("years", "id"))

rm(gdppc, lpjGDPpc)
```

```{r precip}
prec = apply(prec_landuse[,1:12,],c(1,3),sum) #sums values accross the monthly dimension
prec <- as.data.frame(prec)
colnames(prec) <- seq(1901, 2005, by = 1)
prec$id <- seq(1, 67420, by =1)  

prec <- prec %>% 
  pivot_longer(-id, names_to = "years", values_to = "precip")

aei <- aei %>% 
  right_join(., prec, by = c("years", "id"))

rm(prec, prec_landuse)

```

```{r pet}
pet = apply(total_PET[,1:12,],c(1,3),sum) #sums values accross the monthly dimension
pet <- as.data.frame(pet)
colnames(pet) <- seq(1901, 2005, by = 1)
pet$id <- seq(1, 67420, by =1)  

pet <- pet %>% 
  pivot_longer(-id, names_to = "years", values_to = "pet")

aei <- aei %>% 
  right_join(., pet, by = c("years", "id"))

rm(pet, total_PET)
```

Turning year back numeric and adding a year count col. Removing data before 1960 as well. Limited data for several predictors exists prior to this.

```{r}
aei$years <- as.numeric(aei$years)
aei$yearcount <- aei$years - 1960
aei <- aei %>% 
  subset(years >= 1960)
```

Now add additional predictors such as

-   income,

-   percentage of gdp that comes from agriculture,

-   ruggedness

-   others

```{r ruggedness}
library(rethinking)
data("rugged")
aei <- rugged %>% dplyr::select(c("isocode", "rugged")) %>% 
  rename(ISO = isocode) %>% 
  left_join(aei, ., by = c("ISO"))

rm(rugged)
```

```{r cerealyields}
library(wbstats)
aei <- wb_data("AG.YLD.CREL.KG", start_date = 1960, end_date = 2005) %>% 
  dplyr::select(iso3c, date, AG.YLD.CREL.KG) %>% 
  rename(ISO = iso3c, years = date, crl_yld = AG.YLD.CREL.KG) %>% 
  dplyr::select(ISO, years, crl_yld) %>% 
  left_join(aei, ., by = c("ISO", "years")) 
```

```{r perc_ag_gdp}
aei <- wb_data("NV.AGR.TOTL.ZS", start_date = 1960, end_date = 2005) %>% 
  dplyr::select(iso3c, date, NV.AGR.TOTL.ZS) %>% 
  rename(ISO = iso3c, years = date, ag_gdp = NV.AGR.TOTL.ZS) %>% 
  dplyr::select(ISO, years, ag_gdp) %>% 
  left_join(aei, ., by = c("ISO", "years")) 
```

Democracy

```{r democracy}
aei <- read.csv2("/Volumes/RachelExternal2/Thesis/Data_and_Plots/Country_Level_Data/Bjørnskov-Rode-Democracy-data.csv") %>% 
  dplyr::select(-c(6:12,14:19,21:22,24:51)) %>% 
  rename(ISO = country.isocode, years = year) %>% 
  dplyr::select(ISO, years, DD.regime, DD.category, Democracy) %>% 
  mutate(DD.regime = replace_na(DD.regime, 6)) %>% 
  mutate(ISO=replace(ISO, ISO=="GER", "DEU"),
         ISO=replace(ISO, ISO=="ROM", "ROU"), 
         ISO=replace(ISO, ISO=="ZAR", "COD")) %>% 
  left_join(aei, ., by = c("ISO", "years")) 
  
 
```

```{r rice_and_maize}

FAOSTAT_data_11_14_2021_rice_and_maize <- read_csv("/Volumes/RachelExternal2/Thesis/Data_and_Plots/Country_Level_Data/FAOSTAT_data_11-14-2021_rice_and_maize.csv")
FAOSTAT_data_11_15_2021_area <- read_csv("/Volumes/RachelExternal2/Thesis/Data_and_Plots/Country_Level_Data/FAOSTAT_data_11-15-2021_area.csv")
FAOSTAT_data_11_15_2021 <- read_csv("/Volumes/RachelExternal2/Thesis/Data_and_Plots/Country_Level_Data/FAOSTAT_data_11-15-2021.csv")

FAOSTAT_data_11_15_2021_area <- 
  FAOSTAT_data_11_15_2021_area %>% 
  dplyr::select(`Area Code (ISO3)`, Area, Value, Unit) %>% 
  dplyr::rename(ISO = `Area Code (ISO3)`)

FAOSTAT_data_11_15_2021 <- 
  FAOSTAT_data_11_15_2021 %>% 
  rename(ISO = `Area Code (ISO3)`) %>% 
  left_join(., FAOSTAT_data_11_15_2021_area, by = "ISO") %>% 
  dplyr::rename(cereals_ha = "Value.x", 
         total_area = "Value.y", 
         Countryname = "Area.x",
         years = "Year") %>% 
  dplyr::mutate(total_area = total_area*1000) %>% 
  dplyr::select(ISO, Countryname,
                Item, years, cereals_ha, total_area ) %>%
  mutate(cereal_frac = cereals_ha/total_area) %>% 
  tibble()


new <- tibble(FAOSTAT_data_11_15_2021$ISO) %>% unique() %>% rename(ISO = "FAOSTAT_data_11_15_2021$ISO")
new$years <- 1960

FAOSTAT_data_11_15_2021 <- dplyr::bind_rows(new, FAOSTAT_data_11_15_2021)

#fill in that missing year
FAOSTAT_data_11_15_2021<-
  FAOSTAT_data_11_15_2021 %>% 
  dplyr::group_by(ISO) %>% 
  fill(everything(), .direction = "up") %>% 
  dplyr::ungroup() %>% 
  dplyr::select(ISO, years, cereals_ha, total_area, cereal_frac)


  
rm(FAOSTAT_data_11_15_2021_area)


#the join
#will come back later to fix some countries, 
#just want to see how things play out

aei <- aei %>% 
  anti_join(., FAOSTAT_data_11_15_2021, by = c("years", "ISO")) 
```
#Variable transformation

- Irrigation fraction as a proportion of cropland, not cell area
- Humidity, precipitation divided by PET. Both are in the same units of mm/yr

As many cell value for cropland are 0, and dividing by zero produced NaNs, all NaN values in irrcrop will be replaced with 0s. 

```{r variable transformations}

aei$irrcrop <- aei$irrfrac/aei$cropland
aei$humid <- aei$precip/aei$pet

aei$irrcrop[is.nan(aei$irrcrop)]<-0
```



```{r}
summary(aei)

write_csv(aei, "/Volumes/RachelExternal2/Thesis/Data_and_Plots/aei_full.csv")
```



# Removing cells with 0 cropland

```{r}

nocrop <- aei %>% dplyr::group_by(id) %>% filter(cropland > 0) #check for no cropland
zerocrop <- aei %>% dplyr::group_by(id) %>% filter(cropland == 0) #looks great


#extract cell ids
ids <- unique(nocrop$id)


aei_nocrop <- subset(aei, subset = id %in% ids)




p1 <- ggplot(aei_nocrop, aes(x=irrfrac)) + 
  geom_histogram(bins = 100) + ggtitle("irrfrac w/ cells removed")

p2 <- ggplot(aei, aes(x=irrfrac)) + 
  geom_histogram(bins = 100) + ggtitle("irrfrac w/ cells included")

gridExtra::grid.arrange(p1, p2, ncol = 2)



write.csv(aei_nocrop, "/Volumes/RachelExternal2/Thesis/Data_and_Plots/aei_nocrop.csv")
```



```{r}
fullids <- aei %>% dplyr::select(id, lat, lon) %>%  unique()
nocropids <- aei_nocrop %>% dplyr::select(id, lat, lon) %>%  unique()

library(ggmap)
library(raster)
library(ggplot2)

world <- map_data("world") 

ggplot() + geom_polygon(data = world, aes(x=long, y = lat, group = group), fill = NA, color = "black") + 
  geom_point(data = fullids, aes(x=lon, y=lat), size=1,shape=".", color = "red") + 
  geom_point(data = nocropids, aes(x=lon, y=lat), size=1,shape=".", color = "blue")  

ggsave(file = "/Volumes/RachelExternal/Thesis/Data_and_Plots/nocropland.png", dpi = 500, width = 12, height = 8)
```

# Moran's I

```{r morani}
india <- aei_nocrop %>% subset(yearcount == 45 & ISO == "IND") %>% dplyr::select(id, lat, lon, irrfrac) #india for the year 2005
moranfast(india$irrfrac, india$lon, india$lat)
```

```{r}
coordinates(india)= ~ lon+lat
bubble(india, zcol='irrfrac', fill=TRUE, do.sqrt=FALSE, maxsize=3) #check the bubble
# TheVariogram=variogram(irrfrac~1, data=india)
# plot(TheVariogram)
```




#NAs 

Check for NA data in main predictors
```{r}
aei_nocrop %>% subset(., is.na(DD.regime)) %>% dplyr::select(Countryname) %>% unique()
aei_nocrop %>% subset(., is.na(gdppc)) %>% dplyr::select(Countryname) %>% unique()
aei_nocrop %>% subset(., is.na(rugged)) %>% dplyr::select(Countryname) %>% unique()
aei_nocrop %>% subset(., is.na(cereal_frac)) %>% dplyr::select(Countryname) %>% unique()

```

There is a lot of NAs in a couple different predictors. Brms doesn't allow for NAs. I will remove them.

```{r}
aei_no_nas <- 
  aei_nocrop %>% 
  dplyr::select(id, ISO, Countryname, lon, lat, six_regions, eight_regions, years, yearcount, irrfrac, irrcrop, cropland, dist, medinc, precip, pet, humid, rugged, popdens, gdppc, DD.regime, DD.category) %>% 
  drop_na(rugged, gdppc, DD.regime)

nonas <- aei_no_nas %>% dplyr::select(id, lat, lon) %>%  unique()
```


#Sampling grid with min distance to reduce autocorrelation

```{r}
library(sp)
library(rgeos)
library(geosphere)
library(spatialEco)


coordinates(nonas) <- ~lon+lat
#specifying lon lat projection
proj4string(nonas) <- CRS("+proj=longlat +datum=WGS84") 
#converting to UTM (x, y) projection
res <- spTransform(nonas, CRS("+proj=utm +ellps=WGS84 +datum=WGS84 +units=km" )) 

#sampling with a min dist of 3 cells between @equator 0.5deg cell is 55km
#3 cells = 165 km of min dist.
sub.nocropids165a <- subsample.distance(res, size = 2185, d = 165, echo = T)  
sub.nocropids165b <- subsample.distance(res, size = 2185, d = 165, echo = T )  
sub.nocropids100a <- subsample.distance(res, size = 2185, d = 100, echo = T)  
sub.nocropids100b <- subsample.distance(res, size = 2185, d = 100, echo = T)  

library(ggmap)
library(raster)
library(ggplot2)

#flip back to lat lon just to plot
sub.nocropids165aa <- spTransform(sub.nocropids165a, CRS("+proj=longlat +ellps=WGS84 +datum=WGS84" ))
sub.nocropids165bb <- spTransform(sub.nocropids165b, CRS("+proj=longlat +ellps=WGS84 +datum=WGS84" ))
sub.nocropids100aa <- spTransform(sub.nocropids100a, CRS("+proj=longlat +ellps=WGS84 +datum=WGS84" ))
sub.nocropids100bb <- spTransform(sub.nocropids100b, CRS("+proj=longlat +ellps=WGS84 +datum=WGS84" ))

world <- map_data("world") # we already did this, but we can do it again

ggplot() + geom_polygon(data = world, aes(x=long, y = lat, group = group), fill = NA, color = "black") + 
  geom_point(data = as.data.frame(nonas), aes(x=lon, y=lat), size=3,shape=".", color = "red") + 
  geom_point(data = as.data.frame(sub.nocropids165aa), aes(x=lon, y=lat), size=3,shape=".", color = "blue") 

ggsave(plot = last_plot(), dpi = 300, width = 10, height = 6.6, filename = "/Volumes/RachelExternal/Thesis/Data_and_Plots/Plots/subnocropids165aa.png")

ggplot() + geom_polygon(data = world, aes(x=long, y = lat, group = group), fill = NA, color = "black") + 
  geom_point(data = as.data.frame(nonas), aes(x=lon, y=lat), size=3,shape=".", color = "red") + 
  geom_point(data = as.data.frame(sub.nocropids165bb), aes(x=lon, y=lat), size=3,shape=".", color = "blue") 

ggsave(plot = last_plot(), dpi = 300, width = 10, height = 6.6, filename = "/Volumes/RachelExternal/Thesis/Data_and_Plots/Plots/subnocropids165bb.png")

ggplot() + geom_polygon(data = world, aes(x=long, y = lat, group = group), fill = NA, color = "black") + 
  geom_point(data = as.data.frame(nonas), aes(x=lon, y=lat), size=3,shape=".", color = "red") + 
  geom_point(data = as.data.frame(sub.nocropids100aa), aes(x=lon, y=lat), size=3,shape=".", color = "blue") 

ggsave(plot = last_plot(), dpi = 300, width = 10, height = 6.6, filename = "/Volumes/RachelExternal/Thesis/Data_and_Plots/Plots/sub.nocropids100aa.png")

ggplot() + geom_polygon(data = world, aes(x=long, y = lat, group = group), fill = NA, color = "black") + 
  geom_point(data = as.data.frame(nonas), aes(x=lon, y=lat), size=3,shape=".", color = "red") + 
  geom_point(data = as.data.frame(sub.nocropids100bb), aes(x=lon, y=lat), size=3,shape=".", color = "blue") 

ggsave(plot = last_plot(), dpi = 300, width = 10, height = 6.6, filename = "/Volumes/RachelExternal/Thesis/Data_and_Plots/Plots/sub.nocropids100bb.png")

# Check distances	
dm <- spDists(sub.nocropids)
diag(dm) <- NA
cat("\n", "Min distance for subsample", min(dm, na.rm=TRUE), "\n")  
cat("Max distance for subsample", max(dm, na.rm=TRUE), "\n") 
```


Extract cell ids
```{r}
finalids165aa <- sub.nocropids165aa$id
finalids165bb <- sub.nocropids165bb$id
finalids100aa <- sub.nocropids100aa$id
finalids100bb <- sub.nocropids100bb$id

aei_nocrop_samp165aa <- subset(aei_no_nas, subset = id %in% finalids165aa)
aei_nocrop_samp165bb <- subset(aei_no_nas, subset = id %in% finalids165bb)
aei_nocrop_samp100aa <- subset(aei_no_nas, subset = id %in% finalids100aa)
aei_nocrop_samp100bb <- subset(aei_no_nas, subset = id %in% finalids100bb)




write.csv(aei_nocrop_samp165aa, "/Volumes/RachelExternal/Thesis/Data_and_Plots/aei_nocrop_samp165aa.csv")
write.csv(aei_nocrop_samp165bb, "/Volumes/RachelExternal/Thesis/Data_and_Plots/aei_nocrop_samp165bb.csv")
write.csv(aei_nocrop_samp100aa, "/Volumes/RachelExternal/Thesis/Data_and_Plots/aei_nocrop_samp100aa.csv")
write.csv(aei_nocrop_samp100bb, "/Volumes/RachelExternal/Thesis/Data_and_Plots/aei_nocrop_samp100bb.csv")



```



# Slimming to 5 year time steps
No need for this if we have decreased the 

```{r}
aei_nocrop_samp165aa_5yr <- aei_nocrop_samp165aa %>% subset(., subset = yearcount %in% c("0",  "5", "10", "15", "20", "25", "30", "35", "40", "45"))
aei_nocrop_samp165bb_5yr <- aei_nocrop_samp165bb %>% subset(., subset = yearcount %in% c("0",  "5", "10", "15", "20", "25", "30", "35", "40", "45"))
aei_nocrop_samp100aa_5yr <- aei_nocrop_samp100aa %>% subset(., subset = yearcount %in% c("0",  "5", "10", "15", "20", "25", "30", "35", "40", "45"))
aei_nocrop_samp100bb_5yr <- aei_nocrop_samp100bb %>% subset(., subset = yearcount %in% c("0",  "5", "10", "15", "20", "25", "30", "35", "40", "45"))

write.csv(aei_nocrop_samp165aa_5yr, "/Volumes/RachelExternal/Thesis/Data_and_Plots/aei_nocrop_samp165aa_5yr.csv")
write.csv(aei_nocrop_samp165bb_5yr, "/Volumes/RachelExternal/Thesis/Data_and_Plots/aei_nocrop_samp165bb_5yr.csv")
write.csv(aei_nocrop_samp100aa_5yr, "/Volumes/RachelExternal/Thesis/Data_and_Plots/aei_nocrop_samp100aa_5yr.csv")
write.csv(aei_nocrop_samp100bb_5yr, "/Volumes/RachelExternal/Thesis/Data_and_Plots/aei_nocrop_samp100bb_5yr.csv")

summary(aei_nocrop_samp_5yr)
```
  
# Histograms


```{r histdist}
hist(aei_nocrop_samp_5yr$irrfrac, breaks = 50)
```


```{r histdist}
hist(aei_nocrop_samp_5yr$dist, breaks = 1000)
plot(aei_nocrop_samp_5yr$dist, aei_nocrop_samp_5yr$irrfrac)
```

```{r histmedinc}
hist(aei_nocrop_samp_5yr$medinc, breaks = 100000, xlim = c(-1, 1))
plot(aei_nocrop_samp_5yr$medinc, aei_nocrop_samp_5yr$irrfrac)
```

```{r histpopdens}
hist(aei_nocrop_samp_5yr$popdens, breaks = 100)
plot(aei_nocrop_samp_5yr$popdens, aei_nocrop_samp_5yr$irrfrac)
```

```{r histgdp}
hist(aei_nocrop_samp_5yr$gdppc, breaks = 100)
plot(aei_nocrop_samp_5yr$gdppc, aei_nocrop_samp_5yr$irrfrac)
```

```{r histprecip}
hist(aei_nocrop_samp_5yr$precip, breaks = 100)
plot(aei_nocrop_samp_5yr$precip, aei_nocrop_samp_5yr$irrfrac)
```

```{r histpet}
hist(aei_nocrop_samp_5yr$pet, breaks = 100)
plot(aei_nocrop_samp_5yr$pet, aei_nocrop_samp_5yr$irrfrac)
```

```{r}
hist(aei_nocrop_samp_5yr$humid, breaks = 100)
plot(aei_nocrop_samp_5yr$humid, aei_nocrop_samp_5yr$irrfrac)
```


```{r histrugg}
hist(aei_nocrop_samp_5yr$rugged, breaks = 100)
plot(aei_nocrop_samp_5yr$rugged, aei_nocrop_samp_5yr$irrfrac)
```



# `pairs()`



```{r}
pairs(~irrfrac+log(gdppc)+log(popdens)+rugged,data=aei_nocrop_samp_5yr) 
```

```{r}
pairs(~irrfrac + precip + humid + pet,data=aei_nocrop_samp_5yr) 
```

I will be using precip. It fits better conceptually and the relationship here looks stronger than pet. 

```{r}
pairs(~irrfrac + dist + medinc + Democracy + DD.regime,data=aei_nocrop_samp_5yr) 
```


