```{r}
usethis::use_blank_slate()

iFol="/Volumes/RachelExternal/Thesis/Data_and_Plots/Grid_Data_Scripts/"
oFol="/Volumes/RachelExternal/Thesis/Data_and_Plots/Plots/"
load(file=paste0(iFol,"data.RData"))
load(file=paste0(iFol,"cftfrac.RData"))
load(file=paste0(iFol,"countryData_67420.RData"))
load(file=paste0(iFol,"pet_data.RData"))
load(file=paste0(iFol,"landfrac_67420.RData"))
source(file=paste0(iFol,"lpjmliotools_20201104.R"))

#remove some stuff we wont use
rm(mean_increase, discharge_landuse, evap_landuse, interc_landuse, runoff_landuse, total_PET_allcrops, transp_blue_landuse, transp_green_landuse)

```

```{r}
library(dplyr)
library(tidyverse)
```

```{r}
#fix cellarea before we start
cellarea <- (111194.9/2)*(111194.9/2)*cos(lat/180*pi) 
landarea <- landfrac*cellarea #still in m2

yearcount <- seq(1, 105, by = 1 )

```

## Total Country Area and Metadata

```{r}

cl_ha <- NULL
cl_ha <- aggregateLPJmLdata2Country(input=landarea/10000 , #ha conversion
                                    cowList = fullCountryData$ISO,
                                    aggMethod = "sum")


cl_ha <- as.data.frame(cl_ha)
colnames(cl_ha) <- "country_area_ha"

aei_country <- as.data.frame(cl_ha)
aei_country$ISO <- row.names(aei_country)
aei_country$ISO <- as.factor(aei_country$ISO)


#add countrynames
aei_country <- fullCountryData %>% 
  select(ISO, Countryname) %>% 
  unique() %>% 
  right_join(., aei_country, by = "ISO")

#some metadata for later analysis
meta <- read.csv2("/Volumes/RachelExternal/Thesis/Data_and_Plots/Country_Level_Data/GAPminder_ISOcodes.csv")


#fix the factors
meta <- 
  meta%>%
  select(-name) %>% 
    mutate_if(is.character,as.factor)


aei_country <- left_join(
  aei_country, 
  meta, 
  by = "ISO"
)

rm(cl_ha, meta)
```

## Cropland Area

Goal: total cropland area for each country

```{r}
Cropland = apply(cftfrac[,1:32,],c(1,3),sum) #summing all the crop fractions per grid cell
cropland_cl_ha <- NULL
for (i in yearcount){
  new_col <- aggregateLPJmLdata2Country(input=Cropland[,i]*cellarea, #multiplying by cell area
                                        cowList = fullCountryData$ISO,
                                        aggMethod = "sum")
  new_col <- new_col/10000 #m2 to ha conversion
  cropland_cl_ha <- cbind(cropland_cl_ha, new_col)
}
colnames(cropland_cl_ha) <- yearcount
cropland_cl_ha <- as.data.frame(cropland_cl_ha)
cropland_cl_ha$ISO <- row.names(cropland_cl_ha)
cropland_cl_ha$ISO <- as.factor(cropland_cl_ha$ISO)

rm(Cropland)

#pivot for tidy data
cropland_cl_ha <- pivot_longer(
  cropland_cl_ha, 
  -ISO,
  names_to = "yearcount", 
  values_to = "cropland_ha"
)

#merge to country area data
aei_country <- left_join(
  aei_country, 
  cropland_cl_ha, 
  by = "ISO"
)

aei_country$yearcount <- as.numeric(aei_country$yearcount)

rm(cropland_cl_ha)
```

## Irrigation Fraction

Irrigation Fraction for each country.

```{r}
irr_cl_ha <- NULL
for (i in yearcount){
  new_col <- aggregateLPJmLdata2Country(input=irrFrac[,i]*cellarea,
                                        cowList = fullCountryData$ISO,
                                        aggMethod = "sum")
  new_col <- new_col/10000 #m2 to ha conversion
  irr_cl_ha <- cbind(irr_cl_ha, new_col)
}
colnames(irr_cl_ha) <- yearcount


irr_cl_ha <- as.data.frame(irr_cl_ha)
irr_cl_ha$ISO <- row.names(irr_cl_ha)
irr_cl_ha$ISO <- as.factor(irr_cl_ha$ISO)

#pivot for tidy data
irr_cl_ha <- pivot_longer(
  irr_cl_ha, 
  -ISO,
  names_to = "yearcount", 
  values_to = "irrigated_ha"
)

irr_cl_ha$yearcount <- as.numeric(irr_cl_ha$yearcount )

#merge to big dataframe
aei_country <- left_join(
  aei_country, 
  irr_cl_ha, 
  by = c("ISO", "yearcount")
)

rm(irr_cl_ha)
```

## Creation of Irrigation Fraction on with two metrics

```{r}
aei_country <- aei_country %>% 
  mutate(
  if_total_area = irrigated_ha/country_area_ha, 
  if_crop_area = irrigated_ha/cropland_ha
)

#fix the NaNs from dividing by 0 cropland area
aei_country$if_crop_area[is.nan(aei_country$if_crop_area)]<-0
```

Removing countries with less than 5 cells, at the equator. These consist of mainly islands.

```{r}

counts <- 
  fullCountryData %>% 
  group_by(ISO, Countryname) %>% 
  count() %>% 
  subset(n <= 5) 

aei_country <- aei_country %>% 
  anti_join(., counts, by = "ISO")

rm(counts)
```

## Population

Using popdens

```{r}
poptot <- NULL
popdens <- popdens[,-c(106:111)]
for (i in yearcount){
  new_col <- aggregateLPJmLdata2Country(input=popdens[,i]*cellarea*0.000001,
                                        cowList = fullCountryData$ISO,
                                        aggMethod = "sum")
  poptot <- cbind(poptot, new_col)
}
colnames(poptot) <- yearcount


poptot <- as.data.frame(poptot)
poptot$ISO <- row.names(poptot)
poptot$ISO <- as.factor(poptot$ISO)

#pivot for tidy data
poptot <- pivot_longer(
  poptot, 
  -ISO,
  names_to = "yearcount", 
  values_to = "poptotal"
)

poptot$yearcount <- as.numeric(poptot$yearcount )

#merge to big dataframe
aei_country <- left_join(
  aei_country, 
  poptot, 
  by = c("ISO", "yearcount")
)

rm(poptot)
```

## Income

Using lpjgdppc

```{r}
income <- NULL
for (i in yearcount){
  new_col <- aggregateLPJmLdata2Country(input=lpjGDPpc[,i],
                                        cowList = fullCountryData$ISO,
                                        aggMethod = "median")
  income <- cbind(income, new_col)
}
colnames(income) <- yearcount


income <- as.data.frame(income)
income$ISO <- row.names(income)
income$ISO <- as.factor(income$ISO)

#pivot for tidy data
income <- pivot_longer(
  income, 
  -ISO,
  names_to = "yearcount", 
  values_to = "median_income"
)

income$yearcount <- as.numeric(income$yearcount )

#merge to big dataframe
aei_country <- left_join(
  aei_country, 
  income, 
  by = c("ISO", "yearcount")
)

rm(income)
```

## PET

```{r}
PET = apply(total_PET[,1:12,],c(1,3),sum) #summing all PET per year
pet_yr <- NULL
for (i in yearcount){
  new_col <- aggregateLPJmLdata2Country(input=PET[,i], #multiplying by cell area
                                        cowList = fullCountryData$ISO,
                                        aggMethod = "median")
  new_col <- new_col #m2 to ha conversion
  pet_yr <- cbind(pet_yr, new_col)
}
colnames(pet_yr) <- yearcount
pet_yr <- as.data.frame(pet_yr)
pet_yr$ISO <- row.names(pet_yr)
pet_yr$ISO <- as.factor(pet_yr$ISO)



#pivot for tidy data
pet_yr <- pivot_longer(
  pet_yr, 
  -ISO,
  names_to = "yearcount", 
  values_to = "pet_med_yr"
)

pet_yr$yearcount <-as.numeric(pet_yr$yearcount)

#merge to country area data
aei_country <- left_join(
  aei_country, 
  pet_yr, 
  by = c("ISO", "yearcount")
)


rm(pet_yr)

```

## Precipitation

```{r}
precip = apply(prec_landuse[,1:12,],c(1,3),sum) #summing all precip per year
precip_yr <- NULL
for (i in yearcount){
  new_col <- aggregateLPJmLdata2Country(input=precip[,i], #multiplying by cell area
                                        cowList = fullCountryData$ISO,
                                        aggMethod = "median")
  new_col <- new_col #m2 to ha conversion
  precip_yr <- cbind(precip_yr, new_col)
}
colnames(precip_yr) <- yearcount
precip_yr <- as.data.frame(precip_yr)
precip_yr$ISO <- row.names(precip_yr)
precip_yr$ISO <- as.factor(precip_yr$ISO)



#pivot for tidy data
precip_yr <- pivot_longer(
  precip_yr, 
  -ISO,
  names_to = "yearcount", 
  values_to = "precip_med_yr"
)

precip_yr$yearcount <-as.numeric(precip_yr$yearcount)

#merge to country area data
aei_country <- left_join(
  aei_country, 
  precip_yr, 
  by = c("ISO", "yearcount")
)


rm(precip_yr)
```

## Humidity

```{r}
#humidity as precipitation divided by potential evaopotranspiration
humidity = prec_landuse/total_PET

precip_yr <- NULL
for (i in yearcount){
  new_col <- aggregateLPJmLdata2Country(input=precip[,i], #multiplying by cell area
                                        cowList = fullCountryData$ISO,
                                        aggMethod = "median")
  new_col <- new_col #m2 to ha conversion
  precip_yr <- cbind(precip_yr, new_col)
}
colnames(precip_yr) <- yearcount
precip_yr <- as.data.frame(precip_yr)
precip_yr$ISO <- row.names(precip_yr)
precip_yr$ISO <- as.factor(precip_yr$ISO)



#pivot for tidy data
precip_yr <- pivot_longer(
  precip_yr, 
  -ISO,
  names_to = "yearcount", 
  values_to = "precip_med_yr"
)

precip_yr$yearcount <-as.numeric(precip_yr$yearcount)

#merge to country area data
aei_country <- left_join(
  aei_country, 
  precip_yr, 
  by = c("ISO", "yearcount")
)


rm(precip_yr)
```


## Ruggedness

Using data from @rileyTerrainRuggednessIndex1999 contained in the package `rethinking`

```{r}
library(rethinking)
data("rugged")
aei_country <- rugged %>% select(c("isocode", "rugged")) %>% 
  rename(ISO = isocode) %>% 
  left_join(aei_country, ., by = c("ISO"))

rm(rugged)

```

## Polity
