

Packages

```{r}
library(dplyr)
library(tidyr)
library(brms)
library(purrr)
library(ggplot2)
library(GGally)
```


Data

```{r load_data}
aei_small <- read.csv("/Volumes/RachelExternal/Thesis/Data_and_Plots/aei_small.csv")
aei_small_std <- read.csv("/Volumes/RachelExternal/Thesis/Data_and_Plots/aei_small_std.csv")
#aei_std <- read.csv("/p/projects/open/ledig/thesis/aei_std.csv")

#for ease of life
aei_small_std <- aei_small_std %>% filter(id %in% sample(levels(id),50))

summary(aei_small_std)
```



# Individual cell change model

run first model based on time series

```{r}
# nest data
by_id <-
  aei_small_std %>%
  group_by(id) %>%
  nest()
```

use default priors here, they allow things to vary nicely on a per country basis. 
```{r}
job::job({
id_fits_time_zib <-
  brm(data = by_country$data[[1]],
      formula = bf(irrfrac ~ 1 + yearcount,
                   zi ~ 1 + yearcount),
      family = zero_inflated_beta(),
      control = list(adapt_delta = 0.95, 
                     max_treedepth=11),
      cores = 2, 
      iter = 1000, 
      chains = 2,
      seed = 2,
      file = "/Volumes/RachelExternal/Thesis/Thesis/3_pretty_scripts_for_profs/Fits/final_fits/Country_fits_pred_intercept_years_small.Rds")
})

# map to all other countries
job::job({
  yearcount_fits <- 
  by_country %>%
  mutate(model = map(data, ~update(id_fits_time_zib, newdata = ., seed = 2)))
  
  
#save nested df for later analysis on local machine
saveRDS(yearcount_fits, "/Volumes/RachelExternal/Thesis/Thesis/3_pretty_scripts_for_profs/Fits/final_fits/Country_fits_pred_intercept_years_full.Rds")
  })


```



# id Models


## Unconditional Means Model ID

aims to look at general effects across cells without time as a component


### Priors

```{r}
get_prior(data = aei_small_std,
      formula = bf(irrfrac ~ 1 + (1|id),
                   phi ~ 1 + (1|id),
                   zi ~ 1 + (1|id)),
      family = zero_inflated_beta())
```
Visualization of priors on the outcome scale. Remember, the intercept and zi parameters are on a logit scale (use the inverse logit, `plogis()`) and phi is on a log scale (use the `exp()`).


## Model

Set the priors
```{r}
uncond_means_priors_id <- 
  c(set_prior("student_t(3, 0, 2.5)", class = "Intercept"), #Global Intercept
            set_prior("student_t(3, 0, 2.5)", class = "Intercept", dpar = "phi"), #phi Intercept
            set_prior("logistic(0, 1)", class = "Intercept", dpar = "zi")) #slope parameters, knowing that 
    
```



```{r}
job::job({
uncond_means_id <-
  brm(data = aei_small_std,
      formula = bf(irrfrac ~ 1 + (1|id),
                  phi ~ 1 + (1|id),
                   zi ~ 1 + (1|id)),
      family = zero_inflated_beta(link = "logit", 
                                  link_phi = "log", 
                                  link_zi = "logit"),
        control = list(adapt_delta = 0.95, 
                       max_treedepth=11),
      prior = uncond_means_priors_id,
      cores = 2, 
      iter = 1000, 
      chains = 4,
      seed = 2,
      file = "/Volumes/RachelExternal/Thesis/Thesis/3_pretty_scripts_for_profs/Fits/tiny_trials/simple_uncond_means_id_small.Rds")
})
```


## Unconditional Growth Model ID
aims to look at general effects across cells and over time prior to adding other preds

```{r}
get_prior(data = aei_small_std,
      formula = bf(irrfrac ~ 1 + yearcount + (1 + yearcount|id),
                   phi ~ 1 + yearcount + (1 + yearcount|id),
                   zi ~ 1 + yearcount + (1 + yearcount|id)),
      family = zero_inflated_beta(link = "logit", 
                                  link_phi = "log", 
                                  link_zi = "logit"))
```



```{r}
uncond_growth_id <-
  brm(data = aei_small_std,
      formula = bf(irrfrac ~ 1 + yearcount + (1 + yearcount|id),
                   zi ~ 1 + yearcount + (1 + yearcount|id)),
      family = zero_inflated_beta(link = "logit", 
                                  link_phi = "log", 
                                  link_zi = "logit"),
      prior = c(prior(normal(0, 100), class = b)),
        control = list(adapt_delta = 0.95, 
                       max_treedepth=11),
      cores = 2, 
      inits = "0", #had trouble initializing beta_lpdf. second shape parameter was 0. Solution found here https://discourse.mc-stan.org/t/rejecting-initial-value/7152/4
      iter = 1000, 
      chains = 4,
      seed = 348,
      file = "/Volumes/RachelExternal/Thesis/Thesis/3_pretty_scripts_for_profs/Fits/simple_uncond_growth_id_small.Rds")
```


## Full Simple model ID

```{r}
full_simple_id <-
  brm(data = aei_small_std,
      formula = bf(irrfrac ~ 0 + Intercept + yearcount + ISO + ISO:yearcount + (1 + yearcount|id), 
                   zi ~ 0 + Intercept + yearcount + ISO + ISO:yearcount + (1 + yearcount|id)),
      family = zero_inflated_beta(link = "logit", 
                                  link_phi = "log", 
                                  link_zi = "logit"),
      prior = c(prior(normal(0, 100), class = b)),
        control = list(adapt_delta = 0.95, 
                       max_treedepth=11),
      cores = 2, 
      iter = 1000, 
      inits = "0",
      chains = 4,
      seed = 2,
      file = "/Volumes/RachelExternal/Thesis/Thesis/3_pretty_scripts_for_profs/Fits/simple_full_id_small2.Rds")
```


# Complex model 

## Biophysical

### Unconditional means model

```{r}

```

### Unconditional Growth model


### Full model


## Biophysical and Sociological


### Unconditional means model

```{r}

```

### Unconditional Growth model


### Full model




without interactions, at lower levels they are easier to sort out but here its too complex

```{r}
full_simple <-
  brm_multiple(data = folds,
               formula = bf(irrfrac ~ 0 + Intercept + yearcount + DD.regime + dist + popdens + pet + gdppc + rugged + (1 + yearcount + popdens + DD.regime + gdppc|id),
                            zi ~ 0 + Intercept + yearcount + DD.regime + dist + popdens + pet + gdppc + rugged + (1 + yearcount + popdens + DD.regime + gdppc|id)),
               family = zero_inflated_beta(link = "logit", 
                                  link_phi = "log", 
                                  link_zi = "logit"),
               prior = c(prior(normal(0, 100), class = b)),
               control = list(adapt_delta = 0.95, 
                       max_treedepth=11),
               cores = 2, 
               iter = 1000, 
               chains = 4,
               seed = 2,
               file = "/Volumes/RachelExternal/Thesis/Thesis/3_pretty_scripts_for_profs/Fits/complex_full_small.Rds")
```




