

# Packages

```{r}
library(dplyr)
library(tidyr)
library(brms)
library(purrr)
library(ggplot2)
library(GGally)
library(parallel)
library(kableExtra)
library(rethinking)
library(devtools)
#install_github('mcooper/moranfast')
library(moranfast)
library(Rcpp)
library(sp)
library(geoR)
library(gstat)
library(agrmt)

```








# Data

```{r load_data}
aei_cropclean <- read.csv("/Volumes/RachelExternal/Thesis/Data_and_Plots/aei_nocrop.csv")
aei_cropclean_sample <- read.csv("/Volumes/RachelExternal/Thesis/Data_and_Plots/aei_nocropland_sample.csv")
```

```{r morani}
india <- aei_cropclean %>% subset(yearcount == 45 & ISO == "IND") %>% dplyr::select(id, lat, lon, irrfrac) #india for the year 2005
moranfast(india$irrfrac, india$lon, india$lat)
```

```{r}
coordinates(india)= ~ lon+lat
bubble(india, zcol='irrfrac', fill=TRUE, do.sqrt=FALSE, maxsize=3) #check the bubble
# TheVariogram=variogram(irrfrac~1, data=india)
# plot(TheVariogram)
```

# Standardization

```{r}
normalized<-function(y) {
  x<-y[!is.na(y)]
  x<-(x - min(x)) / (max(x) - min(x))
  y[!is.na(y)]<-x
  return(y)
}

centered <-function(y) {
  x<-y[!is.na(y)]
  x<-(x - mean(x)) / (sd(x)*2)
  y[!is.na(y)]<-x
  return(y)
}
```

Some variables need to be logged before centering on 0. Check out the predictors.
```{r}
aei_cropclean_sample %>% dplyr::select(precip, humid,  dist, medinc, rugged, gdppc, popdens) %>% summary()
```
Recall the histograms, only rugged and pet had (semi) normal distributions (they were not normally distributed at all but at least both were not seemingly zero inflated like the others). Knowing that pet will not be used and rugged just needs to be scaled, take the log of the other variables. All predictors to be logged contain zeros (or neg, medinc) except for gdp. All of the 0s here represent different things:

- dist, cells with irrigation have a value of 0. cells with no irrigation have a dist value greater than 22.83487km.
- popdens, here 0 means 0. 
- precip, here 0 means 0 as well. 
- humid, same as precip.
- medinc, this one is a conundrum, as there are are negative median yield increases

Here are the min non zero values for each.
```{r}
minnz(aei_cropclean_sample$dist)
minnz(aei_cropclean_sample$popdens)
minnz(aei_cropclean_sample$precip)
minnz(aei_cropclean_sample$humid)
```

This is controversial, but I will just add 1 to these columns and take the log. In other places, recommendations are made to add half of the minimum non zero value. Although this seems fine, for popdens and precip and humid this would produce negative values on the log scale. The solution of adding 1 is not my most favorite idea but must be done.

```{r}

aei_cropclean_sample <- 
  aei_cropclean_sample %>% 
  mutate(popdens = popdens + 1, 
         dist = dist + 1,
         precip = precip + 1,
         humid = humid+ 1,
         medinc = medinc + 1)

aei_cropclean_sample %>% select(precip, humid, popdens, gdppc, dist, medinc, rugged) %>% summary()

```
 the only issue here will be medinc, still with neg values post log transform. 

```{r}
aei_cropclean_sample_std <-
  aei_cropclean_sample %>% 
  mutate(across(c(17:21,25:26, 31), centered)) %>% 
  mutate(across(c(2,3,6,27, 29), as.factor)) %>% 
  select(lat, lon, id, ISO, Countryname, six_regions, eight_regions, yearcount, irrfrac,
         dist, medinc, precip, humid, rugged, crl_yld, ag_gdp, popdens, gdppc, 
         DD.regime,DD.category, Democracy)


summary(aei_cropclean_sample_std)

write.csv(aei_cropclean_sample_std, "/Volumes/RachelExternal/Thesis/Data_and_Plots/aei_cropclean_sample_std.csv")

```




# Prior Vis


Knowing that the most complicated model formula I will use is the one below, visualize the priors.
```{r}
get_prior(data = df, 
          formula = bf(irrfrac ~ 1 + eight_regions + yearcount + precip + medinc + dist + rugged + popdens + gdppc + Democracy +
                     eight_regions:yearcount + precip:yearcount + medinc:yearcount + dist:yearcount + rugged:yearcount + popdens:yearcount + gdppc:yearcount + Democracy:yearcount + (1 + yearcount |ISO),
                   zi ~ 1 + dist + gdppc),
      family = zero_inflated_beta(link = "logit", 
                                  link_phi = "identity", 
                                  link_zi = "logit"))
```

## Beta

These priors have a logit link. i.e. to visualize them we need and inv logit.

### mu Intercept

```{r}
tibble(`Default Student T(3, 0, 2.5)` = rt(10000, 3, 0)*2.5) %>%  
  mutate(`plogis Outcome Scale` = plogis(`Default Student T(3, 0, 2.5)`)) %>% 
  pivot_longer(cols = c(`Default Student T(3, 0, 2.5)`, `plogis Outcome Scale`), 
               names_to = "Transformation", 
               values_to = "Distribution") %>% 
  ggplot() +
  geom_histogram(aes(x = Distribution, color = Transformation)) + 
  xlab(NULL) + facet_wrap(~Transformation, scales = "free") +
  ggtitle("Default Global Intercept Prior - Student T(3, 0, 2.5)")
```

### mu Slopes


```{r}
tibble(`Proper Normal(0, 1)` = rnorm(10000, 0, 1)) %>%  
  mutate(`plogis Outcome Scale` = plogis(`Proper Normal(0, 1)`)) %>% 
  pivot_longer(cols = c(`Proper Normal(0, 1)`, `plogis Outcome Scale`), 
               names_to = "Transformation", 
               values_to = "Distribution") %>% 
  ggplot() +
  geom_histogram(aes(x = Distribution, color = Transformation)) + 
  xlab(NULL) + facet_wrap(~factor(Transformation, levels=c('Proper Normal(0, 1)','plogis Outcome Scale')), scales = "free") +
  ggtitle("mu Slope Prior - Normal(0, 1)")
```

### phi

As I am not predicting phi in the model formula, it has an identity link, meaning that it can just be visualized. 


```{r}
tibble(`Default Gamma(0.01, 0.01)` = rgamma(10000, rate = 0.01, shape = 0.01)) %>%  #according to the stan documentation
  pivot_longer(cols = c(`Default Gamma(0.01, 0.01)`), 
               names_to = "Transformation", 
               values_to = "Distribution") %>% 
  ggplot() +
  geom_histogram(aes(x = Distribution, color = Transformation), bins = 100) + 
  xlab(NULL) + xlim(-0.1,0.1) +
  ggtitle("Default Zi Intercept Prior - `Default Gamma(0.01, 0.01)`")
```

This is super tight.. I wonder if I am plotting it correctly.


## Bernoullii

These priors have a logit link. i.e. to visualize them we need and inv logit.

### Intercept

```{r}
tibble(`Default Logistic(0, 1)` = rlogis(10000, 0, 1)) %>%  
  mutate(`plogis Outcome Scale` = plogis(`Default Logistic(0, 1)`)) %>% 
  pivot_longer(cols = c(`Default Logistic(0, 1)`, `plogis Outcome Scale`), 
               names_to = "Transformation", 
               values_to = "Distribution") %>% 
  ggplot() +
  geom_histogram(aes(x = Distribution, color = Transformation)) + 
  xlab(NULL) + facet_wrap(~Transformation, scales = "free") +
  ggtitle("Default Zi Intercept Prior - `Default Logistic(0, 1)`")
```

### Slopes


```{r}
tibble(`Proper Normal(0, 1)` = rnorm(10000, 0, 1)) %>%  
  mutate(`plogis Outcome Scale` = plogis(`Proper Normal(0, 1)`)) %>% 
  pivot_longer(cols = c(`Proper Normal(0, 1)`, `plogis Outcome Scale`), 
               names_to = "Transformation", 
               values_to = "Distribution") %>% 
  ggplot() +
  geom_histogram(aes(x = Distribution, color = Transformation)) + 
  xlab(NULL) + facet_wrap(~factor(Transformation, levels=c('Proper Normal(0, 1)','plogis Outcome Scale')), scales = "free") +
  ggtitle("Zi Slope Prior - Normal(0, 1)")
```



## cor and sd



# `brm()` settings
```{r brmsettings}
df = aei_cropclean_sample
ncore = 4 #detectCores()
options(mc.cores = parallel::detectCores())
niter = 2000
nchain = 4
seed = 2

#saving location
folder = "/Volumes/RachelExternal/Thesis/Thesis/final/fits/"
```



# Individual Country Change Model

run first model based on time series

```{r}
# nest data
by_country <-
  aei_small_std %>%
  group_by(Countryname) %>%
  nest()
```

use default priors here, they allow things to vary nicely on a per country basis. 
```{r}
job::job({
id_fits_time_zib <-
  brm(data = by_country$data[[1]],
      formula = irrfrac ~ 1 + yearcount,
      family = gaussian(), 
      control = list(adapt_delta = 0.95, 
                     max_treedepth=11),
      cores = ncore, 
      iter = niter, 
      chains = nchain,
      seed = seed,
      file = "/Volumes/RachelExternal/Thesis/Thesis/3_pretty_scripts_for_profs/Fits/final_fits/Country_fits_pred_intercept_years_small.Rds")
})

# map to all other countries
job::job({
  yearcount_fits <- 
  by_country %>%
  mutate(model = map(data, ~update(id_fits_time_zib, newdata = ., seed = seed)))
  
  
#save nested df for later analysis on local machine
saveRDS(yearcount_fits, "/Volumes/RachelExternal/Thesis/Thesis/3_pretty_scripts_for_profs/Fits/final_fits/Country_fits_pred_intercept_years_full.Rds")
  })


```


# Simple model

## Unconditional Means Model ISO

```{r}
get_prior(data = df,
      formula = bf(irrfrac ~ 1 + (1|ISO), 
                   zi ~ 1 + (1|ISO)),
      family = zero_inflated_beta(link = "logit", 
                                  link_phi = "identity", 
                                  link_zi = "logit"))
```
Using default priors here. 
```{r}
uncond_means_prior_ISO <- 
  c(set_prior("student_t(3, 0, 2.5)", class = "Intercept"), #Global Intercept
    set_prior("gamma(0.01, 0.01)", class = "phi"), #phi Intercept with identity link
    set_prior("logistic(0, 1)", class = "Intercept", dpar = "zi" ), #zi Intercept
    set_prior("student_t(3, 0, 2.5)", class = "sd")) #stdev parameters
```


```{r}
job::job({
uncond_means_ISO <-
  brm(data = df,
      formula = bf(irrfrac ~ 1 + (1|ISO),
                   zi ~ 1 + (1|ISO)),
      family = zero_inflated_beta(link = "logit", 
                                  link_phi = "identity", 
                                  link_zi = "logit"),
      prior = uncond_means_prior_ISO,
      control = list(adapt_delta = 0.99, 
                       max_treedepth=11),
      sample_prior = "yes",
      cores = ncore, 
      iter = niter, 
      chains = nchain,
      seed = ncore,
      threads = threading(2),
      backend = "cmdstanr",
      file = paste0(folder,"ISO_simple_uncondmeans.Rds"))
})
```

5000 samples: 210.0 seconds. 2 chains, 2 cores, threading(2)

```{r}
print(uncond_means_ISO)
```


```{r}
plot(uncond_means_ISO)
```
These are untransformed. 


```{r}
posterior_samples(uncond_means_ISO) %>%
  select(starts_with(c("b_", "Intercept", "sd_", "phi"))) %>% 
  mutate(across(.cols = starts_with("phi"), exp)) %>% #hmmmmmm
  mutate(across(.cols = starts_with(c("b_", "Intercept")), plogis)) %>% 
  posterior_summary() %>% 
  as.data.frame()
```



```{r}
pp_check(uncond_means_ISO, xlim = c(0, 0.1))
```


## Unconditional Growth Model ISO



```{r}
get_prior(data = df,
      formula = bf(irrfrac ~ 1 + yearcount + (1 + yearcount|ISO),
                   zi ~ 1 + yearcount + (1 + yearcount|ISO)),
      family = zero_inflated_beta(link = "logit", 
                                  link_phi = "identity", 
                                  link_zi = "logit"))
```


Using default priors here. 
```{r}
uncond_growth_prior_ISO <- 
  c(set_prior("student_t(3, 0, 2.5)", class = "Intercept"), #Global Intercept
    set_prior("gamma(0.01, 0.01)", class = "phi"), #phi Intercept with identity link
    set_prior("logistic(0, 1)", class = "Intercept", dpar = "zi" ), #zi Intercept
    set_prior("student_t(3, 0, 2.5)", class = "sd"), #stdev parameters
    set_prior("lkj(2)", class = "cor"), #correlation prior
    set_prior("normal(0,1)", class = "b")) #slope prior 
```


```{r}
job::job({
uncond_growth_ISO <-
  brm(data = df,
      formula = bf(irrfrac ~ 1 + yearcount + (1 + yearcount|ISO),
                   zi ~ 1 + yearcount + (1 + yearcount|ISO)),
      family = zero_inflated_beta(link = "logit", 
                                  link_phi = "identity", 
                                  link_zi = "logit"),
      prior = uncond_growth_prior_ISO,
        control = list(adapt_delta = 0.95, 
                       max_treedepth=11),
      sample_prior = "yes",
      cores = ncore, 
      iter = niter, 
      chains = nchain,
      threads = threading(2),
      backend = "cmdstanr",
      inits = "0",
      seed = seed,
      file = paste0(folder,"ISO_simple_uncondgrowth.Rds"))
})
```

5000: Total execution time: 1023.9 seconds. 2 chains, 2 cores, threading(2)

```{r}
summary(uncond_growth_ISO)
```

```{r}
plot(uncond_growth_ISO)
```


```{r}

posterior_samples(uncond_growth_ISO) %>%
  select(starts_with(c("b_", "Intercept", "sd_", "phi", "cor_ISO"))) %>% 
  mutate(across(.cols = starts_with("phi"), exp)) %>% #hmmmmmm
  mutate(across(.cols = starts_with(c("b_", "Intercept")), plogis)) %>% 
  posterior_summary() %>% 
  as.data.frame()

```
```{r}
conditional_effects(uncond_growth_ISO)
```

```{r}
pp_check(uncond_growth_ISO)
```

## Full Simple model ISO


```{r}
get_prior(data = df,
      formula = bf(irrfrac ~ 1 +  + eight_regions + yearcount + eight_regions:yearcount + (1 + yearcount|ISO),
                   zi ~ 1 + eight_regions + yearcount + eight_regions:yearcount + (1 + yearcount|ISO)),
      family = zero_inflated_beta(link = "logit", 
                                  link_phi = "identity", 
                                  link_zi = "logit"))
```

```{r}
full_simple_prior_ISO <- 
  c(set_prior("student_t(3, 0, 2.5)", class = "Intercept"), #Global Intercept
    set_prior("gamma(0.01, 0.01)", class = "phi"), #phi Intercept with identity link
    set_prior("logistic(0, 1)", class = "Intercept", dpar = "zi" ), #zi Intercept
    set_prior("student_t(3, 0, 2.5)", class = "sd"), #stdev parameters
    set_prior("lkj(2)", class = "cor"), #correlation prior
    set_prior("normal(0,1)", class = "b")) #slope prior 
```

```{r}
job::job({
full_simple_ISO <-
  brm(data = df,
      formula = bf(irrfrac ~ 1 +  + eight_regions + yearcount + eight_regions:yearcount + (1 + yearcount|ISO),
                   zi ~ 1 + eight_regions + yearcount + eight_regions:yearcount + (1 + yearcount|ISO)),
      family = zero_inflated_beta(link = "logit", 
                                  link_phi = "identity", 
                                  link_zi = "logit"),
      prior = full_simple_prior_ISO,
      control = list(adapt_delta = 0.95, 
                       max_treedepth=11),
      sample_prior = "yes",
      inits = "0",
      cores = ncore, 
      threads = threading(2),
      backend = "cmdstanr",
      iter = niter, 
      chains = nchain,
      seed = seed,
      file = paste0(folder,"ISO_simple_full_5000.Rds"))
})
```
5000: Total execution time: 1236.6 seconds. 2 chains, 2 cores, threading(2)
```{r}
print(full_simple_ISO)
```


```{r}
posterior_samples(full_simple_ISO) %>%
  select(starts_with(c("b_", "Intercept", "sd_", "phi", "cor_ISO"))) %>% 
  mutate(across(.cols = starts_with("phi"), exp)) %>% #hmmmmmm
  mutate(across(.cols = starts_with(c("b_", "Intercept")), plogis)) %>% 
  posterior_summary() %>% 
  as.data.frame()
```

```{r}
conditional_effects(full_simple_ISO)
```

```{r}
pp_check(full_simple_ISO)
```


# Complex model 


## Biophysical and Sociological

Unconditional means and growth models are the same as above. 

### Full model

```{r}
get_prior(data = df, 
          formula = bf(irrfrac ~ 1 + eight_regions + yearcount + precip + medinc + dist + rugged + popdens + gdppc + Democracy +
                     eight_regions:yearcount + precip:yearcount + medinc:yearcount + dist:yearcount + rugged:yearcount + popdens:yearcount + gdppc:yearcount + Democracy:yearcount + (1 + yearcount |ISO),
                   zi ~ 1 + dist + gdppc),
      family = zero_inflated_beta(link = "logit", 
                                  link_phi = "identity", 
                                  link_zi = "logit"))
```

```{r}
biosoc_full_prior_ISO <- 
  c(set_prior("student_t(3, 0, 2.5)", class = "Intercept"), #Global Intercept
    set_prior("gamma(0.01, 0.01)", class = "phi"), #phi Intercept with identity link
    set_prior("logistic(0, 1)", class = "Intercept", dpar = "zi" ), #zi Intercept
    set_prior("student_t(3, 0, 2.5)", class = "sd"), #stdev parameters
    set_prior("lkj(2)", class = "cor"), #correlation prior
    set_prior("normal(0,1)", class = "b"), 
    set_prior("normal(0,1)", class = "b", dpar = "zi")) #slope prior 
```


```{r}
job::job({
  biosoc_full__ISO <-
  brm(data = df,
      formula = bf(irrfrac ~ 1 + eight_regions + yearcount + precip + medinc + dist + rugged + popdens + gdppc + Democracy +
                     eight_regions:yearcount + precip:yearcount + medinc:yearcount + dist:yearcount + rugged:yearcount + popdens:yearcount + gdppc:yearcount + Democracy:yearcount + (1 + yearcount |ISO),
                   zi ~ 1 + dist + gdppc),
      family = zero_inflated_beta(link = "logit", 
                                  link_phi = "identity", 
                                  link_zi = "logit"),
      prior = biosoc_full_prior_ISO,
        control = list(adapt_delta = 0.95, 
                       max_treedepth=11),
      sample_prior = "yes",
      cores = ncore, 
      iter = niter, 
      chains = nchain,
      seed = seed,
      inits = "0",
      threads = threading(2),
      backend = "cmdstanr",
      file = paste0(folder,"ISO_biosocio_full_5000.Rds"))
})
```
No way: Total execution time: 1668.6 seconds. 2 chains, 2 cores, 5000 obvs. Roughly a half hour.
```{r}
print(biosoc_full__ISO)
```


```{r}
plot(biosoc_full__ISO)
```


```{r}
posterior_samples(biosoc_full__ISO) %>%
  select(starts_with(c("b_", "Intercept", "sd_", "phi", "cor_ISO"))) %>% 
  mutate(across(.cols = starts_with("phi"), exp)) %>% #hmmmmmm
  mutate(across(.cols = starts_with(c("b_", "Intercept")), plogis)) %>% 
  posterior_summary() %>% 
  as.data.frame()
```

```{r}
conditional_effects(biosoc_full__ISO)
```

```{r}
pp_check(full_simple_ISO)

```



## Other less mathematically rigorous models

### Bio only

```{r}
get_prior(data = df,
          formula = bf(irrfrac ~ 1 + yearcount + precip + medinc + dist + rugged + eight_regions + (1 + yearcount|ISO),
                   zi ~ 1 + dist), 
          family = zero_inflated_beta(link = "logit", 
                                  link_phi = "identity", 
                                  link_zi = "logit"))
```


```{r}
bio_only_prior_ISO <- 
  c(set_prior("student_t(3, 0, 2.5)", class = "Intercept"), #Global Intercept
    set_prior("gamma(0.01, 0.01)", class = "phi"), #phi Intercept with identity link
    set_prior("logistic(0, 1)", class = "Intercept", dpar = "zi" ), #zi Intercept
    set_prior("student_t(3, 0, 2.5)", class = "sd"), #stdev parameters
    set_prior("lkj(2)", class = "cor"), #correlation prior
    set_prior("normal(0,1)", class = "b"), 
    set_prior("normal(0,1)", class = "b", dpar = "zi")) #slope prior 
```


```{r}
job::job({
bio_only_ISO <-
  brm(data = df,
      formula = bf(irrfrac ~ 1 + yearcount + precip + medinc + dist + rugged + eight_regions + (1 + yearcount|ISO),
                   zi ~ 1 + dist),
      family = zero_inflated_beta(link = "logit", 
                                  link_phi = "identity", 
                                  link_zi = "logit"),
      prior = bio_only_prior_ISO,
        control = list(adapt_delta = 0.95, 
                       max_treedepth=11),
      sample_prior = "yes",
      cores = ncore, 
      iter = niter, 
      chains = nchain,
      inits = "0",
      threads = threading(2), 
      backend = "cmdstanr",
      seed = seed,
      file = paste0(folder,"ISO_bioonly_5000.Rds"))
})
```

```{r}
print(bio_only_ISO)
```

```{r}
plot(bio_only_ISO)
```


```{r}
posterior_samples(bio_only_ISO) %>%
  select(starts_with(c("b_", "Intercept", "sd_", "phi", "cor_ISO"))) %>% 
  mutate(across(.cols = starts_with("phi"), exp)) %>% #hmmmmmm
  mutate(across(.cols = starts_with(c("b_", "Intercept")), plogis)) %>% 
  posterior_summary() %>% 
  as.data.frame()
```

```{r}
conditional_effects(bio_only_ISO)
```



### Biosoc, no interactions


```{r}
get_prior(data = df,
          formula = bf(irrfrac ~ 1 + yearcount + precip + medinc + dist + rugged + Democracy + gdppc + popdens + eight_regions + (1 + yearcount|ISO),
                   zi ~ 1 + dist + gdppc), 
          family = zero_inflated_beta(link = "logit", 
                                  link_phi = "identity", 
                                  link_zi = "logit"))
```


```{r}
biosoc_no_inter_prior_ISO <- 
  c(set_prior("student_t(3, 0, 2.5)", class = "Intercept"), #Global Intercept
    set_prior("gamma(0.01, 0.01)", class = "phi"), #phi Intercept with identity link
    set_prior("logistic(0, 1)", class = "Intercept", dpar = "zi" ), #zi Intercept
    set_prior("student_t(3, 0, 2.5)", class = "sd"), #stdev parameters
    set_prior("lkj(2)", class = "cor"), #correlation prior
    set_prior("normal(0,1)", class = "b"), 
    set_prior("normal(0,1)", class = "b", dpar = "zi")) #slope prior 
```


```{r}
job::job({
biosoc_no_inter_ISO <-
  brm(data = df,
      formula = bf(irrfrac ~ 1 + yearcount + precip + medinc + dist + rugged + Democracy + gdppc + popdens + eight_regions + (1 + yearcount|ISO),
                   zi ~ 1 + dist + gdppc),
      family = zero_inflated_beta(link = "logit", 
                                  link_phi = "identity", 
                                  link_zi = "logit"),
      prior = biosoc_no_inter_prior_ISO,
        control = list(adapt_delta = 0.95, 
                       max_treedepth=11),
      sample_prior = "yes",
      cores = ncore, 
      iter = niter, 
      chains = nchain,
      inits = "0",
      threads = threading(2), 
      backend = "cmdstanr",
      seed = seed,
      file = paste0(folder,"ISO_biosoc_no_inter_5000.Rds"))
})
```

```{r}
print(biosoc_no_inter_ISO)
```

```{r}
posterior_samples(biosoc_no_inter_ISO) %>%
  select(starts_with(c("b_", "Intercept", "sd_", "phi", "cor_ISO"))) %>% 
  mutate(across(.cols = starts_with("phi"), exp)) %>% #hmmmmmm
  mutate(across(.cols = starts_with(c("b_", "Intercept")), plogis)) %>% 
  posterior_summary() %>% 
  as.data.frame()
```

```{r}
conditional_effects(biosoc_no_inter_ISO)
```

```{r}
pp_check(biosoc_no_inter_ISO)
```


### Biosoc, more varying slopes


```{r}
get_prior(data = df,
          formula = bf(irrfrac ~ 1 + yearcount + precip + medinc + dist + rugged + Democracy + gdppc + popdens + eight_regions +
                         (1 + yearcount + Democracy + gdppc + popdens|ISO),
                   zi ~ 1 + dist + gdppc), 
          family = zero_inflated_beta(link = "logit", 
                                  link_phi = "identity", 
                                  link_zi = "logit"))
```


```{r}
biosoc_v_slopes_prior_ISO <- 
  c(set_prior("student_t(3, 0, 2.5)", class = "Intercept"), #Global Intercept
    set_prior("gamma(0.01, 0.01)", class = "phi"), #phi Intercept with identity link
    set_prior("logistic(0, 1)", class = "Intercept", dpar = "zi" ), #zi Intercept
    set_prior("student_t(3, 0, 2.5)", class = "sd"), #stdev parameters
    set_prior("lkj(2)", class = "cor"), #correlation prior
    set_prior("normal(0,1)", class = "b"), 
    set_prior("normal(0,1)", class = "b", dpar = "zi")) #slope prior 
```


```{r}
job::job({
biosoc_v_slopes_ISO <-
  brm(data = df,
      formula = bf(irrfrac ~ 1 + yearcount + precip + medinc + dist + rugged + Democracy + gdppc + popdens + eight_regions +
                         (1 + yearcount + Democracy + gdppc + popdens|ISO),
                   zi ~ 1 + dist + gdppc),
      family = zero_inflated_beta(link = "logit", 
                                  link_phi = "identity", 
                                  link_zi = "logit"),
      prior = biosoc_v_slopes_prior_ISO,
        control = list(adapt_delta = 0.95, 
                       max_treedepth=11),
      sample_prior = "yes",
      cores = ncore, 
      iter = niter, 
      chains = nchain,
      inits = "0",
      threads = threading(2), 
      backend = "cmdstanr",
      seed = seed,
      file = paste0(folder,"ISO_biosoc_v_slopes_5000.Rds"))
})
```

```{r}
print(biosoc_v_slopes_ISO)
```

```{r}
plot(biosoc_v_slopes_ISO)
```



```{r}
posterior_samples(biosoc_v_slopes_ISO) %>%
  select(starts_with(c("b_", "Intercept", "sd_", "phi", "cor_ISO"))) %>% 
  mutate(across(.cols = starts_with("phi"), exp)) %>% #hmmmmmm
  mutate(across(.cols = starts_with(c("b_", "Intercept")), plogis)) %>% 
  posterior_summary() %>% 
  as.data.frame()
```

```{r}
conditional_effects(biosoc_v_slopes_ISO)
```



#Experiments with Syntax


## Correlated grouping terms

```{r}
get_prior(data = df,
      formula = bf(irrfrac ~ 1 + (1|i|ISO), 
                   zi ~ 1 + (1|i|ISO)),
      family = zero_inflated_beta(link = "logit", 
                                  link_phi = "identity", 
                                  link_zi = "logit"))
```
Using default priors here. 
```{r}
simple_cor_gterms_prior_ISO <- 
  c(set_prior("student_t(3, 0, 2.5)", class = "Intercept"), #Global Intercept
    set_prior("gamma(0.01, 0.01)", class = "phi"), #phi Intercept with identity link
    set_prior("logistic(0, 1)", class = "Intercept", dpar = "zi" ), #zi Intercept
    set_prior("student_t(3, 0, 2.5)", class = "sd"), #stdev parameters
    set_prior("lkj(2)", class = "cor"))
```


```{r}
job::job({
simple_cor_gterms_ISO <-
  brm(data = df,
      formula = bf(irrfrac ~ 1 + (1|i|ISO), 
                   zi ~ 1 + (1|i|ISO)),
      family = zero_inflated_beta(link = "logit", 
                                  link_phi = "identity", 
                                  link_zi = "logit"),
      prior = simple_cor_gterms_prior_ISO,
      control = list(adapt_delta = 0.99, 
                       max_treedepth=11),
      sample_prior = "yes",
      cores = ncore, 
      iter = niter, 
      chains = nchain,
      seed = ncore,
      threads = threading(2),
      backend = "cmdstanr",
      file = paste0(folder,"ISO_simple_cor_gterms_5000.Rds"))
})
```

5000 samples: Total execution time: 328.9 seconds. 2 chains, 2 cores, threading(2)

```{r}
print(simple_cor_gterms_ISO)
```


```{r}
posterior_samples(simple_cor_gterms_ISO, pars = "b_") %>% 
  mutate(across(.cols = starts_with("b_phi"), exp)) %>% 
  mutate(across(.cols = c(- starts_with("b_phi")), plogis)) %>% 
  posterior_summary() %>% 
  as.data.frame() 
```

```{r}
pp_check(simple_cor_gterms_ISO)
```



## Unconditional Growth Model ISO



```{r}
get_prior(data = df,
      formula = bf(irrfrac ~ 1 + yearcount + (1 + yearcount|i|ISO),
                   zi ~ 1 + yearcount + (1 + yearcount|i|ISO)),
      family = zero_inflated_beta(link = "logit", 
                                  link_phi = "identity", 
                                  link_zi = "logit"))
```


Using default priors here. 
```{r}
yrct_cor_gterms_prior_ISO <- 
  c(set_prior("student_t(3, 0, 2.5)", class = "Intercept"), #Global Intercept
    set_prior("gamma(0.01, 0.01)", class = "phi"), #phi Intercept with identity link
    set_prior("logistic(0, 1)", class = "Intercept", dpar = "zi" ), #zi Intercept
    set_prior("student_t(3, 0, 2.5)", class = "sd"), #stdev parameters
    set_prior("lkj(2)", class = "cor"), #correlation prior
    set_prior("normal(0,1)", class = "b")) #slope prior 
```


```{r}
job::job({
yrct_cor_gterms_ISO <-
  brm(data = df,
      formula = bf(irrfrac ~ 1 + yearcount + (1 + yearcount|i|ISO),
                   zi ~ 1 + yearcount + (1 + yearcount|i|ISO)),
      family = zero_inflated_beta(link = "logit", 
                                  link_phi = "identity", 
                                  link_zi = "logit"),
      prior = yrct_cor_gterms_prior_ISO,
        control = list(adapt_delta = 0.95, 
                       max_treedepth=11),
      sample_prior = "yes",
      cores = ncore, 
      iter = niter, 
      chains = nchain,
      threads = threading(2),
      backend = "cmdstanr",
      inits = "0",
      seed = seed,
      file = paste0(folder,"ISO_yrct_cor_gterms_5000.Rds"))
})
```

5000:  2 chains, 2 cores, threading(2)

```{r}
summary(uncond_growth_ISO)
```

```{r}
posterior_samples(uncond_growth_ISO, pars = "b_") %>% 
  mutate(across(.cols = starts_with("b_phi"), exp)) %>% 
  mutate(across(.cols = c(- starts_with("b_phi")), plogis)) %>% 
  posterior_summary() %>% 
  as.data.frame() 
```
```{r}
conditional_effects(uncond_growth_ISO)
```

```{r}
pp_check(uncond_growth_ISO)
```

## Full Simple model ISO


```{r}
get_prior(data = df,
      formula = bf(irrfrac ~ 1 +  + eight_regions + yearcount + eight_regions:yearcount + (1 + yearcount|ISO),
                   zi ~ 1 + eight_regions + yearcount + eight_regions:yearcount + (1 + yearcount|ISO)),
      family = zero_inflated_beta(link = "logit", 
                                  link_phi = "identity", 
                                  link_zi = "logit"))
```

```{r}
full_simple_prior_ISO <- 
  c(set_prior("student_t(3, 0, 2.5)", class = "Intercept"), #Global Intercept
    set_prior("gamma(0.01, 0.01)", class = "phi"), #phi Intercept with identity link
    set_prior("logistic(0, 1)", class = "Intercept", dpar = "zi" ), #zi Intercept
    set_prior("student_t(3, 0, 2.5)", class = "sd"), #stdev parameters
    set_prior("lkj(2)", class = "cor"), #correlation prior
    set_prior("normal(0,1)", class = "b")) #slope prior 
```

```{r}
job::job({
full_simple_ISO <-
  brm(data = df,
      formula = bf(irrfrac ~ 1 +  + eight_regions + yearcount + eight_regions:yearcount + (1 + yearcount|ISO),
                   zi ~ 1 + eight_regions + yearcount + eight_regions:yearcount + (1 + yearcount|ISO)),
      family = zero_inflated_beta(link = "logit", 
                                  link_phi = "identity", 
                                  link_zi = "logit"),
      prior = full_simple_prior_ISO,
      control = list(adapt_delta = 0.95, 
                       max_treedepth=11),
      sample_prior = "yes",
      inits = "0",
      cores = ncore, 
      threads = threading(2),
      backend = "cmdstanr",
      iter = niter, 
      chains = nchain,
      seed = seed,
      file = paste0(folder,"ISO_simple_full_5000.Rds"))
})
```
5000: Total execution time: 1236.6 seconds. 2 chains, 2 cores, threading(2)
```{r}
print(full_simple_ISO)
```


```{r}
posterior_samples(full_simple_ISO, pars = "b_") %>% 
  mutate(across(.cols = starts_with("b_phi"), exp)) %>% 
  mutate(across(.cols = c(- starts_with("b_phi")), plogis)) %>% 
  posterior_summary() %>% 
  as.data.frame() 
```

```{r}
conditional_effects(full_simple_ISO)
```

```{r}
pp_check(full_simple_ISO)
```

