

Packages

```{r}
library(dplyr)
library(tidyr)
library(brms)
library(purrr)
library(ggplot2)
library(GGally)
library(parallel)
library(kableExtra)
```








# Data

```{r load_data}
aei_small <- read.csv("/Volumes/RachelExternal/Thesis/Data_and_Plots/aei_small.csv")
aei_small_std <- read.csv("/Volumes/RachelExternal/Thesis/Data_and_Plots/aei_small_std.csv")
#aei_std <- read.csv("/p/projects/open/ledig/thesis/aei_std.csv")

aei_small_std <- aei_small_std %>% 
  mutate(across(c(4:8, 20:22), as.factor))


#for ease of life
aei_smaller <- aei_small %>% sample_n(5000) #%>% filter(ISO %in% sample(levels(ISO),3))
```

```{r}
#perparing to log transform
aei_smaller <- 
  aei_smaller %>% 
  mutate(popdens = popdens + 1, 
         dist = dist + 1,
         precip = precip + 1,
         humid = humid + 1,
         medinc = medinc + 1)


normalized<-function(y) {
  x<-y[!is.na(y)]
  x<-(x - min(x)) / (max(x) - min(x))
  y[!is.na(y)]<-x
  return(y)
}

centered <-function(y) {
  x<-y[!is.na(y)]
  x<-(x - mean(x)) / (sd(x)*2)
  y[!is.na(y)]<-x
  return(y)
}


aei_smaller_std <-
  aei_smaller %>% 
  mutate(across(c(17:21,25,26,31), centered)) %>% 
  mutate(across(c(24), normalized)) %>% 
  mutate(across(c(2,3,6,27,29), as.factor)) %>% 
  select(lat, lon, id, ISO, Countryname, six_regions, eight_regions, yearcount, irrfrac,
         dist, medinc, precip, humid, rugged, crl_yld, ag_gdp, popdens, gdppc, 
         DD.regime,DD.category, Democracy)


summary(aei_smaller_std)
```


# `brm()` settings
```{r brmsettings}
df = aei_smaller_std
ncore = 2 #detectCores()
options(mc.cores = parallel::detectCores())
niter = 500
nchain = 2
seed = 2

#saving location
folder = "/Volumes/RachelExternal/Thesis/Thesis/final/fits/"
```



# Prior Vis


Knowing that the most complicated model formula I will use is the one below, visualize the priors.
```{r}
get_prior(data = df, 
          formula = bf(irrfrac ~ 1 + eight_regions + yearcount + precip + medinc + dist + rugged + popdens + gdppc + Democracy +
                     eight_regions:yearcount + precip:yearcount + medinc:yearcount + dist:yearcount + rugged:yearcount + popdens:yearcount + gdppc:yearcount + Democracy:yearcount + (1 + yearcount |ISO),
                   zi ~ 1 + dist + gdppc),
      family = zero_inflated_beta(link = "logit", 
                                  link_phi = "identity", 
                                  link_zi = "logit"))
```

## Beta

These priors have a logit link. i.e. to visualize them we need and inv logit.

### mu Intercept

```{r}
tibble(`Default Student T(3, 0, 2.5)` = rt(10000, 3, 0)*2.5) %>%  
  mutate(`plogis Outcome Scale` = plogis(`Default Student T(3, 0, 2.5)`)) %>% 
  pivot_longer(cols = c(`Default Student T(3, 0, 2.5)`, `plogis Outcome Scale`), 
               names_to = "Transformation", 
               values_to = "Distribution") %>% 
  ggplot() +
  geom_histogram(aes(x = Distribution, color = Transformation)) + 
  xlab(NULL) + facet_wrap(~Transformation, scales = "free") +
  ggtitle("Default Global Intercept Prior - Student T(3, 0, 2.5)")
```

### mu Slopes


```{r}
tibble(`Proper Normal(0, 1)` = rnorm(10000, 0, 1)) %>%  
  mutate(`plogis Outcome Scale` = plogis(`Proper Normal(0, 1)`)) %>% 
  pivot_longer(cols = c(`Proper Normal(0, 1)`, `plogis Outcome Scale`), 
               names_to = "Transformation", 
               values_to = "Distribution") %>% 
  ggplot() +
  geom_histogram(aes(x = Distribution, color = Transformation)) + 
  xlab(NULL) + facet_wrap(~factor(Transformation, levels=c('Proper Normal(0, 1)','plogis Outcome Scale')), scales = "free") +
  ggtitle("mu Slope Prior - Normal(0, 1)")
```

### phi

As I am not predicting phi in the model formula, it has an identity link, meaning that it can just be visualized. 


```{r}
tibble(`Default Gamma(0.01, 0.01)` = rgamma(10000, rate = 0.01, shape = 0.01)) %>%  #according to the stan documentation
  pivot_longer(cols = c(`Default Gamma(0.01, 0.01)`), 
               names_to = "Transformation", 
               values_to = "Distribution") %>% 
  ggplot() +
  geom_density(aes(x = Distribution, color = Transformation)) + 
  xlab(NULL) + xlim(-1,1) +
  ggtitle("Default Zi Intercept Prior - `Default Gamma(0.01, 0.01)`")
```

This is super tight.. I wonder if I am plotting it correctly.


## Bernoullii

These priors have a logit link. i.e. to visualize them we need and inv logit.

### Intercept

```{r}
tibble(`Default Logistic(0, 1)` = rlogis(10000, 0, 1)) %>%  
  mutate(`plogis Outcome Scale` = plogis(`Default Logistic(0, 1)`)) %>% 
  pivot_longer(cols = c(`Default Logistic(0, 1)`, `plogis Outcome Scale`), 
               names_to = "Transformation", 
               values_to = "Distribution") %>% 
  ggplot() +
  geom_histogram(aes(x = Distribution, color = Transformation)) + 
  xlab(NULL) + facet_wrap(~Transformation, scales = "free") +
  ggtitle("Default Zi Intercept Prior - `Default Logistic(0, 1)`")
```

### Slopes


```{r}
tibble(`Proper Normal(0, 1)` = rnorm(10000, 0, 1)) %>%  
  mutate(`plogis Outcome Scale` = plogis(`Proper Normal(0, 1)`)) %>% 
  pivot_longer(cols = c(`Proper Normal(0, 1)`, `plogis Outcome Scale`), 
               names_to = "Transformation", 
               values_to = "Distribution") %>% 
  ggplot() +
  geom_histogram(aes(x = Distribution, color = Transformation)) + 
  xlab(NULL) + facet_wrap(~factor(Transformation, levels=c('Proper Normal(0, 1)','plogis Outcome Scale')), scales = "free") +
  ggtitle("Zi Slope Prior - Normal(0, 1)")
```



## cor and sd



# Individual Country Change Model

run first model based on time series

```{r}
# nest data
by_country <-
  aei_small_std %>%
  group_by(Countryname) %>%
  nest()
```

use default priors here, they allow things to vary nicely on a per country basis. 
```{r}
job::job({
id_fits_time_zib <-
  brm(data = by_country$data[[1]],
      formula = irrfrac ~ 1 + yearcount,
      family = gaussian(), 
      control = list(adapt_delta = 0.95, 
                     max_treedepth=11),
      cores = ncore, 
      iter = niter, 
      chains = nchain,
      seed = seed,
      file = "/Volumes/RachelExternal/Thesis/Thesis/3_pretty_scripts_for_profs/Fits/final_fits/Country_fits_pred_intercept_years_small.Rds")
})

# map to all other countries
job::job({
  yearcount_fits <- 
  by_country %>%
  mutate(model = map(data, ~update(id_fits_time_zib, newdata = ., seed = seed)))
  
  
#save nested df for later analysis on local machine
saveRDS(yearcount_fits, "/Volumes/RachelExternal/Thesis/Thesis/3_pretty_scripts_for_profs/Fits/final_fits/Country_fits_pred_intercept_years_full.Rds")
  })


```


# Simple model

## Unconditional Means Model ISO

```{r}
get_prior(data = df,
      formula = bf(irrfrac ~ 1 + (1|ISO), 
                   zi ~ 1 + (1|ISO)),
      family = zero_inflated_beta(link = "logit", 
                                  link_phi = "identity", 
                                  link_zi = "logit"))
```
Using default priors here. 
```{r}
uncond_means_prior_ISO <- 
  c(set_prior("student_t(3, 0, 2.5)", class = "Intercept"), #Global Intercept
    set_prior("gamma(0.01, 0.01)", class = "phi"), #phi Intercept with identity link
    set_prior("logistic(0, 1)", class = "Intercept", dpar = "zi" ), #zi Intercept
    set_prior("student_t(3, 0, 2.5)", class = "sd")) #stdev parameters
```


```{r}
job::job({
uncond_means_ISO <-
  brm(data = df,
      formula = bf(irrfrac ~ 1 + (1|ISO),
                   zi ~ 1 + (1|ISO)),
      family = zero_inflated_beta(link = "logit", 
                                  link_phi = "identity", 
                                  link_zi = "logit"),
      prior = uncond_means_prior_ISO,
      control = list(adapt_delta = 0.99, 
                       max_treedepth=11),
      sample_prior = "yes",
      cores = ncore, 
      iter = niter, 
      chains = nchain,
      seed = ncore,
      threads = threading(2),
      backend = "cmdstanr",
      file = paste0(folder,"ISO_simple_uncondmeans_5000.Rds"))
})
```

5000 samples: 210.0 seconds. 2 chains, 2 cores, threading(2)

```{r}
print(uncond_means_ISO)
```


```{r}
posterior_samples(uncond_means_ISO, pars = "b_") %>% 
  mutate(across(.cols = starts_with("b_phi"), exp)) %>% 
  mutate(across(.cols = c(- starts_with("b_phi")), plogis)) %>% 
  posterior_summary() %>% 
  as.data.frame() 
```

```{r}
pp_check(uncond_means_ISO)
```



## Unconditional Growth Model ISO



```{r}
get_prior(data = df,
      formula = bf(irrfrac ~ 1 + yearcount + (1 + yearcount|ISO),
                   zi ~ 1 + yearcount + (1 + yearcount|ISO)),
      family = zero_inflated_beta(link = "logit", 
                                  link_phi = "identity", 
                                  link_zi = "logit"))
```


Using default priors here. 
```{r}
uncond_growth_prior_ISO <- 
  c(set_prior("student_t(3, 0, 2.5)", class = "Intercept"), #Global Intercept
    set_prior("gamma(0.01, 0.01)", class = "phi"), #phi Intercept with identity link
    set_prior("logistic(0, 1)", class = "Intercept", dpar = "zi" ), #zi Intercept
    set_prior("student_t(3, 0, 2.5)", class = "sd"), #stdev parameters
    set_prior("lkj(2)", class = "cor"), #correlation prior
    set_prior("normal(0,1)", class = "b")) #slope prior 
```


```{r}
job::job({
uncond_growth_ISO <-
  brm(data = df,
      formula = bf(irrfrac ~ 1 + yearcount + (1 + yearcount|ISO),
                   zi ~ 1 + yearcount + (1 + yearcount|ISO)),
      family = zero_inflated_beta(link = "logit", 
                                  link_phi = "identity", 
                                  link_zi = "logit"),
      prior = uncond_growth_prior_ISO,
        control = list(adapt_delta = 0.95, 
                       max_treedepth=11),
      sample_prior = "yes",
      cores = ncore, 
      iter = niter, 
      chains = nchain,
      threads = threading(2),
      backend = "cmdstanr",
      inits = "0",
      seed = seed,
      file = paste0(folder,"ISO_simple_uncondgrowth_5000.Rds"))
})
```

5000: Total execution time: 1023.9 seconds. 2 chains, 2 cores, threading(2)

```{r}
summary(uncond_growth_ISO)
```

```{r}
posterior_samples(uncond_growth_ISO, pars = "b_") %>% 
  mutate(across(.cols = starts_with("b_phi"), exp)) %>% 
  mutate(across(.cols = c(- starts_with("b_phi")), plogis)) %>% 
  posterior_summary() %>% 
  as.data.frame() 
```
```{r}
conditional_effects(uncond_growth_ISO)
```

```{r}
pp_check(uncond_growth_ISO)
```

## Full Simple model ISO


```{r}
get_prior(data = df,
      formula = bf(irrfrac ~ 1 +  + eight_regions + yearcount + eight_regions:yearcount + (1 + yearcount|ISO),
                   zi ~ 1 + eight_regions + yearcount + eight_regions:yearcount + (1 + yearcount|ISO)),
      family = zero_inflated_beta(link = "logit", 
                                  link_phi = "identity", 
                                  link_zi = "logit"))
```

```{r}
full_simple_prior_ISO <- 
  c(set_prior("student_t(3, 0, 2.5)", class = "Intercept"), #Global Intercept
    set_prior("gamma(0.01, 0.01)", class = "phi"), #phi Intercept with identity link
    set_prior("logistic(0, 1)", class = "Intercept", dpar = "zi" ), #zi Intercept
    set_prior("student_t(3, 0, 2.5)", class = "sd"), #stdev parameters
    set_prior("lkj(2)", class = "cor"), #correlation prior
    set_prior("normal(0,1)", class = "b")) #slope prior 
```

```{r}
job::job({
full_simple_ISO <-
  brm(data = df,
      formula = bf(irrfrac ~ 1 +  + eight_regions + yearcount + eight_regions:yearcount + (1 + yearcount|ISO),
                   zi ~ 1 + eight_regions + yearcount + eight_regions:yearcount + (1 + yearcount|ISO)),
      family = zero_inflated_beta(link = "logit", 
                                  link_phi = "identity", 
                                  link_zi = "logit"),
      prior = full_simple_prior_ISO,
      control = list(adapt_delta = 0.95, 
                       max_treedepth=11),
      sample_prior = "yes",
      inits = "0",
      cores = ncore, 
      threads = threading(2),
      backend = "cmdstanr",
      iter = niter, 
      chains = nchain,
      seed = seed,
      file = paste0(folder,"ISO_simple_full_5000.Rds"))
})
```
5000: Total execution time: 1236.6 seconds. 2 chains, 2 cores, threading(2)
```{r}
print(full_simple_ISO)
```


```{r}
posterior_samples(full_simple_ISO, pars = "b_") %>% 
  mutate(across(.cols = starts_with("b_phi"), exp)) %>% 
  mutate(across(.cols = c(- starts_with("b_phi")), plogis)) %>% 
  posterior_summary() %>% 
  as.data.frame() 
```

```{r}
conditional_effects(full_simple_ISO)
```

```{r}
pp_check(full_simple_ISO)
```


# Complex model 


## Biophysical and Sociological

Unconditional means and growth models are the same as above. 

### Full model

```{r}
get_prior(data = df, 
          formula = bf(irrfrac ~ 1 + eight_regions + yearcount + precip + medinc + dist + rugged + popdens + gdppc + Democracy +
                     eight_regions:yearcount + precip:yearcount + medinc:yearcount + dist:yearcount + rugged:yearcount + popdens:yearcount + gdppc:yearcount + Democracy:yearcount + (1 + yearcount |ISO),
                   zi ~ 1 + dist + gdppc),
      family = zero_inflated_beta(link = "logit", 
                                  link_phi = "identity", 
                                  link_zi = "logit"))
```

```{r}
biosoc_full_prior_ISO <- 
  c(set_prior("student_t(3, 0, 2.5)", class = "Intercept"), #Global Intercept
    set_prior("gamma(0.01, 0.01)", class = "phi"), #phi Intercept with identity link
    set_prior("logistic(0, 1)", class = "Intercept", dpar = "zi" ), #zi Intercept
    set_prior("student_t(3, 0, 2.5)", class = "sd"), #stdev parameters
    set_prior("lkj(2)", class = "cor"), #correlation prior
    set_prior("normal(0,1)", class = "b"), 
    set_prior("normal(0,1)", class = "b", dpar = "zi")) #slope prior 
```


```{r}
job::job({
  full_simple_ISO <-
  brm(data = df,
      formula = bf(irrfrac ~ 1 + eight_regions + yearcount + precip + medinc + dist + rugged + popdens + gdppc + Democracy +
                     eight_regions:yearcount + precip:yearcount + medinc:yearcount + dist:yearcount + rugged:yearcount + popdens:yearcount + gdppc:yearcount + Democracy:yearcount + (1 + yearcount |ISO),
                   zi ~ 1 + dist + gdppc),
      family = zero_inflated_beta(link = "logit", 
                                  link_phi = "identity", 
                                  link_zi = "logit"),
      prior = biosoc_full_prior_ISO,
        control = list(adapt_delta = 0.95, 
                       max_treedepth=11),
      sample_prior = "yes",
      cores = ncore, 
      iter = niter, 
      chains = nchain,
      seed = seed,
      inits = "0",
      threads = threading(2),
      backend = "cmdstanr",
      file = paste0(folder,"ISO_biosocio_full_5000.Rds"))
})
```
No way: Total execution time: 1668.6 seconds. 2 chains, 2 cores, 5000 obvs. Roughly a half hour.
```{r}
print(full_simple_ISO)
```

```{r}
posterior_samples(full_simple_ISO, pars = "b_") %>% 
  mutate(across(.cols = starts_with("b_phi"), exp)) %>% 
  mutate(across(.cols = c(- starts_with("b_phi")), plogis)) %>% 
  posterior_summary() %>% 
  as.data.frame() 
```

```{r}
conditional_effects(full_simple_ISO)
```

```{r}
kfold(full_simple_ISO, chains = 1)
```



## Other less mathematically rigorous models

### Bio only

```{r}
get_prior(data = df,
          formula = bf(irrfrac ~ 1 + yearcount + precip + medinc + dist + rugged + eight_regions + (1 + yearcount|ISO),
                   zi ~ 1 + dist), 
          family = zero_inflated_beta(link = "logit", 
                                  link_phi = "identity", 
                                  link_zi = "logit"))
```


```{r}
bio_only_prior_ISO <- 
  c(set_prior("student_t(3, 0, 2.5)", class = "Intercept"), #Global Intercept
    set_prior("gamma(0.01, 0.01)", class = "phi"), #phi Intercept with identity link
    set_prior("logistic(0, 1)", class = "Intercept", dpar = "zi" ), #zi Intercept
    set_prior("student_t(3, 0, 2.5)", class = "sd"), #stdev parameters
    set_prior("lkj(2)", class = "cor"), #correlation prior
    set_prior("normal(0,1)", class = "b"), 
    set_prior("normal(0,1)", class = "b", dpar = "zi")) #slope prior 
```


```{r}
job::job({
bio_only_ISO <-
  brm(data = df,
      formula = bf(irrfrac ~ 1 + yearcount + precip + medinc + dist + rugged + eight_regions + (1 + yearcount|ISO),
                   zi ~ 1 + dist),
      family = zero_inflated_beta(link = "logit", 
                                  link_phi = "identity", 
                                  link_zi = "logit"),
      prior = bio_only_prior_ISO,
        control = list(adapt_delta = 0.95, 
                       max_treedepth=11),
      sample_prior = "yes",
      cores = ncore, 
      iter = niter, 
      chains = nchain,
      inits = "0",
      threads = threading(2), 
      backend = "cmdstanr",
      seed = seed,
      file = paste0(folder,"ISO_bioonly_5000.Rds"))
})
```

```{r}
print(bio_only_ISO)
```

```{r}
posterior_samples(bio_only_ISO, pars = "b_") %>% 
  mutate(across(.cols = starts_with("b_phi"), exp)) %>% 
  mutate(across(.cols = c(- starts_with("b_phi")), plogis)) %>% 
  posterior_summary() %>% 
  as.data.frame() 
```

```{r}
conditional_effects(bio_only_ISO)
```




### Biosoc, no interactions


```{r}
get_prior(data = df,
          formula = bf(irrfrac ~ 1 + yearcount + precip + medinc + dist + rugged + Democracy + gdppc + popdens + eight_regions + (1 + yearcount|ISO),
                   zi ~ 1 + dist + gdppc), 
          family = zero_inflated_beta(link = "logit", 
                                  link_phi = "identity", 
                                  link_zi = "logit"))
```


```{r}
biosoc_no_inter_prior_ISO <- 
  c(set_prior("student_t(3, 0, 2.5)", class = "Intercept"), #Global Intercept
    set_prior("gamma(0.01, 0.01)", class = "phi"), #phi Intercept with identity link
    set_prior("logistic(0, 1)", class = "Intercept", dpar = "zi" ), #zi Intercept
    set_prior("student_t(3, 0, 2.5)", class = "sd"), #stdev parameters
    set_prior("lkj(2)", class = "cor"), #correlation prior
    set_prior("normal(0,1)", class = "b"), 
    set_prior("normal(0,1)", class = "b", dpar = "zi")) #slope prior 
```


```{r}
job::job({
biosoc_no_inter_ISO <-
  brm(data = df,
      formula = bf(irrfrac ~ 1 + yearcount + precip + medinc + dist + rugged + Democracy + gdppc + popdens + eight_regions + (1 + yearcount|ISO),
                   zi ~ 1 + dist + gdppc),
      family = zero_inflated_beta(link = "logit", 
                                  link_phi = "identity", 
                                  link_zi = "logit"),
      prior = biosoc_no_inter_prior_ISO,
        control = list(adapt_delta = 0.95, 
                       max_treedepth=11),
      sample_prior = "yes",
      cores = ncore, 
      iter = niter, 
      chains = nchain,
      inits = "0",
      threads = threading(2), 
      backend = "cmdstanr",
      seed = seed,
      file = paste0(folder,"ISO_biosoc_no_inter_5000.Rds"))
})
```

```{r}
print(biosoc_no_inter_ISO)
```

```{r}
posterior_samples(biosoc_no_inter_ISO, pars = "b_") %>% 
  mutate(across(.cols = starts_with("b_phi"), exp)) %>% 
  mutate(across(.cols = c(- starts_with("b_phi")), plogis)) %>% 
  posterior_summary() %>% 
  as.data.frame() 
```

```{r}
conditional_effects(biosoc_no_inter_ISO)
```

```{r}
pp_check(biosoc_no_inter_ISO)
```


### Biosoc, more varying slopes


```{r}
get_prior(data = df,
          formula = bf(irrfrac ~ 1 + yearcount + precip + medinc + dist + rugged + Democracy + gdppc + popdens + eight_regions +
                         (1 + yearcount + Democracy + gdppc + popdens|ISO),
                   zi ~ 1 + dist + gdppc), 
          family = zero_inflated_beta(link = "logit", 
                                  link_phi = "identity", 
                                  link_zi = "logit"))
```


```{r}
biosoc_v_slopes_prior_ISO <- 
  c(set_prior("student_t(3, 0, 2.5)", class = "Intercept"), #Global Intercept
    set_prior("gamma(0.01, 0.01)", class = "phi"), #phi Intercept with identity link
    set_prior("logistic(0, 1)", class = "Intercept", dpar = "zi" ), #zi Intercept
    set_prior("student_t(3, 0, 2.5)", class = "sd"), #stdev parameters
    set_prior("lkj(2)", class = "cor"), #correlation prior
    set_prior("normal(0,1)", class = "b"), 
    set_prior("normal(0,1)", class = "b", dpar = "zi")) #slope prior 
```


```{r}
job::job({
biosoc_v_slopes_ISO <-
  brm(data = df,
      formula = bf(irrfrac ~ 1 + yearcount + precip + medinc + dist + rugged + Democracy + gdppc + popdens + eight_regions +
                         (1 + yearcount + Democracy + gdppc + popdens|ISO),
                   zi ~ 1 + dist + gdppc),
      family = zero_inflated_beta(link = "logit", 
                                  link_phi = "identity", 
                                  link_zi = "logit"),
      prior = biosoc_v_slopes_prior_ISO,
        control = list(adapt_delta = 0.95, 
                       max_treedepth=11),
      sample_prior = "yes",
      cores = ncore, 
      iter = niter, 
      chains = nchain,
      inits = "0",
      threads = threading(2), 
      backend = "cmdstanr",
      seed = seed,
      file = paste0(folder,"ISO_biosoc_v_slopes_5000.Rds"))
})
```

```{r}
print(biosoc_v_slopes_ISO)
```

```{r}
posterior_samples(biosoc_v_slopes_ISO, pars = "b_") %>% 
  mutate(across(.cols = starts_with("b_phi"), exp)) %>% 
  mutate(across(.cols = c(- starts_with("b_phi")), plogis)) %>% 
  posterior_summary() %>% 
  as.data.frame() 
```

```{r}
conditional_effects(biosoc_v_slopes_ISO)
```



#Experiments with Syntax


## Correlated grouping terms

```{r}
get_prior(data = df,
      formula = bf(irrfrac ~ 1 + (1|i|ISO), 
                   zi ~ 1 + (1|i|ISO)),
      family = zero_inflated_beta(link = "logit", 
                                  link_phi = "identity", 
                                  link_zi = "logit"))
```
Using default priors here. 
```{r}
simple_cor_gterms_prior_ISO <- 
  c(set_prior("student_t(3, 0, 2.5)", class = "Intercept"), #Global Intercept
    set_prior("gamma(0.01, 0.01)", class = "phi"), #phi Intercept with identity link
    set_prior("logistic(0, 1)", class = "Intercept", dpar = "zi" ), #zi Intercept
    set_prior("student_t(3, 0, 2.5)", class = "sd"), #stdev parameters
    set_prior("lkj(2)", class = "cor"))
```


```{r}
job::job({
simple_cor_gterms_ISO <-
  brm(data = df,
      formula = bf(irrfrac ~ 1 + (1|i|ISO), 
                   zi ~ 1 + (1|i|ISO)),
      family = zero_inflated_beta(link = "logit", 
                                  link_phi = "identity", 
                                  link_zi = "logit"),
      prior = simple_cor_gterms_prior_ISO,
      control = list(adapt_delta = 0.99, 
                       max_treedepth=11),
      sample_prior = "yes",
      cores = ncore, 
      iter = niter, 
      chains = nchain,
      seed = ncore,
      threads = threading(2),
      backend = "cmdstanr",
      file = paste0(folder,"ISO_simple_cor_gterms_5000.Rds"))
})
```

5000 samples: Total execution time: 328.9 seconds. 2 chains, 2 cores, threading(2)

```{r}
print(simple_cor_gterms_ISO)
```


```{r}
posterior_samples(simple_cor_gterms_ISO, pars = "b_") %>% 
  mutate(across(.cols = starts_with("b_phi"), exp)) %>% 
  mutate(across(.cols = c(- starts_with("b_phi")), plogis)) %>% 
  posterior_summary() %>% 
  as.data.frame() 
```

```{r}
pp_check(simple_cor_gterms_ISO)
```



## Unconditional Growth Model ISO



```{r}
get_prior(data = df,
      formula = bf(irrfrac ~ 1 + yearcount + (1 + yearcount|i|ISO),
                   zi ~ 1 + yearcount + (1 + yearcount|i|ISO)),
      family = zero_inflated_beta(link = "logit", 
                                  link_phi = "identity", 
                                  link_zi = "logit"))
```


Using default priors here. 
```{r}
yrct_cor_gterms_prior_ISO <- 
  c(set_prior("student_t(3, 0, 2.5)", class = "Intercept"), #Global Intercept
    set_prior("gamma(0.01, 0.01)", class = "phi"), #phi Intercept with identity link
    set_prior("logistic(0, 1)", class = "Intercept", dpar = "zi" ), #zi Intercept
    set_prior("student_t(3, 0, 2.5)", class = "sd"), #stdev parameters
    set_prior("lkj(2)", class = "cor"), #correlation prior
    set_prior("normal(0,1)", class = "b")) #slope prior 
```


```{r}
job::job({
yrct_cor_gterms_ISO <-
  brm(data = df,
      formula = bf(irrfrac ~ 1 + yearcount + (1 + yearcount|i|ISO),
                   zi ~ 1 + yearcount + (1 + yearcount|i|ISO)),
      family = zero_inflated_beta(link = "logit", 
                                  link_phi = "identity", 
                                  link_zi = "logit"),
      prior = yrct_cor_gterms_prior_ISO,
        control = list(adapt_delta = 0.95, 
                       max_treedepth=11),
      sample_prior = "yes",
      cores = ncore, 
      iter = niter, 
      chains = nchain,
      threads = threading(2),
      backend = "cmdstanr",
      inits = "0",
      seed = seed,
      file = paste0(folder,"ISO_yrct_cor_gterms_5000.Rds"))
})
```

5000:  2 chains, 2 cores, threading(2)

```{r}
summary(uncond_growth_ISO)
```

```{r}
posterior_samples(uncond_growth_ISO, pars = "b_") %>% 
  mutate(across(.cols = starts_with("b_phi"), exp)) %>% 
  mutate(across(.cols = c(- starts_with("b_phi")), plogis)) %>% 
  posterior_summary() %>% 
  as.data.frame() 
```
```{r}
conditional_effects(uncond_growth_ISO)
```

```{r}
pp_check(uncond_growth_ISO)
```

## Full Simple model ISO


```{r}
get_prior(data = df,
      formula = bf(irrfrac ~ 1 +  + eight_regions + yearcount + eight_regions:yearcount + (1 + yearcount|ISO),
                   zi ~ 1 + eight_regions + yearcount + eight_regions:yearcount + (1 + yearcount|ISO)),
      family = zero_inflated_beta(link = "logit", 
                                  link_phi = "identity", 
                                  link_zi = "logit"))
```

```{r}
full_simple_prior_ISO <- 
  c(set_prior("student_t(3, 0, 2.5)", class = "Intercept"), #Global Intercept
    set_prior("gamma(0.01, 0.01)", class = "phi"), #phi Intercept with identity link
    set_prior("logistic(0, 1)", class = "Intercept", dpar = "zi" ), #zi Intercept
    set_prior("student_t(3, 0, 2.5)", class = "sd"), #stdev parameters
    set_prior("lkj(2)", class = "cor"), #correlation prior
    set_prior("normal(0,1)", class = "b")) #slope prior 
```

```{r}
job::job({
full_simple_ISO <-
  brm(data = df,
      formula = bf(irrfrac ~ 1 +  + eight_regions + yearcount + eight_regions:yearcount + (1 + yearcount|ISO),
                   zi ~ 1 + eight_regions + yearcount + eight_regions:yearcount + (1 + yearcount|ISO)),
      family = zero_inflated_beta(link = "logit", 
                                  link_phi = "identity", 
                                  link_zi = "logit"),
      prior = full_simple_prior_ISO,
      control = list(adapt_delta = 0.95, 
                       max_treedepth=11),
      sample_prior = "yes",
      inits = "0",
      cores = ncore, 
      threads = threading(2),
      backend = "cmdstanr",
      iter = niter, 
      chains = nchain,
      seed = seed,
      file = paste0(folder,"ISO_simple_full_5000.Rds"))
})
```
5000: Total execution time: 1236.6 seconds. 2 chains, 2 cores, threading(2)
```{r}
print(full_simple_ISO)
```


```{r}
posterior_samples(full_simple_ISO, pars = "b_") %>% 
  mutate(across(.cols = starts_with("b_phi"), exp)) %>% 
  mutate(across(.cols = c(- starts_with("b_phi")), plogis)) %>% 
  posterior_summary() %>% 
  as.data.frame() 
```

```{r}
conditional_effects(full_simple_ISO)
```

```{r}
pp_check(full_simple_ISO)
```

