The purpose of this script is to create the data and STAN code necessary to run grid cell models on PIK's cluster. 

Adjust your expectations accordingly. 

```{r packages}
library(dplyr)
library(rstan)
library(brms)
```

Load data. 
This data is the 5 year time step data. Computationally it doesn't make sense to run all time steps. 
```{r load_data}
aei_small_std <- read.csv("/Volumes/RachelExternal/Thesis/Data_and_Plots/aei_small_std.csv")
#aei_std <- read.csv("/p/projects/open/ledig/thesis/aei_std.csv")

summary(aei_small_std)
```

# id Models
These models follow the logic that each individual cell has its own trajectory in terms of growth and paramaterizaiton.


## Unconditional Means Model ID
Unconditional means models aim to look at general effects across grid cells without time as a component.


Save the data as STAN data. i.e. a list. 
```{r stan_data_means_id}
uncond_means_data_id <- 
  make_standata(data = aei_small_std,
                formula = bf(irrfrac ~ 1 + (1|id),
                             zi ~ 1 + (1|id)),
                family = zero_inflated_beta())

saveRDS(uncond_means_data_id, file = "/Volumes/RachelExternal/Thesis/Thesis/3_pretty_scripts_for_profs/PIK/data/uncond_means_data_id.Rdata")

```

Look at priors for this model
```{r}
get_prior(data = aei_small_std,
            formula = bf(irrfrac ~ 1 + (1|id),
                             zi ~ 1 + (1|id)),
                family = zero_inflated_beta())
```
Student T will work fine for the intercept. It creates a nice U shaped, perfect for a ZIB model. 
```{r}
uncond_means_priors_id <- 
  set_prior()
```



```{r}
uncond_means_code_id <-
  make_stancode(data = aei_small_std,
                formula = bf(irrfrac ~ 1 + (1|id),
                             phi ~ 1 + (1|id),
                             zi ~ 1 + (1|id)),
                family = zero_inflated_beta())
```



```{r}
uncond_means_stan_id <-
  stan(model_code = uncond_means_code_id, 
       data = uncond_means_data_id, 
       iter = 2000, 
       chains = 4,
       cores = 4,
       seed = 17,
       control = list(adapt_delta = 0.9, 
                      max_treedepth=11))

saveRDS(uncond_means_stan_id, "/p/projects/open/ledig/thesis/fits/stan_uncond_means_id.rds")
```




## Unconditional Growth Model ID
aims to look at general effects across cells and over time prior to adding other preds

```{r}
uncond_growth_data_id <- 
  make_standata(data = aei_small_std,
                formula = bf(irrfrac ~ 1 + yearcount + (1 + yearcount|id),
                             zi ~ 1 + yearcount + (1 + yearcount|id)),
                family = zero_inflated_beta())

saveRDS(uncond_growth_data_id, file = "/Volumes/RachelExternal/Thesis/Thesis/3_pretty_scripts_for_profs/PIK/data/uncond_growth_data_id.Rdata")

```

```{r}
uncond_growth_code_id <-
  make_stancode(data = aei_small_std,
                formula = bf(irrfrac ~ 1 + yearcount + (1 + yearcount|id),
                             zi ~ 1 + yearcount + (1 + yearcount|id)),
                family = zero_inflated_beta(),
                prior = c(prior(normal(0, 100), class = b)),
                sample_prior = "only")


```

```{r}
uncond_growth_stan_id <-
  stan(model_code = uncond_growth_code_id,
       data = uncond_growth_data_id,
       iter = 2000, 
       chains = 4,
       cores = 4,
       seed = 17,
       control = list(adapt_delta = 0.9, 
                      max_treedepth=11))

saveRDS(uncond_growth_stan_id, "/p/projects/open/ledig/thesis/fits/stan_uncond_growth_id.rds")

```


## Full Simple model ID 

```{r}
full_simple_data_id <- 
  make_standata(data = aei_small_std,
                formula = bf(irrfrac ~ 0 + Intercept + yearcount + ISO + ISO:yearcount + (1 + yearcount|id),
                                       zi ~ 0 + Intercept + yearcount + ISO + ISO:yearcount + (1 + yearcount|id)),
                family = zero_inflated_beta())

saveRDS(full_simple_data_id, file = "/Volumes/RachelExternal/Thesis/Thesis/3_pretty_scripts_for_profs/PIK/data/full_simple_data_id.Rdata")
```

```{r}
full_simple_code_id <-
  make_stancode(data = aei_small_std,
                formula = formula = bf(irrfrac ~ 0 + Intercept + yearcount + ISO + ISO:yearcount + (1 + yearcount|id),
                                       zi ~ 0 + Intercept + yearcount + ISO + ISO:yearcount + (1 + yearcount|id)),
                family = zero_inflated_beta(),
                prior = c(prior(normal(0, 100), class = b)),
                sample_prior = "only")
```


```{r}
full_simple_stan_id <-
  stan(model_code = full_simple_code_id,
       data = full_simple_data_id,
       iter = 2000, 
       chains = 4,
       cores = 4,
       seed = 17,
       control = list(adapt_delta = 0.9, 
                      max_treedepth=11))

saveRDS(full_simple_stan_id, "/p/projects/open/ledig/thesis/fits/stan_full_simple_id.rds")

```



# ISO Models

##Unconditional Means Model ISO


```{r}
uncond_means_data_ISO <- 
  make_standata(data = aei_small_std,
                formula = bf(irrfrac ~ 1 + (1|ISO),
                             zi ~ 1 + (1|ISO)),
                family = zero_inflated_beta())

saveRDS(uncond_means_data_ISO, file = "/Volumes/RachelExternal/Thesis/Thesis/3_pretty_scripts_for_profs/PIK/data/uncond_means_data_ISO.Rdata")
```


```{r}
uncond_means_code_ISO <-
  make_stancode(data = aei_small_std,
                formula = bf(irrfrac ~ 1 + (1|ISO),
                             zi ~ 1 + (1|ISO)),
                family = zero_inflated_beta(),
                sample_prior = "only")
```


```{r}
uncond_means_stan_ISO <-
  stan(model_code = uncond_means_code_ISO, 
       data = uncond_means_data_ISO, 
       iter = 2000, 
       chains = 4,
       cores = 4,
       seed = 17,
       control = list(adapt_delta = 0.9, 
                      max_treedepth=11))

saveRDS(uncond_means_stan_id, "/p/projects/open/ledig/thesis/fits/stan_uncond_means_ISO.rds")
```





## Unconditional Growth Model ISO

```{r}
uncond_growth_data_ISO <- 
  make_standata(data = aei_small_std,
                formula = bf(irrfrac ~ 1 + yearcount + (1 + yearcount|ISO),
                             zi ~ 1 + yearcount + (1 + yearcount|ISO)),
                family = zero_inflated_beta())

saveRDS(uncond_growth_data_ISO, file = "/Volumes/RachelExternal/Thesis/Thesis/3_pretty_scripts_for_profs/PIK/data/uncond_growth_data_ISO.Rdata")
```


```{r}
uncond_growth_code_ISO <-
  make_stancode(data = aei_small_std,
                formula = bf(irrfrac ~ 1 + yearcount + (1 + yearcount|ISO),
                             zi ~ 1 + yearcount + (1 + yearcount|ISO)),
                family = zero_inflated_beta(),
                prior = c(prior(normal(0, 100), class = b)),
                sample_prior = "only")

```


```{r}
uncond_growth_stan_ISO <-
  stan(model_code = uncond_growth_code_ISO,
       data = uncond_growth_data_ISO,
       iter = 2000, 
       chains = 4,
       cores = 4,
       seed = 17,
       control = list(adapt_delta = 0.9, 
                      max_treedepth=11))

saveRDS(uncond_growth_stan_id, "/p/projects/open/ledig/thesis/fits/stan_uncond_growth_ISO.rds")
```




## Full Simple model ISO

```{r}
full_simple_data_ISO <- 
  make_standata(data = aei_small_std,
                formula = bf(irrfrac ~ 0 + Intercept + yearcount + six_regions + six_regions:yearcount + (1 + yearcount|ISO),
                                       zi ~ 0 + Intercept + yearcount + six_regions + six_regions:yearcount + (1 + yearcount|ISO)),
                family = zero_inflated_beta())

saveRDS(full_simple_data_ISO, file = "/Volumes/RachelExternal/Thesis/Thesis/3_pretty_scripts_for_profs/PIK/data/full_simple_data_ISO.Rdata")
```

```{r}
full_simple_code_ISO <-
  make_stancode(data = aei_small_std,
                formula = bf(irrfrac ~ 0 + Intercept + yearcount + six_regions + six_regions:yearcount + (1 + yearcount|ISO),
                                       zi ~ 0 + Intercept + yearcount + six_regions + six_regions:yearcount + (1 + yearcount|ISO)),
                family = zero_inflated_beta(),
                prior = c(prior(normal(0, 100), class = b)),
                sample_prior = "only")
```



```{r}
full_simple_stan_id <-
  stan(model_code = full_simple_code_ISO,
       data = full_simple_data_ISO,
       iter = 2000, 
       chains = 4,
       cores = 4,
       seed = 17,
       control = list(adapt_delta = 0.9, 
                      max_treedepth=11))

saveRDS(full_simple_stan_id, "/p/projects/open/ledig/thesis/fits/stan_full_simple_ISO.rds")
```






# Complex model

 without interactions, at lower levels they are easier to sort out but here its too complex
```{r}
full_complex_data <- 
  make_standata(data = aei_small_std,
                formula = bf(irrfrac ~ 0 + Intercept + yearcount + DD.regime + dist + popdens + pet + gdppc + rugged + (1 + yearcount + popdens + DD.regime + gdppc|id),
                             zi ~ 0 + Intercept + yearcount + DD.regime + dist + popdens + pet + gdppc + rugged + (1 + yearcount + popdens + DD.regime + gdppc|id)),
                family = zero_inflated_beta())

saveRDS(full_complex_data, file = "/Volumes/RachelExternal/Thesis/Thesis/3_pretty_scripts_for_profs/PIK/data/full_complex_data.Rdata")
```

```{r}
full_complex_code <-
  make_stancode(data = aei_small_std,
                formula = bf(irrfrac ~ 0 + Intercept + yearcount + DD.regime + dist + popdens + pet + gdppc + rugged + (1 + yearcount + popdens + DD.regime + gdppc|id),
                             zi ~ 0 + Intercept + yearcount + DD.regime + dist + popdens + pet + gdppc + rugged + (1 + yearcount + popdens + DD.regime + gdppc|id)),
                family = zero_inflated_beta(),
                prior = c(prior(normal(0, 100), class = b)),
                sample_prior = "only")


```



```{r}
full_complex_stan <-
  stan(model_code = full_complex_code,
       data = full_complex_data,
       iter = 2000, 
       chains = 4,
       cores = 4,
       seed = 17,
       control = list(adapt_delta = 0.9, 
                      max_treedepth=11))

saveRDS(full_complex_stan, "/p/projects/open/ledig/thesis/fits/stan_full_complex.rds")
```







